---
title: "00-01-Start"
author: "Michael Arbeider"
date: "2026-03-06"
---

Welcome to the Fraser River and Interior Area Sockeye Calibration Report Code.
This code was written for the Technical Document, "Assessment and redevelopment of visual survey expansion factors and a summary of current interpolation methods for Fraser River Sockeye (Oncorhynchus nerka) spawner estimates."

The code is separated into 5 main sections.
- 00 Series are the introduction and functions that are referenced in later
sections.
- 01 Series are model development and running of the leave-one-out cross-
validations.
- 02 Calculates the model performance metrics for each set-level, resulting in
rank-sets.
- 03 Series applies the calibration algorithm to the training data and compares it to 
the default 1.8 expansion factor.
- 04 Is the calibration application "tool" and test application to the 2024 
data.

If you are running the code for the very first time, this mark-down must be run
because it initiates a file that is then populated throughout the 01 series.

Otherwise, this Start file is primarily to show some simple visualizations of
the data.

# Set-up
## Packages
```{r}
rm(list = ls())
library(tidyverse) # most %>% 
theme_set(theme_bw()) # ggplot
library(scales) # scale_x_continuous
library(patchwork) # for multi plot ggplot graphs
source("00-Functions.R") # scale_x_reordered, rank.summary(), round2()
```

## Data
This is the paired calibration database.
There is an excel version in the data folder with a readme tab that defines
the fields.
```{r}
cal <- readRDS(file = "r_objects/sockeye_calibration_data.rds")
glimpse(cal)
```

## Dataframe for Model Specs for Later
This dataframe will be populated during the 01 series with model specs
It is later referenced in 03-01-Model-Code to create a function that 
automatically applies model specific transformations.
```{r}
modout <- "./r_objects/models/model_specifications.rds"

specifications_out <- data.frame(model = c(""),
                                 target = c(""),
                                 order = c(0),
                                 type = c(""),
                                 transform = c(0))

saveRDS(specifications_out, file = modout)
```

# Summaries
The following are a few quick summaries used for tables in the report

## Samples Per Year
```{r}
sample.systems <- cal %>% 
  group_by(CU, gazetted_stream_name) %>% 
  summarize(n = n(),
            earliest = min(year),
            recent = max(year))

write.csv(sample.systems, file="./tables/00-Sample Size Summary Table - CUs and Systems.csv", row.names = F)

```

## Visual Count Ranges
```{r}
data.range <- cal %>% 
  group_by(size, water_clarity, lpe_method, substrate_shade) %>% 
  summarise(n = n(),
            median = median(lpe),
            min = min(lpe),
            max = max(lpe)) %>% 
  mutate_at(c(6:8), ~ round2(., 0))

write.csv(data.range, "./tables/00-Regression_LPE_Ranges_Full.csv", row.names = F)
saveRDS(data.range, "./r_objects/00-Regression_LPE_Ranges_FUll.rds")

data.range2 <- cal %>% 
  group_by(size, water_clarity, lpe_method) %>% 
  summarise(n = n(),
            min10 = min(lpe)*0.9/10,
            max10 = max(lpe)*1.1/10) %>% 
  mutate_at(c(5:6), ~ round2(., 0)*10)

write.csv(data.range2, "./tables/00-Regression_LPE_Ranges_Clarity_Method.csv", row.names = F)
saveRDS(data.range2, "./r_objects/00-Regression_LPE_Ranges_Clarity_Method.rds")

data.range3 <- cal %>% 
  group_by(size, lpe_method) %>% 
  summarise(n = n(),
            min10 = min(lpe)*0.9/10,
            max10 = max(lpe)*1.1/10) %>% 
  mutate_at(c(4:5), ~ round2(., 0)*10)

write.csv(data.range3, "./tables/00-Regression_LPE_Ranges.csv", row.names = F)
saveRDS(data.range3, "./r_objects/00-Regression_LPE_Ranges.rds")
```


# Plots
## Size Colours
https://stackoverflow.com/questions/57153428/r-plot-color-combinations-that-are-colorblind-accessible
```{r colour pallet}
size.cols = c("#E69F00", # Large
              "#56B4E9", # Medium (including the smalls)
              "#D55E00", # Small
              "#CC79A7") # X-Large
```

## HPE vs LPE By Size and Clarity
```{r}
plt.raw <- cal %>% 
  ggplot(aes(x = lpe, y = hpe, colour = size)) +
  geom_point() +
  labs(x = "Visual Count", 
       y = "High Accuracy Estimate") +
  scale_colour_manual(values = size.cols)

plt.log <- cal %>% 
  ggplot(aes(x = lnlpe.sc, y = log(hpe), colour = size)) +
  geom_point() +
  labs(x = "ln(Scaled - Visual Count)", 
       y = "ln(High Accuracy Estimate)") +
  scale_colour_manual(values = size.cols)

plt.raw + plt.log + plot_layout(guides = 'collect')

ggsave("figures/SK Calibration - HPE vs LPE with Size as Colour.png",
       units = "in", height = 4, width = 9, dpi = 150)
```

## Surveys through Time
```{r}
# By Conservation Unit
cal %>% 
  ggplot(aes(x = year, fill = CU)) +
  geom_bar(stat = "count") +
  scale_fill_manual(values = c("#000000","#004949","#009292","#ff6db6", "#ffb6db",
                               "#490092","#006ddb","#b66dff","#6db6ff", "#b6dbff", 
                               "#920000","#924900","#db6d00","#24ff24"))

ggsave("figures/SK Calibration - Count by Year and CU.png",
       units = "in", height = 5, width = 9, dpi = 96)

# By Management Unit
cal %>% 
  ggplot(aes(x = year, fill = run_timing_group)) +
  geom_bar(stat = "count") +
  ggokabeito::scale_fill_okabe_ito(name = "Management Unit") +
  labs(x = "Year", y = "Count of Paired Data")

ggsave("figures/SK Calibration - Count by Year and MU.png",
       units = "in", height = 5, width = 9, dpi = 96)
```

## Histogram by Size, Substrate & Clarity
```{r}
cal %>% 
  ggplot(aes(lpe, fill = size)) +
  geom_histogram() +
  scale_fill_manual(values = size.cols,
                      name = "Size") +
  facet_grid(substrate_shade~water_clarity) +
  geom_vline(xintercept = c(25000, 75000),
             colour = "grey", lty = 2) +
  scale_x_continuous(labels = label_comma()) +
  labs(x = "Visual Count", y = "Count",
       title = "Water Clarity by Substrate Shade and Size")

ggsave("figures/SK Calibration - Count by Size Substrate Clarity.png",
       units = "in", height = 6, width = 5.5, dpi = 96)
```

