---
title: "01-01-Mean-of-Ratios"
author: "Michael Arbeider"
date: "2026-03-06"
---

From the 2019 Calibration Report (Welch and Leaf 2019):
contingent on the criteria that escapements to individual streams would only be 
reassessed if they fell within a stream size/water clarity/assessment method 
strata having either a minimum of five data points with an error (CV) of 25% or 
less or having a total of 10 or more data points. 

*Major Difference between 01-02 and 01-01 is that 01-02 is in log-space
*01-01 is in un-transformed space.

# Set-up
## Packages
```{r}
rm(list = ls())
library(tidyverse) # most %>% 
theme_set(theme_bw()) # ggplot
library(patchwork) # for multi plot ggplot graphs
source("00-Functions.R") # scale_x_reordered, rank.summary(), round2()
```

## Data
```{r}
cal <-readRDS("r_objects/sockeye_calibration_data.rds")   

glimpse(cal)
```

## Specifications FILE
```{r}
modout <- "./r_objects/models/model_specifications.rds"
spec.out <- readRDS(modout)
```

## Base Plots
https://stackoverflow.com/questions/57153428/r-plot-color-combinations-that-are-colorblind-accessible
```{r colour pallet}
cal$size <- factor(cal$size, levels=c("Small", "Medium", "Large", "X-Large"), ordered=T)

size.cols = c("#D55E00", # V. Small
              "#56B4E9", # Medium (including the smalls)
              "#E69F00", # Large
              "#CC79A7") # X-Large
```

#1 Ratios
## Water Clarity
Only considering Method-Size-Clarity combinations of > 20 n
```{r}
cal.keep <- count(cal,
                  lpe_method, size, water_clarity)
```
```{r}
cal.ratios <- cal %>% 
  # creating ratios by size and water clarity
  group_by(lpe_method, size, water_clarity) %>%
  summarise(n = n(), # sample size
            Average.Pop = mean(hpe),
            MoR = mean(hpe/lpe), # MoR = Mean of Ratios
            MoR.Min = min(hpe/lpe),
            MoR.Max = max(hpe/lpe),
            MoR.SD = sqrt(var(hpe/lpe) + mean((hpe_error^2)*(1/lpe)^2)), # Table 3 in McPherson et al. 2000 https://www.arlis.org/docs/vol1/ADFG/FMS/FMS00-02.pdf
            MoR.CV = MoR.SD/MoR,
            q75 = quantile(hpe/lpe, 0.75),
            q25 = quantile(hpe/lpe, 0.25)) %>% 
  # needed to calculated interquartile range
  mutate(iqr = q75-q25,
         iqr.5 = iqr*1.5, # for calculating outliers
         q75.5 = round2(q75+iqr.5, 2), # whisker max
         q25.5 = round2(q25-iqr.5, 2), # whisker min
         # creating flag of when to use MoR
          useable.MoR = case_when(
    n > 20 & MoR.CV <= 0.50 ~ "Yes",
    TRUE ~ "No"
    ))

# filtering based on MoR selection criteria
cal.ratios.keep <- filter(cal.ratios,
                          useable.MoR == "Yes")
```

## Table Outputs
```{r}
table01 <- cal.ratios %>% 
  dplyr::select(lpe_method:MoR.CV) %>% 
  mutate(Average.Pop = round2(Average.Pop, 0),
         MoR = round2(MoR, 2),
         MoR.Min = round2(MoR.Min, 2),
         MoR.Max = round2(MoR.Max, 2),
         MoR.SD = round2(MoR.SD, 2),
         MoR.CV = round2(MoR.CV, 3)*100)

write.csv(table01, "./tables/01_01-Mean-of-Ratios-summary.csv")

outliers <- cal %>% 
  left_join(.,
            dplyr::select(cal.ratios,
                          lpe_method:water_clarity, q75.5, q25.5)) %>% 
  dplyr::select(sid:water_clarity, lpe_method, hpe, lpe, q75.5, q25.5) %>% 
  mutate(ratio = round2(hpe/lpe, 2),
    outlier = case_when(
    ratio < q25.5 ~ "Outlier",
    ratio > q75.5 ~ "Outlier",
    TRUE ~ NA
  )) %>% 
  filter(outlier == "Outlier")
```


### Figure of Ratios by Clarity and Method
```{r}
f1 <- cal %>% 
  ggplot(aes(x = size, y = hpe/lpe, fill = size, colour = water_clarity)) +
  geom_hline(yintercept = 1.8, col = "red", alpha = 0.5) +
  geom_boxplot() +
  scale_fill_manual(values = size.cols) +
  scale_colour_manual(values = c("gray","black")) +
  facet_wrap(~lpe_method) +
  labs(y = "High Accuracy / Peak Count") +
  geom_text(data = cal.keep,
            aes(x = size, y = 16,
                 label = n), 
            position = position_dodge(width = 0.75)) 
f1
ggsave("./figures/Clarity_Size_Method_Ratios.png",
       plot = f1, height = 4.5, width = 10, dpi = 150)

ggsave("./figures/Clarity_Size_Method_Ratios.png",
       plot = f1, height = 7, width = 8, dpi = 150)
```

## Substrate
### Filterting
Only considering Method-Size-Substrate combinations of > 20 n
```{r}
cal$substrate_shade <- factor(cal$substrate_shade, levels=c("Light", "Medium", "Dark"), ordered=T)

cal.keep.sub <- count(cal,
                  lpe_method, size, substrate_shade)
```

```{r}
f2 <- cal %>% 
  ggplot(aes(x = size, y = hpe/lpe, fill = size, colour = substrate_shade)) +
  geom_hline(yintercept = 1.8, col = "red", alpha = 0.5) +
  geom_boxplot() +
  scale_fill_manual(values = size.cols) +
  scale_colour_manual(values = c("grey", "gray25", "black")) +
  facet_wrap(~lpe_method) +
  labs(y = "High Accuracy / Peak Count") +
  geom_text(data = cal.keep.sub,
            aes(x = size, y = 16,
                 label = n), 
            position = position_dodge(width = 0.75)) 

#ggsave("./figures/Substrate_Size_Method_Ratios.png",
#       plot = f2, height = 8, width = 10, dpi = 150)
```

Deciding not to run any substrate specific MoR models.

## Leave One Out
This is leave one out model fitting for each of MoR, RoM, and LM MoR
```{r}
# dataframe to fill with results.
model.test <- data.frame(model=character(),
                 usid=character(), 
                 fit=numeric(), 
                 se.fit=numeric()) 
```

```{r}
# saving each model type for later steps
saveRDS(cal.ratios.keep, file = "./r_objects/models/MoR_Clarity_Model.rds")
#saveRDS(cal.ratios.keep, file = "./r_objects/models/RoM_Clarity_Model.rds")
#saveRDS(rom.lm, file = "./r_objects/models/MoR_LM_Clarity_Model.rds")

# all models have no transformations, so just making a generic MoR_RoM shared specifications file
specifications_out <- data.frame(model = c("MoR_Clarity_Models"),
                                 target = c(NA),
                                 order = c(NA),
                                 type = c(NA),
                                 transform = c(NA))

spec.out <- spec.out %>% 
  filter(model != "MoR_Clarity_Models") %>% 
  rbind(., specifications_out)

saveRDS(spec.out, file=modout)
```
Need to add a running of the model without the one data point.

system - method - model
mor_clarity
```{r}
# creating a data frame that only contains data points part of MoR useable = Yes clarity_size sets
ratio.test.set <- left_join(cal, 
                            dplyr::select(cal.ratios.keep,
                                          lpe_method, size, water_clarity, useable.MoR)) %>% 
  filter(useable.MoR == "Yes")

for(i in 1:nrow(cal.ratios.keep)){ # for each method-size-clarity set with MoR = Yes
  loop.method <- cal.ratios.keep$lpe_method[i]
  loop.size <- cal.ratios.keep$size[i]
  loop.clarity <- cal.ratios.keep$water_clarity[i]
  
  loop.dat = filter(ratio.test.set,
                      lpe_method == loop.method,
                      size == loop.size,
                      water_clarity == loop.clarity)
  
            # only run the ratios for samples greater than 5
    if(nrow(loop.dat)>5){
      for(l in 1:nrow(loop.dat)){
        temp.dat <- loop.dat[-l, ]
        temp.usid <- loop.dat$usid[l]
      
        # create the MoR from the temporary loop data set
          loop.fit <- temp.dat %>% 
            group_by(lpe_method, size, water_clarity) %>%
            summarise(MoR = mean(hpe/lpe),
                    MoR.SD = sd(hpe/lpe))
          
          # estimate fit and error and bind it to the output (model.test) for MoR
           model.test <- rbind(model.test,
                        data.frame(model = paste(loop.size, loop.method, 
                                                 loop.clarity, "MoR_Clarity_Model", sep = "_"),
                                   usid = loop.dat$usid[l],
                                   fit = loop.fit$MoR*loop.dat$lpe[l],
                                   # var(fit) = lpe^2 * var(MoR) = lpe^2 * SD^2 = lpe * SD
                                   se.fit = loop.fit$MoR.SD*loop.dat$lpe[l],
                                   calibrated.estimate = loop.fit$MoR*loop.dat$lpe[l]))
      } # l of each datapoint
    } # if > 5
} # end of i loop for cal.ratio set
```

```{r}
model.rank <- left_join(model.test, 
                        dplyr::select(cal,
                                      usid, hpe, watershed_group:year, lpe_method)) %>% 
  mutate(error = calibrated.estimate - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error),
         rank.set.level = 4)

```
### Plots
```{r box and whisker of percent error by system}
model.rank %>% 
  ggplot(aes(x = reorder(gazetted_stream_name, abs.per.error, FUN = median), 
             # reorder FUN = median, reorders based on the median
             y = abs.per.error,
             fill = model)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_point(position = "jitter", alpha = 0.5,
             aes(colour = model)) +
  geom_hline(yintercept = 0.5, colour = "grey") +
  coord_flip() +
  labs(title = "All Data Model Performance",
       y = "Absolute percent error", x = "Streams")
```

#### Save Model Ranks
```{r}
count(model.rank, model)

MoR_Clarity_ranks <- model.rank %>% 
  filter(grepl("MoR_Clarity", model))

saveRDS(MoR_Clarity_ranks, "r_objects/01_Model_Ranks_from_MoR_Clarity_Model.rds")
```

# 2 Early Stuart
```{r}
cal.ratios <- cal %>% 
  filter(run_timing_group == "Early Stuart") %>% 
  # creating ratios by strea
  group_by(lpe_method, gazetted_stream_name) %>%
  summarise(n = n(), # sample size
            Average.Pop = mean(hpe),
            MoR = mean(hpe/lpe), # MoR = Mean of Ratios
            MoR.Min = min(hpe/lpe),
            MoR.Max = max(hpe/lpe),
            MoR.SD = sqrt(var(hpe/lpe) + mean((hpe_error^2)*(1/lpe)^2)), # Table 3 in McPherson et al. 2000 https://www.arlis.org/docs/vol1/ADFG/FMS/FMS00-02.pdf
            MoR.CV = MoR.SD/MoR,
            q75 = quantile(hpe/lpe, 0.75),
            q25 = quantile(hpe/lpe, 0.25)) %>% 
  # needed to calculated interquartile range
  mutate(iqr = q75-q25,
         iqr.5 = iqr*1.5, # for calculating outliers
         q75.5 = q75+iqr.5, # whisker max
         q25.5 = q25-iqr.5  # whisker min
         ) %>% 
  filter(n > 10)

# total ESTU average with just 3 streams
cal.ratios.estu <- cal %>% 
  # just based off of the three main streams
  filter(gazetted_stream_name %in% cal.ratios$gazetted_stream_name) %>% 
  group_by(lpe_method) %>% 
  summarise(n = n(), # sample size
            Average.Pop = mean(hpe),
            MoR = mean(hpe/lpe), # MoR = Mean of Ratios
            MoR.Min = min(hpe/lpe),
            MoR.Max = max(hpe/lpe),
            MoR.SD = sqrt(var(hpe/lpe) + mean((hpe_error^2)*(1/lpe)^2)), # Table 3 in McPherson et al. 2000 https://www.arlis.org/docs/vol1/ADFG/FMS/FMS00-02.pdf
            MoR.CV = MoR.SD/MoR,
            q75 = quantile(hpe/lpe, 0.75),
            q25 = quantile(hpe/lpe, 0.25)) %>% 
  # needed to calculated interquartile range
  mutate(iqr = q75-q25,
         iqr.5 = iqr*1.5, # for calculating outliers
         q75.5 = q75+iqr.5, # whisker max
         q25.5 = q25-iqr.5,  # whisker min
         gazetted_stream_name = "Forfar, Gluske, O'Ne-Ell")

# total ESTU average with all clear data
cal.ratios.estu.all <- cal %>% 
  # just based off of the three main streams
  filter(run_timing_group == "Early Stuart", water_clarity == "Clear") %>% 
  group_by(lpe_method) %>% 
  summarise(n = n(), # sample size
            Average.Pop = mean(hpe),
            MoR = mean(hpe/lpe), # MoR = Mean of Ratios
            MoR.Min = min(hpe/lpe),
            MoR.Max = max(hpe/lpe),
            MoR.SD = sqrt(var(hpe/lpe) + mean((hpe_error^2)*(1/lpe)^2)), # Table 3 in McPherson et al. 2000 https://www.arlis.org/docs/vol1/ADFG/FMS/FMS00-02.pdf
            MoR.CV = MoR.SD/MoR,
            q75 = quantile(hpe/lpe, 0.75),
            q25 = quantile(hpe/lpe, 0.25)) %>% 
  # needed to calculated interquartile range
  mutate(iqr = q75-q25,
         iqr.5 = iqr*1.5, # for calculating outliers
         q75.5 = q75+iqr.5, # whisker max
         q25.5 = q25-iqr.5,  # whisker min
         gazetted_stream_name = "All Early Stuart")
```

## Table Outputs
```{r}
table02 <- cal.ratios %>% 
  rbind(., cal.ratios.estu, cal.ratios.estu.all) %>% # add on the Estu Average
  dplyr::select(lpe_method:MoR.CV) %>% 
  mutate(Average.Pop = round2(Average.Pop, 0),
         MoR = round2(MoR, 2),
         MoR.Min = round2(MoR.Min, 2),
         MoR.Max = round2(MoR.Max, 2),
         MoR.SD = round2(MoR.SD, 2),
         MoR.CV = round2(MoR.CV, 3)*100)

write.csv(table02, "./tables/01_01-Mean-of-Ratios-summary-Early-Stuart.csv")
```

## Leave One Out
This is leave one out model fitting for each of MoR, RoM, and LM MoR
```{r}
# dataframe to fill with results.
model.test <- data.frame(model=character(),
                 usid=character(), 
                 fit=numeric(), 
                 se.fit=numeric()) 
```


```{r}
# saving each model type for later steps
saveRDS(cal.ratios, file = "./r_objects/models/MoR_ESTU_Single_Model.rds")
#saveRDS(cal.ratios.keep, file = "./r_objects/models/RoM_Clarity_Model.rds")
#saveRDS(rom.lm, file = "./r_objects/models/MoR_LM_Clarity_Model.rds")

# all models have no transformations, so just making a generic MoR_RoM shared specifications file
specifications_out <- data.frame(model = c("MoR_ESTU_Single_Model"),
                                 target = c(NA),
                                 order = c(NA),
                                 type = c(NA),
                                 transform = c(NA))

spec.out <- spec.out %>% 
  filter(model != "MoR_ESTU_Single_Model") %>% 
  rbind(., specifications_out)

saveRDS(spec.out, file=modout)
```

```{r}
for(i in 1:nrow(cal.ratios)){ # for each method-size-clarity set with MoR = Yes
  loop.system <- cal.ratios$gazetted_stream_name[i]
  
  loop.dat = filter(ratio.test.set,
                      gazetted_stream_name == loop.system)
  
            # only run the ratios for samples greater than 5
    if(nrow(loop.dat)>5){
      for(l in 1:nrow(loop.dat)){
        temp.dat <- loop.dat[-l, ]
        temp.usid <- loop.dat$usid[l]
      
        # create the MoR from the temporary loop data set
          loop.fit <- temp.dat %>% 
            group_by(lpe_method, gazetted_stream_name) %>%
            summarise(MoR = mean(hpe/lpe),
                    MoR.SD = sqrt(var(hpe/lpe) + mean((hpe_error^2)*(1/lpe)^2)))
          
          # estimate fit and error and bind it to the output (model.test) for MoR
           model.test <- rbind(model.test,
                        data.frame(model = paste(loop.system, "MoR_ESTU_Single_Model", sep = "_"),
                                   usid = loop.dat$usid[l],
                                   fit = loop.fit$MoR*loop.dat$lpe[l],
                                   # var(fit) = lpe^2 * var(MoR) = lpe^2 * SD^2 = lpe * SD
                                   se.fit = loop.fit$MoR.SD*loop.dat$lpe[l],
                                   calibrated.estimate = loop.fit$MoR*loop.dat$lpe[l]))
      } # l of each datapoint
    } # if > 5
} # end of i loop for cal.ratio set
```

```{r}
model.rank <- left_join(model.test, 
                        dplyr::select(cal,
                                      usid, hpe, watershed_group:year, lpe_method)) %>% 
  mutate(error = calibrated.estimate - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error),
         rank.set.level = 5)

```

```{r box and whisker of percent error by system}
model.rank %>% 
  ggplot(aes(x = reorder(gazetted_stream_name, abs.per.error, FUN = median), 
             # reorder FUN = median, reorders based on the median
             y = abs.per.error,
             fill = model)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_point(position = "jitter", alpha = 0.5,
             aes(colour = model)) +
  geom_hline(yintercept = 0.5, colour = "grey") +
  coord_flip() +
  labs(title = "All Data Model Performance",
       y = "Absolute percent error", x = "Streams")
```

#### Save Model Ranks
```{r}
count(model.rank, model)

saveRDS(model.rank, "r_objects/01_Model_Ranks_from_MoR_ESTU_Single_Model.rds")
```

# 3 Extra - Low Abundance Ratios
```{r}
cal.ratios.all <- cal %>% 
  # creating ratios by size and water clarity
  group_by(lpe_method, size = "All", water_clarity) %>%
  summarise(n = n(), # sample size
            min.pop = min(hpe),
            Average.Pop = mean(hpe),
            max.pop = max(hpe),
            MoR = mean(hpe/lpe), # MoR = Mean of Ratios
            MoR.Min = min(hpe/lpe),
            MoR.Max = max(hpe/lpe),
            MoR.SD = sqrt(var(hpe/lpe) + mean((hpe_error^2)*(1/lpe)^2)), # Table 3 in McPherson et al. 2000 https://www.arlis.org/docs/vol1/ADFG/FMS/FMS00-02.pdf
            MoR.CV = MoR.SD/MoR)

cal.ratios.25 <- cal %>% 
  filter(lpe < 25000) %>% 
  # creating ratios by size and water clarity
  group_by(lpe_method, size = "< 25k", water_clarity) %>%
  summarise(n = n(), # sample size
            min.pop = min(hpe),
            Average.Pop = mean(hpe),
            max.pop = max(hpe),
            MoR = mean(hpe/lpe), # MoR = Mean of Ratios
            MoR.Min = min(hpe/lpe),
            MoR.Max = max(hpe/lpe),
            MoR.SD = sqrt(var(hpe/lpe) + mean((hpe_error^2)*(1/lpe)^2)), # Table 3 in McPherson et al. 2000 https://www.arlis.org/docs/vol1/ADFG/FMS/FMS00-02.pdf
            MoR.CV = MoR.SD/MoR)

cal.ratios.25.sm <- cal %>% 
  filter(lpe < 25000,
         size %in% c("Small", "Medium")) %>% 
  # creating ratios by size and water clarity
  group_by(lpe_method, size = "< 25k SM", water_clarity) %>%
  summarise(n = n(), # sample size
            min.pop = min(hpe),
            Average.Pop = mean(hpe),
            max.pop = max(hpe),
            MoR = mean(hpe/lpe), # MoR = Mean of Ratios
            MoR.Min = min(hpe/lpe),
            MoR.Max = max(hpe/lpe),
            MoR.SD = sqrt(var(hpe/lpe) + mean((hpe_error^2)*(1/lpe)^2)), # Table 3 in McPherson et al. 2000 https://www.arlis.org/docs/vol1/ADFG/FMS/FMS00-02.pdf
            MoR.CV = MoR.SD/MoR)

cal.ratios.10.sm <- cal %>% 
  filter(lpe < 10000,
         size %in% c("Small", "Medium")) %>% 
  # creating ratios by size and water clarity
  group_by(lpe_method, size = "< 10k SM", water_clarity) %>%
  summarise(n = n(), # sample size
            min.pop = min(hpe),
            Average.Pop = mean(hpe),
            max.pop = max(hpe),
            MoR = mean(hpe/lpe), # MoR = Mean of Ratios
            MoR.Min = min(hpe/lpe),
            MoR.Max = max(hpe/lpe),
            MoR.SD = sqrt(var(hpe/lpe) + mean((hpe_error^2)*(1/lpe)^2)), # Table 3 in McPherson et al. 2000 https://www.arlis.org/docs/vol1/ADFG/FMS/FMS00-02.pdf
            MoR.CV = MoR.SD/MoR)

cal.ratios3 <- rbind(cal.ratios.all,
                     cal.ratios.25,
                     cal.ratios.25.sm,
                     cal.ratios.10.sm)
```

## Table Outputs
```{r}
table.extra <- cal.ratios3 %>% 
  mutate(min.pop = round2(min.pop, 0),
         Average.Pop = round2(Average.Pop, 0),
         max.pop = round2(max.pop, 0),
         MoR = round2(MoR, 2),
         MoR.Min = round2(MoR.Min, 2),
         MoR.Max = round2(MoR.Max, 2),
         MoR.SD = round2(MoR.SD, 2),
         MoR.CV = round2(MoR.CV, 3)*100)

write.csv(table.extra, "./tables/01_01-Mean-of-Ratios-Small-Abundance-Extra-summary.csv")
```

## Save Low Abundance MoR Models
```{r}
# saving each model type for later steps
saveRDS(cal.ratios.25.sm, file = "./r_objects/models/MoR_Low_Abundance_Models.rds")

# all models have no transformations, so just making a generic MoR_RoM shared specifications file
specifications_out <- data.frame(model = c("MoR_Low_Abundance_Models"),
                                 target = c(NA),
                                 order = c(NA),
                                 type = c(NA),
                                 transform = c(NA))

spec.out <- spec.out %>% 
  filter(model != "MoR_Low_Abundance_Models") %>% 
  rbind(., specifications_out)

saveRDS(spec.out, file=modout)
```
## Leave One Out
This is leave one out model fitting for each of MoR, RoM, and LM MoR
```{r}
# dataframe to fill with results.
model.test <- data.frame(model=character(),
                 usid=character(), 
                 fit=numeric(), 
                 se.fit=numeric()) 
```

```{r}
# creating a data frame that only contains data points part of MoR useable = Yes clarity_size sets
loop.dat <- cal %>% 
  filter(lpe < 25000,
         size %in% c("Small", "Medium"))

    for(l in 1:nrow(loop.dat)){
      temp.clarity <- loop.dat$water_clarity[l]
      temp.method <- loop.dat$lpe_method[l]
        temp.dat <- loop.dat[-l, ] %>% 
          filter(water_clarity == temp.clarity,
                 lpe_method == temp.method)
        temp.usid <- loop.dat$usid[l]
      
        # create the MoR from the temporary loop data set
          loop.fit <- temp.dat %>% 
            summarise(MoR = mean(hpe/lpe),
                    MoR.SD = sd(hpe/lpe))
          
          # estimate fit and error and bind it to the output (model.test) for MoR
           model.test <- rbind(model.test,
                        data.frame(model = "MoR_Low_Abundance_Models",
                                   usid = loop.dat$usid[l],
                                   fit = loop.fit$MoR*loop.dat$lpe[l],
                                   # var(fit) = lpe^2 * var(MoR) = lpe^2 * SD^2 = lpe * SD
                                   se.fit = loop.fit$MoR.SD*loop.dat$lpe[l],
                                   calibrated.estimate = loop.fit$MoR*loop.dat$lpe[l]))
      } # l of each datapoint
```

```{r}
  model.rank <- left_join(model.test, 
                        dplyr::select(cal,
                                      usid, hpe, watershed_group:year, lpe_method)) %>% 
  mutate(error = calibrated.estimate - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error),
         rank.set.level = 4.1)

```

```{r}
saveRDS(model.rank, "r_objects/01_Model_Ranks_from_MoR_Low_Abundance_Models.rds")
```

# 4 Harrison
```{r}
cal.ratios.25 <- cal %>% 
  filter(gazetted_stream_name == "Harrison River") %>% 
  # creating ratios by size and water clarity
  group_by(lpe_method, water_clarity) %>%
  summarise(n = n(), # sample size
            min.pop = min(hpe),
            Average.Pop = mean(hpe),
            max.pop = max(hpe),
            MoR = mean(hpe/lpe), # MoR = Mean of Ratios
            MoR.Min = min(hpe/lpe),
            MoR.Max = max(hpe/lpe),
            MoR.SD = sqrt(var(hpe/lpe) + mean((hpe_error^2)*(1/lpe)^2)), # Table 3 in McPherson et al. 2000 https://www.arlis.org/docs/vol1/ADFG/FMS/FMS00-02.pdf
            MoR.CV = MoR.SD/MoR)

cal.harrison <- cal %>% 
  filter(gazetted_stream_name == "Harrison River") %>% 
  # creating ratios by size and water clarity
  group_by(lpe_method) %>%
  summarise(n = n(), # sample size
            min.pop = min(hpe),
            Average.Pop = mean(hpe),
            max.pop = max(hpe),
            MoR = mean(hpe/lpe), # MoR = Mean of Ratios
            MoR.Min = min(hpe/lpe),
            MoR.Max = max(hpe/lpe),
            MoR.SD = sqrt(var(hpe/lpe) + mean((hpe_error^2)*(1/lpe)^2)), # Table 3 in McPherson et al. 2000 https://www.arlis.org/docs/vol1/ADFG/FMS/FMS00-02.pdf
            MoR.CV = MoR.SD/MoR)
```

## Leave One Out
This is leave one out model fitting for each of MoR, RoM, and LM MoR
```{r}
# dataframe to fill with results.
model.test <- data.frame(model=character(),
                 usid=character(), 
                 fit=numeric(), 
                 se.fit=numeric()) 
```

```{r}
# saving each model type for later steps
saveRDS(cal.harrison, file = "./r_objects/models/MoR_Harrison_Model.rds")

# all models have no transformations, so just making a generic MoR_RoM shared specifications file
specifications_out <- data.frame(model = c("MoR_Harrison_Models"),
                                 target = c(NA),
                                 order = c(NA),
                                 type = c(NA),
                                 transform = c(NA))

spec.out <- spec.out %>% 
  filter(model != "MoR_Harrison_Models") %>% 
  rbind(., specifications_out)

saveRDS(spec.out, file=modout)
```

```{r}
# creating a data frame that only contains data points part of MoR useable = Yes clarity_size sets
loop.dat <- cal %>% 
  filter(gazetted_stream_name == "Harrison River")

    for(l in 1:nrow(loop.dat)){
        temp.dat <- loop.dat[-l, ]
        temp.usid <- loop.dat$usid[l]
      
        # create the MoR from the temporary loop data set
          loop.fit <- temp.dat %>% 
            group_by(lpe_method) %>%
            summarise(MoR = mean(hpe/lpe),
                    MoR.SD = sd(hpe/lpe))
          
          # estimate fit and error and bind it to the output (model.test) for MoR
           model.test <- rbind(model.test,
                        data.frame(model = "MoR_Harrison_Models",
                                   usid = loop.dat$usid[l],
                                   fit = loop.fit$MoR*loop.dat$lpe[l],
                                   # var(fit) = lpe^2 * var(MoR) = lpe^2 * SD^2 = lpe * SD
                                   se.fit = loop.fit$MoR.SD*loop.dat$lpe[l],
                                   calibrated.estimate = loop.fit$MoR*loop.dat$lpe[l]))
      } # l of each datapoint
```

```{r}
  model.rank <- left_join(model.test, 
                        dplyr::select(cal,
                                      usid, hpe, watershed_group:year, lpe_method)) %>% 
  mutate(error = calibrated.estimate - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error),
         rank.set.level = 5)

```

```{r}
saveRDS(model.rank, "r_objects/01_Model_Ranks_from_MoR_Harrison_Model.rds")
```

