---
title: "01-03-Inverse-Regressions"
author: "Michael Arbeider"
date: "2026-03-06"
---

# Set-up
## Packages
```{r}
rm(list = ls())
library(tidyverse) # most %>% 
library(broom) # augment()
theme_set(theme_bw()) # ggplot
source("00-Functions.R") # scale_x_reordered, rank.summary(), round2()
library(patchwork) # for multi plot ggplot graphs
```

## Data
```{r}
cal <-readRDS("r_objects/sockeye_calibration_data.rds")   

glimpse(cal)

cal$lnhpe <- log(cal$hpe)
```

## Model Specifications FILE
```{r}
modout <- "./r_objects/models/model_specifications.rds"
spec.out <- readRDS(modout)
```

# Inverse Function
This Code is adapted from Kleiven et al. 2023 - Using camera traps to monitor 
cyclic vole populations
It is to apply the calibrate function from the investr package and return 
statistics on a data point using Inverse Prediction
```{r}
cal_4apply <- function(y0, # Low Precision Estimate
                       mod, # simple linear model 
                       mean.res=FALSE, interval="Wald")
  # Apply calibration function from INVESTR PACKAGE
{
  x0 <- investr::calibrate(mod,y0,interval=interval, mean.response = mean.res)
  return(x0[1:4]) # return the High Precision estimate, confidence interval, and Standard Error
}
```

# Models
## No Large Model
### Aerial + Ground
subset data to exclude Large and xL data
```{r}
cal.SM <- filter(cal,
                  !size %in% c("X-Large", "Large"))
```

Simple Linear Model
```{r}
lm.SM <- lm(lnlpe~lnhpe, data = cal.SM)
summary(lm.SM)

# what the linear model looks like
ggplot(cal.SM,
       aes(x = lnhpe, y = lnlpe)) +
  geom_point() +
  geom_smooth(method = "lm")
```

#### Leave One Out
```{r}
saveRDS(lm.SM, file = "./r_objects/models/Inverse_SM_Both_Data_Model.rds")

specifications_out <- data.frame(model = c("Inverse_SM_Both_Data_Model"),
                                 target = c("lpe","lpe","lpe"),
                                 order = c(1),
                                 type = c("log"),
                                 transform = c(NA))

spec.out <- spec.out %>% 
  filter(model != "Inverse_SM_Both_Data_Model") %>% 
  rbind(., specifications_out)

saveRDS(spec.out, file=modout)
```

```{r}
model.test <- data.frame(model=character(),
                 usid=character(), 
                 fit=numeric(), 
                 se.fit=numeric(),
                 calibrated.estimate=numeric()) 
```

```{r}
model = lm.SM
  for(i in 1:nrow(cal.SM)){
    m.temp <- lm(lnlpe~lnhpe, data = cal.SM[-i, ])
    
    p.temp <- matrix(unlist(t(sapply(cal.SM$lnlpe[i], cal_4apply, mod=m.temp))), ncol = 4)
    p.temp <- as.data.frame( p.temp )
    colnames(p.temp) <- c("fit", "lower", "upper", "se.fit")
    
    model.test <- rbind(model.test, data.frame(model = paste0(as.character(model$call)[2], " Inverse B"),
                                 usid = cal.SM$usid[i],
                                 fit = p.temp$fit,
                                 se.fit = p.temp$se.fit,
                                 calibrated.estimate = exp(p.temp$fit)))
  }
```

```{r}
model.rank <- left_join(model.test, 
                        dplyr::select(cal,
                                      usid, hpe, watershed_group:year, lpe_method)) %>% 
  mutate(error = calibrated.estimate - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error),
         rank.set.level = 2)


```

```{r plot of percent error}
ggplot(model.rank,
       aes(x = log(hpe), y = abs.per.error)) +
  geom_point()
```

#### Save Model Ranks
```{r}
saveRDS(model.rank, "r_objects/01_Model_Ranks_from_Inverse_SM_Both_Data_Model.rds")
```

### Ground
subset data to exclude Aerial
```{r}
cal.SMG <- filter(cal.SM,
                  lpe_method == "Ground")
```

Simple Linear Model
```{r}
lm.SMG <- lm(lnlpe~lnhpe, data = cal.SMG)
summary(lm.SMG)

# what the linear model looks like
ggplot(cal.SMG,
       aes(x = lnhpe, y = lnlpe)) +
  geom_point() +
  geom_smooth(method = "lm")
```

#### Leave One Out
```{r}
saveRDS(lm.SMG, file = "./r_objects/models/Inverse_SM_Ground_Data_Model.rds")

specifications_out <- data.frame(model = c("Inverse_SM_Ground_Data_Model"),
                                 target = c("lpe","lpe","lpe"),
                                 order = c(1),
                                 type = c("log"),
                                 transform = c(NA))

spec.out <- spec.out %>% 
  filter(model != "Inverse_SM_Ground_Data_Model") %>% 
  rbind(., specifications_out)

saveRDS(spec.out, file=modout)
```

```{r}
model.test <- data.frame(model=character(),
                 usid=character(), 
                 fit=numeric(), 
                 se.fit=numeric(),
                 calibrated.estimate=numeric()) 
```

```{r}
model = lm.SMG
  for(i in 1:nrow(cal.SMG)){
    m.temp <- lm(lnlpe~lnhpe, data = cal.SMG[-i, ])
    
    p.temp <- matrix(unlist(t(sapply(cal.SMG$lnlpe[i], cal_4apply, mod=m.temp))), ncol = 4)
    p.temp <- as.data.frame( p.temp )
    colnames(p.temp) <- c("fit", "lower", "upper", "se.fit")
    
    model.test <- rbind(model.test, data.frame(model = paste0(as.character(model$call)[2], " Inverse G"),
                                 usid = cal.SMG$usid[i],
                                 fit = p.temp$fit,
                                 se.fit = p.temp$se.fit,
                                 calibrated.estimate = exp(p.temp$fit)))
  }
```

```{r}
model.rank <- left_join(model.test, 
                        dplyr::select(cal,
                                      usid, hpe, watershed_group:year, lpe_method)) %>% 
  mutate(error = calibrated.estimate - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error), 
         rank.set.level = 2)


```

```{r plot of percent error}
ggplot(model.rank,
       aes(x = log(hpe), y = abs.per.error)) +
  geom_point()
```

#### Save Model Ranks
```{r}
saveRDS(model.rank, "r_objects/01_Model_Ranks_from_Inverse_SM_Ground_Data_Model.rds")
```

### Aerial
subset data to exclude Ground
```{r}
cal.SMA <- filter(cal.SM,
                  lpe_method == "Aerial")
```

Simple Linear Model
```{r}
lm.SMA <- lm(lnlpe~lnhpe, data = cal.SMA)
summary(lm.SMA)

# what the linear model looks like
ggplot(cal.SMA,
       aes(x = lnhpe, y = lnlpe)) +
  geom_point() +
  geom_smooth(method = "lm")
```

#### Leave One Out
```{r}
saveRDS(lm.SMA, file = "./r_objects/models/Inverse_SM_Aerial_Data_Model.rds")

specifications_out <- data.frame(model = c("Inverse_SM_Aerial_Data_Model"),
                                 target = c("lpe","lpe","lpe"),
                                 order = c(1),
                                 type = c("log"),
                                 transform = c(NA))

spec.out <- spec.out %>% 
  filter(model != "Inverse_SM_Aerial_Data_Model") %>% 
  rbind(., specifications_out)

saveRDS(spec.out, file=modout)
```

```{r}
model.test <- data.frame(model=character(),
                 usid=character(), 
                 fit=numeric(), 
                 se.fit=numeric(),
                 calibrated.estimate=numeric()) 
```

```{r}
model = lm.SMA
  for(i in 1:nrow(cal.SMA)){
    m.temp <- lm(lnlpe~lnhpe, data = cal.SMA[-i, ])
    
    p.temp <- matrix(unlist(t(sapply(cal.SMA$lnlpe[i], cal_4apply, mod=m.temp))), ncol = 4)
    p.temp <- as.data.frame( p.temp )
    colnames(p.temp) <- c("fit", "lower", "upper", "se.fit")
    
    model.test <- rbind(model.test, data.frame(model = paste0(as.character(model$call)[2], " Inverse A"),
                                 usid = cal.SMA$usid[i],
                                 fit = p.temp$fit,
                                 se.fit = p.temp$se.fit,
                                 calibrated.estimate = exp(p.temp$fit)))
  }
```

```{r}
model.rank <- left_join(model.test, 
                        dplyr::select(cal,
                                      usid, hpe, watershed_group:year, lpe_method)) %>% 
  mutate(error = calibrated.estimate - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error), 
         rank.set.level = 2)


```

```{r plot of percent error}
ggplot(model.rank,
       aes(x = log(hpe), y = abs.per.error)) +
  geom_point()
```

#### Save Model Ranks
```{r}
saveRDS(model.rank, "r_objects/01_Model_Ranks_from_Inverse_SM_Aerial_Data_Model.rds")
```

## Large Model
### Aerial + Ground
subset data to exclude Large and xL data
```{r}
cal.LXL <- filter(cal,
                  size %in% c("X-Large", "Large"))
```

Simple Linear Model
```{r}
lm.LXL <- lm(lnlpe~lnhpe, data = cal.LXL)
summary(lm.LXL)

# what the linear model looks like
ggplot(cal.LXL,
       aes(x = lnhpe, y = lnlpe)) +
  geom_point() +
  geom_smooth(method = "lm")
```

#### Leave One Out
```{r}
saveRDS(lm.LXL, file = "./r_objects/models/Inverse_LXL_Both_Data_Model.rds")

specifications_out <- data.frame(model = c("Inverse_LXL_Both_Data_Model"),
                                 target = c("lpe","lpe","lpe"),
                                 order = c(1),
                                 type = c("log"),
                                 transform = c(NA))

spec.out <- spec.out %>% 
  filter(model != "Inverse_LXL_Both_Data_Model") %>% 
  rbind(., specifications_out)

saveRDS(spec.out, file=modout)
```

```{r}
model.test <- data.frame(model=character(),
                 usid=character(), 
                 fit=numeric(), 
                 se.fit=numeric(),
                 calibrated.estimate=numeric()) 
```

```{r}
model = lm.LXL
  for(i in 1:nrow(cal.LXL)){
    m.temp <- lm(lnlpe~lnhpe, data = cal.LXL[-i, ])
    
    p.temp <- matrix(unlist(t(sapply(cal.LXL$lnlpe[i], cal_4apply, mod=m.temp))), ncol = 4)
    p.temp <- as.data.frame( p.temp )
    colnames(p.temp) <- c("fit", "lower", "upper", "se.fit")
    
    model.test <- rbind(model.test, data.frame(model = paste0(as.character(model$call)[2], " Inverse B"),
                                 usid = cal.LXL$usid[i],
                                 fit = p.temp$fit,
                                 se.fit = p.temp$se.fit,
                                 calibrated.estimate = exp(p.temp$fit)))
  }
```

```{r}
model.rank <- left_join(model.test, 
                        dplyr::select(cal,
                                      usid, hpe, watershed_group:year, lpe_method)) %>% 
  mutate(error = calibrated.estimate - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error), 
         rank.set.level = 2)


```

```{r plot of percent error}
ggplot(model.rank,
       aes(x = log(hpe), y = abs.per.error)) +
  geom_point()
```

#### Save Model Ranks
```{r}
saveRDS(model.rank, "r_objects/01_Model_Ranks_from_Inverse_LXL_Both_Data_Model.rds")
```

### Ground
subset data to exclude Aerial
```{r}
cal.LXLG <- filter(cal.LXL,
                  lpe_method == "Ground")
```

Simple Linear Model
```{r}
lm.LXLG <- lm(lnlpe~lnhpe, data = cal.LXLG)
summary(lm.LXLG)

# what the linear model looks like
ggplot(cal.LXLG,
       aes(x = lnhpe, y = lnlpe)) +
  geom_point() +
  geom_smooth(method = "lm")
```

#### Leave One Out
```{r}
saveRDS(lm.LXLG, file = "./r_objects/models/Inverse_LXL_Ground_Data_Model.rds")

specifications_out <- data.frame(model = c("Inverse_LXL_Ground_Data_Model"),
                                 target = c("lpe","lpe","lpe"),
                                 order = c(1),
                                 type = c("log"),
                                 transform = c(NA))

spec.out <- spec.out %>% 
  filter(model != "Inverse_LXL_Ground_Data_Model") %>% 
  rbind(., specifications_out)

saveRDS(spec.out, file=modout)
```

```{r}
model.test <- data.frame(model=character(),
                 usid=character(), 
                 fit=numeric(), 
                 se.fit=numeric(),
                 calibrated.estimate=numeric()) 
```

```{r}
model = lm.LXLG
  for(i in 1:nrow(cal.LXLG)){
    m.temp <- lm(lnlpe~lnhpe, data = cal.LXLG[-i, ])
    
    p.temp <- matrix(unlist(t(sapply(cal.LXLG$lnlpe[i], cal_4apply, mod=m.temp))), ncol = 4)
    p.temp <- as.data.frame( p.temp )
    colnames(p.temp) <- c("fit", "lower", "upper", "se.fit")
    
    model.test <- rbind(model.test, data.frame(model = paste0(as.character(model$call)[2], " Inverse G"),
                                 usid = cal.LXLG$usid[i],
                                 fit = p.temp$fit,
                                 se.fit = p.temp$se.fit,
                                 calibrated.estimate = exp(p.temp$fit)))
  }
```

```{r}
model.rank <- left_join(model.test, 
                        dplyr::select(cal,
                                      usid, hpe, watershed_group:year, lpe_method)) %>% 
  mutate(error = calibrated.estimate - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error), 
         rank.set.level = 2)


```

```{r plot of percent error}
ggplot(model.rank,
       aes(x = log(hpe), y = abs.per.error)) +
  geom_point()
```

#### Save Model Ranks
```{r}
saveRDS(model.rank, "r_objects/01_Model_Ranks_from_Inverse_LXL_Ground_Data_Model.rds")
```

### Aerial
subset data to exclude Ground
```{r}
cal.LXLA <- filter(cal.LXL,
                  lpe_method == "Aerial")
```

Simple Linear Model
```{r}
lm.LXLA <- lm(lnlpe~lnhpe, data = cal.LXLA)
summary(lm.LXLA)

# what the linear model looks like
ggplot(cal.LXLA,
       aes(x = lnhpe, y = lnlpe)) +
  geom_point() +
  geom_smooth(method = "lm")
```

#### Leave One Out
```{r}
saveRDS(lm.LXLA, file = "./r_objects/models/Inverse_LXL_Aerial_Data_Model.rds")

specifications_out <- data.frame(model = c("Inverse_LXL_Aerial_Data_Model"),
                                 target = c("lpe","lpe","lpe"),
                                 order = c(1),
                                 type = c("log"),
                                 transform = c(NA))

spec.out <- spec.out %>% 
  filter(model != "Inverse_LXL_Aerial_Data_Model") %>% 
  rbind(., specifications_out)

saveRDS(spec.out, file=modout)
```

```{r}
model.test <- data.frame(model=character(),
                 usid=character(), 
                 fit=numeric(), 
                 se.fit=numeric(),
                 calibrated.estimate=numeric()) 
```

```{r}
model = lm.LXLA
  for(i in 1:nrow(cal.LXLA)){
    m.temp <- lm(lnlpe~lnhpe, data = cal.LXLA[-i, ])
    
    p.temp <- matrix(unlist(t(sapply(cal.LXLA$lnlpe[i], cal_4apply, mod=m.temp))), ncol = 4)
    p.temp <- as.data.frame( p.temp )
    colnames(p.temp) <- c("fit", "lower", "upper", "se.fit")
    
    model.test <- rbind(model.test, data.frame(model = paste0(as.character(model$call)[2], " Inverse A"),
                                 usid = cal.LXLA$usid[i],
                                 fit = p.temp$fit,
                                 se.fit = p.temp$se.fit,
                                 calibrated.estimate = exp(p.temp$fit)))
  }
```

```{r}
model.rank <- left_join(model.test, 
                        dplyr::select(cal,
                                      usid, hpe, watershed_group:year, lpe_method)) %>% 
  mutate(error = calibrated.estimate - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error), 
         rank.set.level = 2)


```

```{r plot of percent error}
ggplot(model.rank,
       aes(x = log(hpe), y = abs.per.error)) +
  geom_point()
```

#### Save Model Ranks
```{r}
saveRDS(model.rank, "r_objects/01_Model_Ranks_from_Inverse_LXL_Aerial_Data_Model.rds")
```

# Table01
```{r}
# making a blank table - partly did this way to think about what data I wanted to extract
table01 <- data.frame(
  Method = NA,
  Set = NA,
  SizeOrSystem = NA,
  SampleSize = NA,
  Intercept = NA,
  Int.Error = NA,
  Slope = NA,
  Slp.Error = NA,
  R2 = NA
)

# Generate the list of lm models
models <- mget(ls(pattern = "^lm\\.")) # grab models from above

for(i in 1:length(models)){ # for every model
  if(i>1) table01 <- rbind(table01, table01[1,]) # add a new blank row if not the first row
  # then pull out and add the various variables
  table01$Set[i] = names(models)[i]
  table01$SampleSize[i] = sum(summary(models[[i]])$df[c(1,2)])
  table01$Intercept[i] = summary(models[[i]])$coefficients[1,1]
  table01$Int.Error[i] = summary(models[[i]])$coefficients[1,2]
  table01$Slope[i] = summary(models[[i]])$coefficients[2,1]
  table01$Slp.Error[i] = summary(models[[i]])$coefficients[2,2]
  table01$R2[i] = summary(models[[i]])$r.squared*100
}

table01$Method = c("Both", "Aerial", "Ground",
                   "Both", "Aerial", "Ground")
table01$SizeOrSystem = c(rep("Large-X-Large", 3),
                         rep("Small-Medium", 3))
table01$Set = "Size-Aggregate (2)"
```
## Residual Plots
Residual plots for each lm model.
```{r}
pL <- ggplot(augment(lm.LXL),
             aes(x = .fitted, y = .resid)) +
    geom_hline(yintercept = 0, colour = "grey") +
  geom_point() +
  labs(x = "Fitted Values", y = "Response Residuals",
       title = "Large-X-Large-Both")

pLA <- ggplot(augment(lm.LXLA),
             aes(x = .fitted, y = .resid)) +
    geom_hline(yintercept = 0, colour = "grey") +
  geom_point() +
  labs(x = "Fitted Values", y = "Response Residuals",
       title = "Large-X-Large-Aerial")

pLG <- ggplot(augment(lm.LXLG),
             aes(x = .fitted, y = .resid)) +
    geom_hline(yintercept = 0, colour = "grey") +
  geom_point() +
  labs(x = "Fitted Values", y = "Response Residuals",
       title = "Large-X-Large-Ground")

pSM <- ggplot(augment(lm.SM),
             aes(x = .fitted, y = .resid)) +
    geom_hline(yintercept = 0, colour = "grey") +
  geom_point() +
  labs(x = "Fitted Values", y = "Response Residuals",
       title = "Small-Medium-Both")

pSMA <- ggplot(augment(lm.SMA),
             aes(x = .fitted, y = .resid)) +
    geom_hline(yintercept = 0, colour = "grey") +
  geom_point() +
  labs(x = "Fitted Values", y = "Response Residuals",
       title = "Small-Medium-Aerial")

pSMG <- ggplot(augment(lm.SMG),
             aes(x = .fitted, y = .resid)) +
    geom_hline(yintercept = 0, colour = "grey") +
  geom_point() +
  labs(x = "Fitted Values", y = "Response Residuals",
       title = "Small-Medium-Ground")


(pL + pSM + pLA + pSMA + pLG + pSMG) + 
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2, axis = "collect") +
  theme(plot.tag = element_text(size = 6))

ggsave("./figures/01_Model_Diagnostics/01-03-Inverse-Regression-01.png",
       height = 10, width = 8, dpi = 95)
```

# Individual
```{r}
(cal.fil.5 <- count(cal, gazetted_stream_name, lpe_method) %>% 
  filter(n>5) %>% 
  mutate(filter5 = paste0(gazetted_stream_name, "_", lpe_method)))

cal5 <- cal %>% 
  mutate(filter5 = paste0(gazetted_stream_name, "_", lpe_method)) %>% 
  filter(filter5 %in% cal.fil.5$filter5)

cal5 %>% 
  ggplot(aes(x = lnlpe, y = log(hpe),
             colour = filter5)) +
  geom_point()
```

## Analysis
```{r}
t1n <- nrow(table01) # number of rows of table01 from above
models <- list()

for(k in 1:nrow(cal.fil.5)){ # for every individual system
  print(k)
  # make temporary data with just one system
  cal5.temp <- filter(cal5, filter5 == cal.fil.5$filter5[k])
  # make temporary model with just one system
  
  m.temp <- lm(lnlpe~lnhpe, data = cal5.temp)
  
  table01 <- rbind(table01, table01[1,]) # add a row to fill in
  # then pull out and add the various variables
  # # need to add the base row number, t1n, to the k of this loop to keep buidling it up
  table01$Set[t1n+k] = "Individual (5)"
  table01$Method[t1n+k] = as.character(cal.fil.5$lpe_method[k])
  table01$SizeOrSystem[t1n+k] = cal.fil.5$gazetted_stream_name[k]
  table01$SampleSize[t1n+k] = sum(summary(m.temp)$df[c(1,2)])
  table01$Intercept[t1n+k] = summary(m.temp)$coefficients[1,1]
  table01$Int.Error[t1n+k] = summary(m.temp)$coefficients[1,2]
  table01$Slope[t1n+k] = summary(m.temp)$coefficients[2,1]
  table01$Slp.Error[t1n+k] = summary(m.temp)$coefficients[2,2]
  table01$R2[t1n+k] = summary(m.temp)$r.squared*100
  
  model.test <- data.frame(model=character(),
                 usid=character(), 
                 fit=numeric(), 
                 se.fit=numeric(),
                 calibrated.estimate=numeric())
  models[[k]] <- m.temp
    
  saveRDS(m.temp, file = paste0("./r_objects/models/", cal.fil.5$filter5[k], "-Individual_Inverse_Data_Model.rds"))

    for(i in 1:nrow(cal5.temp)){ # for every data point in cal5.temp
      
      m.temp <- lm(lnlpe~lnhpe, data = cal5.temp[-i, ])
      
      p.temp <- matrix(unlist(t(sapply(cal5.temp$lnlpe[i], cal_4apply, mod=m.temp))), ncol = 4)
     p.temp <- as.data.frame( p.temp )
     colnames(p.temp) <- c("fit", "lower", "upper", "se.fit")
    
     model.test <- rbind(model.test, data.frame(model = paste0(as.character(m.temp$call)[2], " Inverse Individual"),
                                 usid = cal5.temp$usid[i],
                                 fit = p.temp$fit,
                                 se.fit = p.temp$se.fit,
                                 calibrated.estimate = exp(p.temp$fit)))
  } # end of i loop
  model.rank.temp <- left_join(model.test, 
                        dplyr::select(cal5.temp,
                                      usid, hpe, watershed_group:year, lpe_method)) %>% 
    mutate(error = calibrated.estimate - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error),
         rank.set.level = 5)
  
  if(k == 1) model.rank = model.rank.temp
  if(k != 1) model.rank = rbind(model.rank, model.rank.temp)
} # end of k loop

```

# Save Model Ranks
```{r}
saveRDS(model.rank, "r_objects/01_Model_Ranks_from_Individual_Inverse_Data_Model.rds")
```

```{r}
specifications_out <- data.frame(model = c("Individual_Inverse_Data_Model"),
                                 target = c("lpe","lpe","lpe"),
                                 order = c(1),
                                 type = c("log"),
                                 transform = c(NA))

spec.out <- spec.out %>% 
  filter(model != "Individual_Inverse_Data_Model") %>% 
  rbind(., specifications_out)

saveRDS(spec.out, file=modout)
```

#Table01 Save
```{r}
table01[,5:9] <- round2(table01[,5:9], 2)

write.csv(table01, "./tables/01_03-Inverse-Regression-summary.csv",
          row.names = F)
```

## Residual Plots
```{r}
plist <- list() # list for plots

for(i in 1:length(models)){
  plist[[i]] <- ggplot(augment(models[[i]]),
             aes(x = .fitted, y = .resid)) +
    geom_hline(yintercept = 0, colour = "grey") +
  geom_point() +
  labs(x = "Fitted Values", y = "Response Residual",
       title = cal.fil.5$gazetted_stream_name[i])
}

plist[[1]] +  plist[[2]]  +  plist[[3]]+  
  plist[[4]]+  plist[[5]]+  plist[[6]]+  
  plist[[7]]+  plist[[8]] +
  plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2, axis = "collect") +
  theme(plot.tag = element_text(size = 6))

ggsave("./figures/01_Model_Diagnostics/01-03-Inverse-Regression-02.png",
       height = 11, width = 8, dpi = 95)
```

# Remove Some Models from the Candidate Set
Based on residual patterns and small sample sizes.
Removing models with a sample size less than 11 (i.e. > 10)
```{r}
# deleting models
file.remove("./r_objects/models/Harrison River_Aerial-Individual_Inverse_Data_Model.rds")
file.remove("./r_objects/models/Adams River (Lower)_Ground-Individual_Inverse_Data_Model.rds")
file.remove("./r_objects/models/Inverse_LXL_Ground_Data_Model.rds")

# deleting Model Rank files
file.remove("./r_objects/01_Model_Ranks_from_Inverse_LXL_Ground_Data_Model.rds")

model.rank <- model.rank %>% 
  filter(!gazetted_stream_name %in% c("Harrison River", "Adams River (Lower)"))
saveRDS(model.rank, "r_objects/01_Model_Ranks_from_Individual_Inverse_Data_Model.rds")
```

