---
title: "01-05-Early-Stu-Specific"
author: "Michael Arbeider"
date: "2026-03-06"
---

# Set-up
## Packages
```{r}
rm(list = ls())
library(tidyverse) # most %>% 
theme_set(theme_bw()) # ggplot
library(MASS) # boxcox - masked from ‘package:dplyr’: select()
library(lme4) # generalized linear regressions
# methods(class = "lmerMod") # shows what functionality is available
library(MuMIn) # dredge
library(PerformanceAnalytics) # correlation figure
library(car)       # for vif()
library(patchwork) # for multi plot ggplot graphs
source("00-Functions.R") # scale_x_reordered, rank.summary(), round2()
```

This analysis is based off of an analysis done by Scott Decker.
Original Summary: Early_Stuart_esc_estimation_issues.docx
Original Code: Early_Stuart_calibration.R

ground_cuml_carc_above = Cumulative Carcasses up till the max count date.
ground_carcs_total_above = cumulative carcasses for the whole program.

## Data
```{r}
modout <- "./r_objects/models/model_specifications.rds" # for saving specifications later
spec.out <- readRDS(modout)
cal <-readRDS("r_objects/sockeye_calibration_data.rds")   

glimpse(cal)

cal <- cal %>% filter(run_timing_group == "Early Stuart", # Early Stuart
                      lpe_method == "Ground", # Ground Surveys
                      !is.na(ground_carcs_total_above)) %>%  # Need Total Carcs 
  mutate(protraction = ground_cuml_carc_above/ground_carcs_total_above, # "First Ratio" in the report
         deadvlive = ground_carcs_total_above/peak_live_above) # "Second Ratio" in the report
```
### Test for Normality
High precision estimates
```{r}
hist(cal$hpe)
shapiro.test(cal$hpe) # p < 0.01-0.05 means that it is NOT normal
b <- boxcox(lm(cal$hpe ~ 1))
b$x[which.max(b$y)] # lambda
hist(log(cal$hpe))
shapiro.test(log(cal$hpe)) # acceptable
hist(sqrt(cal$hpe))
shapiro.test(sqrt(cal$hpe)) # not as good
hist((cal$hpe)^(1/3))
shapiro.test((cal$hpe)^(1/3)) # best
```
low precision estimates
```{r}  
hist(cal$lpe)
shapiro.test(cal$lpe) # p < 0.01-0.05 means that it is NOT normal
b <- boxcox(lm(cal$lpe ~ 1))
b$x[which.max(b$y)] # lambda
hist(log(cal$lpe))
shapiro.test(log(cal$lpe)) # not very acceptable
hist(sqrt(cal$lpe))
shapiro.test(sqrt(cal$lpe)) # better
hist((cal$lpe)^(1/3))
shapiro.test((cal$lpe)^(1/3)) # best
```
total ground carcs
```{r}
hist(cal$ground_carcs_total_above)
shapiro.test(cal$ground_carcs_total_above) # p < 0.01-0.05 means that it is NOT normal
b <- boxcox(lm(cal$ground_carcs_total_above ~ 1))
b$x[which.max(b$y)] # lambda
hist(log(cal$ground_carcs_total_above))
shapiro.test(log(cal$ground_carcs_total_above)) # not very acceptable
hist(sqrt(cal$ground_carcs_total_above))
shapiro.test(sqrt(cal$ground_carcs_total_above)) # not very acceptable
hist((cal$ground_carcs_total_above)^(1/3))
shapiro.test((cal$ground_carcs_total_above)^(1/3)) # best
```
protration index
```{r}
hist(cal$protraction)
shapiro.test(cal$protraction) # p < 0.01-0.05 means that it is NOT normal
b <- boxcox(lm(cal$protraction ~ 1))
b$x[which.max(b$y)] # lambda
hist(log(cal$protraction))
shapiro.test(log(cal$protraction)) # best
hist(sqrt(cal$protraction))
shapiro.test(sqrt(cal$protraction)) # not acceptable
hist((cal$protraction)^(1/3))
shapiro.test((cal$protraction)^(1/3)) # ok
```

dead vs live index
```{r}
hist(cal$deadvlive)
shapiro.test(cal$deadvlive) # p < 0.01-0.05 means that it is NOT normal
b <- boxcox(lm(cal$deadvlive ~ 1))
b$x[which.max(b$y)] # lambda
hist(log(cal$deadvlive))
shapiro.test(log(cal$deadvlive)) # good
hist(sqrt(cal$deadvlive))
shapiro.test(sqrt(cal$deadvlive)) # great
hist((cal$deadvlive)^(1/3))
shapiro.test((cal$deadvlive)^(1/3)) # best
```

cube-root transformation does the best job of normalizing most of the E Stu data
with the exception of the protraction index
```{r}  
cal <- cal %>% 
  mutate(Cubehpe = hpe^(1/3),
         Cubelpe = lpe^(1/3),
         CubeTotalCarv = ground_carcs_total_above^(1/3),
         lnprotract = log(protraction),
         Cubedeadvlive = deadvlive^(1/3),
         Cubelpe.sc = scale(Cubelpe, center=T, scale=T)[1:nrow(cal)],
         CubeTotalCarc.sc = scale(CubeTotalCarv, center=T, scale=T)[1:nrow(cal)],
         lnprotract.sc = scale(lnprotract, center=T, scale=T)[1:nrow(cal)],
         Cubedeadvlive.sc = scale(Cubedeadvlive, center=T, scale=T)[1:nrow(cal)])
```
### Correlation
note that multiple variables have only 1 state or low variability:
size
stability
water clarity
canopy cover
method
```{r correlation figure}
cal.cor <- round(cor(dplyr::select(cal,
                 substrate_shade_recode, lwd_recode,
                 Cubelpe.sc, CubeTotalCarc.sc, lnprotract.sc,
                 Cubedeadvlive.sc)), 3)
cal.cor

#png(file="figures/E. Stu Sockeye Calibration Correlation Plot.png",
#    units = "in", width=10, height=5, res = 250)

chart.Correlation(dplyr::select(cal,
                 substrate_shade_recode, lwd_recode,
                 Cubelpe.sc, CubeTotalCarc.sc, lnprotract.sc,
                 Cubedeadvlive.sc), 
                  histogram = TRUE, method = "pearson")
#dev.off()
```

Cubed Total Carc is the most correlated. Removing from analysis.


### Covariate Selection
```{r}
m.g <- lmer(Cubehpe ~ Cubelpe.sc + Cubedeadvlive.sc + lnprotract.sc +
              substrate_shade + large_woody_debris + (1|gazetted_stream_name), 
    data=cal,  
    #family = gaussian(link = "identity"), # if using glm
    na.action = "na.fail")

summary(m.g)

plot((fitted(m.g)), residuals(m.g, type = "response"))
# transforming back to "real"
plot((fitted(m.g)^3), residuals(m.g, type = "response"))

m.g.coef <- as.data.frame(summary(m.g)$coefficients)
m.g.coef$parameter = row.names(m.g.coef)
colnames(m.g.coef)[2] = "std.error"

ggplot(m.g.coef,
       aes(x = parameter, y = Estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = Estimate-std.error, ymax = Estimate+std.error)) +
  geom_hline(yintercept = 0) +
  coord_flip()

# Only lpe and deadvsalive appear to be significant

m.global <- lmer(Cubehpe ~ Cubelpe.sc + 
                   Cubedeadvlive.sc*Cubelpe.sc + lnprotract.sc*Cubelpe.sc +
              substrate_shade*Cubelpe.sc + large_woody_debris*Cubelpe.sc + 
                (1|gazetted_stream_name), 
    data=cal,  
    REML = FALSE, # pg 122 of Zuur 2009 - Use ML to compare fixed effects. refit with REML for final parameter estimates.
    na.action = "na.fail")

summary(m.global)
```

#### Variance Inflation Factors
general definition:
https://online.stat.psu.edu/stat462/node/180/
on interactions and categorical variables:
https://statisticalhorizons.com/multicollinearity/
```{r}
vif(m.g)
```
All below 2, which is "OK". Keep all in global model.

#### Dredge
```{r}
m.dredge <- dredge(m.global, beta = "sd", rank = "AICc")
m.dredge

m.dredge <- as.data.frame(m.dredge %>% filter(delta < 4)) %>% 
  dplyr::select(Cubedeadvlive.sc:delta) %>% 
  dplyr::select(-df, -logLik)

write.csv(m.dredge, file = "./tables/01_05-EarlyStuart-Specialized.csv", na = "", row.names = F)
```

Cubedeadvlive interaction is the only one in the top 2.

# Top Model Selection
```{r}
m.interact <- lmer(Cubehpe ~ Cubelpe.sc*Cubedeadvlive.sc + (1|gazetted_stream_name), 
    data=cal,  
    REML = TRUE,
    na.action = "na.fail")

png(file = "./figures/01_Model_Diagnostics/01-05-Early-Stuart-Residuals.png",
    height = 5, width = 5, units = "in", res = 96)
plot(fitted(m.interact), residuals(m.interact),
     xlab = "Fitted Values", ylab = "Residuals",
     main = "Early Stuart Specialized (6) Model")
abline(h = 0)
dev.off()

# possible outliers
cal$usid[which(abs(residuals(m.interact))>3)]
# Influential Data
plot(cooks.distance(m.interact))
abline(h = 4/nrow(cal), col = "blue") # Identifies points with High Influence
abline(h = 3*mean(cooks.distance(m.interact)), col = "red")  # Identifies points with High Influence
# which system-years are possible issues or outliers?
# high leverage
cal$usid[which(cooks.distance(m.interact) > 3*mean(cooks.distance(m.interact)))]

summary(m.interact)

m.interact.coef <- as.data.frame(summary(m.interact)$coefficients)
m.interact.coef$parameter = row.names(m.interact.coef)
colnames(m.interact.coef)[2] = "std.error"

ggplot(filter(m.interact.coef, parameter != "(Intercept)"),
       aes(x = reorder_within(parameter, Estimate,
                              parameter, mean), 
           y = Estimate)) +
  geom_point() +
  scale_x_reordered() +
  geom_errorbar(aes(ymin = Estimate-std.error, ymax = Estimate+std.error),
                width = 0.5) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  labs(x = "Parameter")

ggsave("./figures/01_Model_Diagnostics/01-05-EarlyStuart-Parameters.png",
       width = 5, height = 5, dpi = 96)

m.interact.coef <- dplyr::select(m.interact.coef, parameter, Estimate, std.error) %>% 
  arrange(desc(abs(Estimate))) 
write.csv(m.interact.coef, file = "./tables/01_05-EarlyStuart-Parameters.csv",
          row.names = F)
```

# Leave One Out
https://www.rdocumentation.org/packages/lme4/versions/1.1-23/topics/predict.merMod
https://stats.stackexchange.com/questions/467543/including-random-effects-in-prediction-with-linear-mixed-model
https://stats.stackexchange.com/questions/262277/why-would-you-predict-from-a-mixed-effect-model-without-including-random-effects
```{r}
dplyr::select(cal[1, ], Cubelpe.sc, Cubedeadvlive.sc, gazetted_stream_name)
predict(m.interact, type = "response", se.fit = T)
fitted(m.interact)
p1 <- predict(m.interact,  
              newdata = as.data.frame(dplyr::select(cal[1, ], Cubelpe.sc, 
                                                    Cubedeadvlive.sc)), # not including stream name to remove random effect correction
              re.form=NA, # this allows for "out of group" data to be used, i.e. from a group not part of the random effect
              type="response",
              se.fit = T) # se.fit = F must be used if including the random effect
p1$fit
```

```{r}
model.test <- data.frame(model=character(),
                 usid=character(), 
                 fit=numeric(),
                 se.fit=numeric(),
                 calibrated.estimate=numeric()) 
```

```{r}
models <- list(m.interact)

saveRDS(m.interact, file = "./r_objects/models/Early_Stu_DeadAlive_Model.rds")

specifications_out <- data.frame(model = "Early_Stu_DeadAlive_Model",
                                 target = c("lpe","lpe","lpe",
                                            "deadvlive","deadvlive","deadvlive",
                                            "prediction"),
                                 order = c(1,2,3,
                                           1,2,3,
                                           1),
                                 type = c("cuberoot","center","scale",
                                          "cuberoot","center","scale",
                                          "cube"),
                                 transform = c(NA, mean(cal$Cubelpe), sd(cal$Cubelpe),
                                               NA, mean(cal$Cubedeadvlive), sd(cal$Cubedeadvlive),
                                               NA)) # note that in the future, may want to re-center and scale using the XLONLY Data instead of the whole dataset


spec.out <- spec.out %>% 
  filter(model != "Early_Stu_DeadAlive_Model") %>% 
  rbind(., specifications_out)

saveRDS(spec.out, file=modout)
```
Need to add a running of the model without the one data point.
```{r}
for(j in 1:length(models)){
  model = models[[j]]
  for(i in 1:nrow(cal)){
    m.temp <- lmer(formula(model),
                     data=cal[-i, ],
                     REML = TRUE)
    p.temp <- predict(m.temp, type="response", cal[i, -7], # -7 is to remove the gazetted_stream_name
                      re.form=NA, se.fit=T)
    model.test <- rbind(model.test, data.frame(model = as.character(formula(model))[3],
                                 usid = cal$usid[i],
                                 fit = p.temp$fit,
                                 se.fit = p.temp$se.fit,
                                 calibrated.estimate = p.temp$fit^3))
  }
}
```

```{r}
model.rank <- left_join(model.test, 
                        dplyr::select(cal,
                                      usid, hpe, watershed_group:year, lpe_method)) %>% 
  mutate(error = calibrated.estimate - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error),
         rank.set.level = 6)

```
### Plots
```{r plot of absolute percent error by system}
set.seed(234)
model.rank %>% 
  ggplot(aes(x = reorder(gazetted_stream_name, abs.per.error, FUN = median), 
             # reorder FUN = median, reorders based on the median
             y = abs.per.error,
             colour = model)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_hline(yintercept = 0.5, colour = "grey") +
  geom_point(position = "jitter", alpha = 0.5) +
  coord_flip() +
  labs(title = "Early Stuart Dead/Alive Model Performance",
       y = "Absolute percent error", x = "Streams",
       colour = "Model Types")

#ggsave("figures/Model Performance/02_Model_Performance_Early_Stu_DeadAlive_Model.png",
#       height = 6, width = 8, dpi = 150)
```

```{r}
filter(model.rank,
       abs.per.error > 0.4)
```

# Save Model Ranks
```{r}
saveRDS(model.rank, "r_objects/01_Model_Ranks_from_Early_Stu_DeadAlive_Model.rds")
```

```{r}

```

# Plots for Presentations
```{r}
summary(m.interact)

plot.dat <- expand.grid(
  lpe = c(0:10000),
  deadvlive = c(0.25, 0.5, 1, 2)
) %>% 
  mutate(Cubelpe = lpe^(1/3),
         Cubedeadvlive = deadvlive^(1/3),
         Cubelpe.sc = (Cubelpe - mean(cal$Cubelpe))/sd(cal$Cubelpe),
         Cubedeadvlive.sc = (Cubedeadvlive - mean(cal$Cubedeadvlive))/sd(cal$Cubedeadvlive))

p.temp <- predict(m.interact, type="response", plot.dat, # -7 is to remove the gazetted_stream_name
                      re.form=NA, se.fit=T)

plot.dat$fit = p.temp$fit


```

