---
title: "01-02-Ratio-of-Means"
author: "Michael Arbeider"
date: "2026-03-06"
---

# Set-up
## Packages
```{r}
rm(list = ls())
library(tidyverse) # most %>% 
theme_set(theme_bw()) # ggplot
```

## Functions
```{r}
source("00-Functions.R") # scale_x_reordered, rank.summary(), round2()
```

## Data
```{r}
cal <-readRDS("r_objects/sockeye_calibration_data.rds")   

glimpse(cal)
```

## Specifications FILE
```{r}
modout <- "./r_objects/models/model_specifications.rds"
spec.out <- readRDS(modout)
```

#1d
## Ratio-of-Means
Based on prior methods, looking at ratio-of-means at different groupings.
lpe_method first, then by size, then by substrate_shade.
It is recommended to have 30+ samples per ratio. (Cochran 1977)
Likely few combinations at that "fine detail" with 30 or even 20 + samples.

It is also recommended that the CV of the mean X and Y is < 10% (Corchran 1977)
I think this is only possible when the estimates are log transformed.

Relationships are also not linear with normal distributions when in regular space
Need to transform into log space for this method to work.

rom = "ratio-of-means"
```{r}
count(cal, lpe_method, size, water_clarity) %>% 
  mutate(rom = n >= 20)

count(cal, lpe_method, size, substrate_shade) %>% 
  mutate(rom = n >= 20)

count(cal, lpe_method, size) %>% 
  mutate(rom = n >= 20)

count(cal, size) %>% 
  mutate(rom = n >= 20)

# showing that the CVs of the X and Y are > 10%
sd(log(cal$hpe)) / mean(log(cal$hpe)) * 100
sd((cal$hpe)) / mean((cal$hpe)) * 100 # super huge if not log transformed
sd(log(cal$lpe)) / mean(log(cal$lpe)) * 100

```

Need to decide on Ratio-of-Mean groups.
Most General = All Data
Mid = Aggregate-Size and lpe_method
Specialized = Size and lpe_method, only consider n>20
WC = Water Clarity, Only Consider n>20

```{r}
cal <- cal %>% 
  mutate(rom.gen = paste("All", lpe_method, sep = "."),
         rom.agg = ifelse(size %in% c("Large", "X-Large"), "LXL", "SM"),
         rom.agg = paste(rom.agg, lpe_method, sep = "."),
         rom.size = paste(size, lpe_method, sep = "."),
         rom.wc = paste(size, lpe_method, water_clarity, sep = "."))

count(cal, rom.agg)
count(cal, rom.size)
count(cal, rom.wc)
```

```{r}
# pivoting longer to get a more loop-able data set
cal.long <- cal %>% 
  dplyr::select(sid:water_clarity, lpe_method, hpe, lpe, hpe_error, rom.gen:rom.wc) %>% 
  pivot_longer(
    cols = starts_with("rom."),
    names_to = "model",
    values_to = "grouping")

# Creating several additional subsets
# no rom.gen both methods was made. need to make one
cal.long.add <- cal.long %>% 
  # make everything named the same
  mutate(model = "rom.genB",
         grouping = "All.Both") %>% 
  # filtering to unique results in the OG cal data set
  unique()

# creating rom.estu single models - i.e. 1 ROM model per Estu System
cal.estu.single <- cal.long %>% 
  filter(gazetted_stream_name %in% c("O' Ne-Ell Creek",
                                     "Gluske Creek",
                                     "Forfar Creek")) %>% 
  mutate(model = "rom.estu.single",
         grouping = gazetted_stream_name) %>% 
  # filtering to unique results in the OG cal data set
  unique()
  
# creating rom.estu aggregate model - i.e. 1 ROM model for all of Estu
cal.estu.all <- cal.long %>% 
  filter(run_timing_group == "Early Stuart",
         lpe_method == "Ground",
         water_clarity == "Clear") %>% 
  mutate(model = "rom.estu.all",
         grouping = "Early Stuart") %>% 
  # filtering to unique results in the OG cal data set
  unique()

## Add subsets to the list  
# adding rom.gen to the rest of the model combinations
cal.long <- rbind(cal.long.add, cal.long, cal.estu.single, cal.estu.all)

model.sets <- cal.long %>% 
  group_by(model, grouping) %>% 
  summarise(average = mean(log(hpe))) %>% 
  mutate(rom = NA, n = NA, var = NA, se = NA)

for(i in 1:nrow(model.sets)){
  temp.cal <- cal.long %>% 
    filter(model == model.sets$model[i],
           grouping == model.sets$grouping[i])
  
  m.temp <- temp.cal %>% 
      summarise(ulow = mean(log(lpe)),
                uhigh = mean(log(hpe)),
                rom = mean(log(hpe))/mean(log(lpe)),
                varhi = var(log(hpe)) + sum((hpe_error^2+1)/(hpe^2))/n()^2, # alternative or additional variance? essentially the sum of the transformed variances / the square of a multiplier, which is 1/n for a mean
                # log variance is https://stats.stackexchange.com/questions/418313/variance-of-x-and-variance-of-logx-how-to-relate-them
                varlo = var(log(lpe)),
                covhilo = cov(log(hpe), log(lpe)),
                n = n()) %>% 
      mutate(var.fit = varhi / (ulow^2) + 
      uhigh^2 * varlo / (ulow^4) -
      2 * uhigh * covhilo / ulow^3 )
  if(m.temp$n > 3){
      model.sets$rom[i] <- m.temp$rom
  model.sets$n[i] = m.temp$n
  model.sets$var[i] = m.temp$var.fit
  model.sets$se[i] = sqrt(m.temp$var.fit)
  }
}

model.sets <- model.sets %>% 
  filter(n > 3)
```

### Table of Outputs
```{r}
table01 <- model.sets %>% 
  mutate(average = round2(average, 2),
         rom = round2(rom, 3),
         se = round2(se, 3), 
         CV = round2(sqrt(exp(se^2)-1)*100, 1)) %>% 
  dplyr::select(-var) %>% 
  rename(Set = model, Expansion.Factor = rom, 
         Sample.Size = n, Standard.Error = se) %>% 
  arrange(Set, grouping) %>% 
  mutate(Set = case_when(
    Set == "rom.genB" ~ "Generalized (1)",
    Set == "rom.gen" ~ "Generalized (1)",
    Set == "rom.agg" ~ "Size-Aggregate (2)",
    Set == "rom.size" ~ "Size (3)",
    Set == "rom.wc" ~ "Size-Clarity (4)",
    Set == "rom.estu.single" ~ "Individual (5)",
    Set == "rom.estu.all" ~ "Specialized (6)"
  ),
  grouping = sub(pattern = "V. ", "", grouping),
  grouping = sub(pattern = "LXL", "Large-X-Large", grouping),
  grouping = sub(pattern = "SM", "Small-Medium", grouping)) %>% 
  separate_wider_delim(
    cols = grouping,
    delim = ".",
    names = c("Size", "Method", "Water.Clarity"),
    too_few = "align_start"
  ) %>% 
  dplyr::select(Set:Water.Clarity,Expansion.Factor, Standard.Error,Sample.Size,average,CV); table01

write.csv(table01, "./tables/01_02-Ratio-of-Means-summary.csv",
          row.names = F)
```

Testing the expansion with or without a variance correction
https://stats.stackexchange.com/questions/241187/calculating-standard-deviation-after-log-transformation
```{r}
cal.long <- cal.long %>% 
  left_join(., model.sets) %>% 
  filter(!is.na(rom)) %>% 
  mutate(cal1 = exp(log(lpe)*rom),
         cal2 = exp(log(lpe)*rom+var/2))

cor(dplyr::select(cal.long, hpe, cal1, cal2))
```

```{r}
model.sets <- filter(model.sets,
                     n >= 13, # changed from 20 to 13 to inlcude O' Ne-Ell Creek
                     model != "rom.estu.all") # removing estu.all because it is just too similar with the ROM_WC model
saveRDS(filter(model.sets,
               model == "rom.genB"), 
               file = "./r_objects/models/All_ROM_GenB_Model.rds")
saveRDS(filter(model.sets,
               model == "rom.gen"), 
               file = "./r_objects/models/All_ROM_Gen_Model.rds")
saveRDS(filter(model.sets,
               model == "rom.agg"), 
               file = "./r_objects/models/All_ROM_Agg_Model.rds")
saveRDS(filter(model.sets,
               model == "rom.size"), 
               file = "./r_objects/models/All_ROM_Size_Model.rds")
saveRDS(filter(model.sets,
               model == "rom.wc"), 
               file = "./r_objects/models/All_ROM_WC_Model.rds")
saveRDS(filter(model.sets,
               model == "rom.estu.single"), 
               file = "./r_objects/models/All_ROM_Estu_Model.rds")

specifications_out <- data.frame(model = c("All_ROM_Model"),
                                 target = c("lpe", "hpe"),
                                 order = c(1, 1),
                                 type = c("log", "log"),
                                 transform = c(NA))

spec.out <- spec.out %>% 
  filter(model != "All_ROM_Model") %>% 
  rbind(., specifications_out)

saveRDS(spec.out, file=modout)
```

## Leave One Out
```{r}
model.test <- data.frame(model=character(),
                 usid=character(), 
                 fit=numeric(), 
                 se.fit=numeric(),
                 calibrated.estimate = numeric()) 
```

```{r}
for(i in 1:nrow(model.sets)){
  temp.cal <- cal.long %>% 
    filter(model == model.sets$model[i],
           grouping == model.sets$grouping[i])
  
  for(j in 1:nrow(temp.cal)){
    
    m.temp <- temp.cal[-j, ] %>% 
      summarise(ulow = mean(log(lpe)),
                uhigh = mean(log(hpe)),
                rom = mean(log(hpe))/mean(log(lpe)),
                rom = mean(log(hpe))/mean(log(lpe)),
                varhi = var(log(hpe)),
                varlo = var(log(lpe)),
                covhilo = cov(log(hpe), log(lpe)),
                n = n()) %>% 
      mutate(var.fit = varhi / (ulow^2) + 
      uhigh^2 * varlo / (ulow^4) -
      2 * uhigh * covhilo / ulow^3 )
    
    p.temp <- data.frame(fit = log(temp.cal$lpe[j])*m.temp$rom, 
                         se.fit = sqrt(m.temp$var.fit))
    
    model.test <- rbind(model.test, data.frame(model = model.sets$model[i],
                                 usid = temp.cal$usid[j],
                                 fit = p.temp$fit,
                                 se.fit = p.temp$se.fit,
                        calibrated.estimate = exp(p.temp$fit)))
  }
}
```

```{r}
model.rank <- left_join(model.test, 
                        dplyr::select(cal,
                                      usid, hpe, watershed_group:year, lpe_method)) %>% 
  mutate(error = calibrated.estimate - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error),
         rank.set.level = case_when(
           model == "rom.genB" ~ 1,
           model == "rom.gen" ~ 1,
           model == "rom.agg" ~ 2,
           model == "rom.size" ~ 3,
           model == "rom.wc" ~ 4,
           model == "rom.estu.single" ~ 5
         ))
```

```{r box and whisker of percent error by system}
model.rank %>% 
  ggplot(aes(x = reorder(gazetted_stream_name, abs.per.error, FUN = median), 
             # reorder FUN = median, reorders based on the median
             y = abs.per.error,
             fill = model)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_point(position = "jitter", alpha = 0.5,
             aes(colour = model)) +
  geom_hline(yintercept = 0.5, colour = "grey") +
  coord_flip() +
  labs(title = "All Data Model Performance",
       y = "Absolute percent error", x = "Streams")
```

#### Save Model Ranks
```{r}
saveRDS(filter(model.rank,
               model == "rom.genB"), "r_objects/01_Model_Ranks_from_All_ROM_GenB_Model.rds")
saveRDS(filter(model.rank,
               model == "rom.gen"), "r_objects/01_Model_Ranks_from_All_ROM_Gen_Model.rds")
saveRDS(filter(model.rank,
               model == "rom.agg"), "r_objects/01_Model_Ranks_from_All_ROM_Agg_Model.rds")
saveRDS(filter(model.rank,
               model == "rom.size"), "r_objects/01_Model_Ranks_from_All_ROM_Size_Model.rds")
saveRDS(filter(model.rank,
               model == "rom.wc"), "r_objects/01_Model_Ranks_from_All_ROM_WC_Model.rds")
saveRDS(filter(model.rank,
               model == "rom.estu.single"), "r_objects/01_Model_Ranks_from_All_ROM_Estu_Model.rds")
```

```{r}

```

