---
title: "01-04-Linear-Regressions"
author: "Michael Arbeider"
date: "2026-03-06"
---

There are three Set-Levels where GLM models are developed in this script.
Two models developed at the Generlized (1) Level;
Two models developed at the Size-Aggregate (2) Level and
Eight models developed at the Individual (5) Level, although these models are
developed within the framework of a loop.

Each model section follows the same steps:
Plot correlations between independent variables
Make a Global Model with and without interaction terms
Calculate the Variance Inflation Factors
Drop covariates based on correlations and VIF
Develop a reduced model and compare it against the global models with AICc
Dredge the reduced model
Select a "top" model
Investigate its residuals and outliers and summarize its parameters
Save the model as an rds and calculate the leave-one-out fits
Save the "Model Ranks", which are actually the files used to create model ranks

# Set-up
## Packages
```{r}
rm(list = ls())
library(tidyverse) # most %>% 
theme_set(theme_bw()) # ggplot
library(MASS) # boxcox
library(lme4) # generalized linear regressions
library(MuMIn) # dredge
library(PerformanceAnalytics) # correlation figure
library(car)       # for vif()
library(patchwork) # for multi plot ggplot graphs
library(scales) # labels = percent
```

## Functions
```{r}
source("00-Functions.R") # scale_x_reordered, rank.summary(), round2()
```

## Data
```{r}
cal <-readRDS("r_objects/sockeye_calibration_data.rds")   

glimpse(cal)
```

## Specifications FILE
```{r}
modout <- "./r_objects/models/model_specifications.rds"
spec.out <- readRDS(modout)
```

## Base Plots
https://stackoverflow.com/questions/57153428/r-plot-color-combinations-that-are-colorblind-accessible
```{r colour pallet}
size.cols = c("#E69F00", # Large
              "#56B4E9", # Medium (including the smalls)
              "#D55E00", # V. Small
              "#CC79A7") # X-Large
```

# All Data Linear Regression
Multiple Linear regression models will be explored, starting with ones that
include all data points to "maximize" the amount of information that is
available.
Later, data will be explicitly broken apart to test if covariates behave
differently for different categories of system based on size.

#1
## Global Model
### Link Selection
```{r}
glm.pois <- glm(hpe ~ lpe, data = cal, family = poisson(link = "log"))
glm.nb <- glm.nb(hpe ~ lpe, data = cal, link="log")

summary(glm.pois)
# For Poisson models, it is important to look at residual deviance divided by 
# the degrees of freedom. When this quotient is larger than 1, 
# the negative binomial should be considered.
glm.pois$deviance / glm.pois$df.residual

summary(glm.nb)
glm.nb$deviance / glm.nb$df.residual

arrange(MuMIn::AICc(glm.pois, glm.nb), AICc) %>% 
  mutate(delta = AICc - min(AICc))
```

This is a rough test to determine the type of over dispersion may be in a model
```{r negative binomial}
dg <- cal %>%
  group_by(size) %>%
  summarise(m = mean(hpe), v = var(hpe))
ggplot(dg, aes(m, v)) +
  geom_smooth(method = "lm", formula = y ~ x - 1, se = F, color = "blue") + # neg binom 1
  geom_smooth(method = "lm", 
    formula = y ~ I(x^2) + offset(x) - 1, colour = "red", se = F) + # neg binom 2
  geom_abline(intercept = 0, slope = 1, lty = 2) +
  geom_point()
```
Looks like neither are necessarily supported highly, 
but maybe negative binomial 2 is better for the early data points.

### Covariate Selection
Canopy Cover was deemed to be less helpful due to sudden changes by fire or
other year-to-year variation and that it wouldn't be that influential on
observer efficiency. (statement from a prior report)
```{r}
m.global <- glm.nb(hpe ~ size*lnlpe.sc + system_stability + water_clarity*lnlpe.sc + 
                           substrate_shade + large_woody_debris + 
                           lpe_method*lnlpe.sc 
                   + lnlpe.sc + size + water_clarity + lpe_method, 
                   na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                   data=cal,
                   link="log")
# glm.nb default negative binomial family is neg binom 2

# Model Checking
# https://www.sfu.ca/sasdoc/sashtml/insight/chap39/sect56.htm
# Pearson residual is more appropriate for a glm that doesn't assume gaussian error distribution
plot(log(fitted(m.global)), residuals(m.global, type = "pearson"))
# Model Summary
summary(m.global)

# Global Model with No Interaction for VIFF
m.global.no.interaction <- glm.nb(hpe ~ size + system_stability + water_clarity + 
                           substrate_shade + large_woody_debris + 
                           lpe_method + lnlpe.sc, 
                   na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                   data=cal,  
                   link="log")

# plot(log(fitted(m.global.no.interaction)), residuals(m.global.no.interaction, type = "response"))
plot(log(fitted(m.global.no.interaction)), residuals(m.global.no.interaction, type = "pearson"))
# looks worse than with the interactions

summary(m.global.no.interaction)
```

#### Correlation Coefficients
```{r correlation figure}
cal.cor <- round2(cor(dplyr::select(cal, size_recode, stability_recode, water_clarity_recode, 
                 substrate_shade_recode, canopy_cover_recode, lwd_recode,
                 lpe_method_recode, lnlpe.sc)), 3)
cal.cor

chart.Correlation(dplyr::select(cal, size_recode, stability_recode, 
                                water_clarity_recode, substrate_shade_recode, 
                                canopy_cover_recode, lwd_recode, 
                                lpe_method_recode, lnlpe.sc), 
                  histogram = TRUE, method = "pearson")
```
Size is >0.6 correlated with stability, canopy, large woody debris, lpe_method, and lpe
Stability is >0.6 with canopy and size
Water clarity is 0.4-0.5 with size, stability, canopy, and method
Substrate is < 0.3 with all.
Canopy is > 0.6 with size, stability, lwd, method and >0.5 with lpe
LWD is >0.6 with size, canopy, and lpe
Method is >0.6 with size and canopy
lpe is >0.6 with size, and lwd, and >0.5 with canopy.

#### Variance Inflation Factors
general definition:
https://online.stat.psu.edu/stat462/node/180/
on interactions and categorical variables:
https://statisticalhorizons.com/multicollinearity/
```{r}
vif(m.global)
vif(m.global.no.interaction)
```

Summary:
lnlpe.sc has the highest VIF, but likely because it is directly interacting
with size.
Generally, categorical variables may also have high VIFs due to their nature.

For now, I am keeping with the interpretation without the interaction term.

Summary without the interaction term:
system stability had the highest VIF and is highly correlated with size.
size is also highly correlated with large woody debris. Size has the second
highest VIF after system stability.

One option would be to remove size, but I propose that it is more important than
stability and LWD for estimating observer efficiency for Sockeye. I could see
LWD being important for Coho but not necessarily for Sockeye.

Considering Size as a primary variable of interest, only water clarity and
substrate shade are < 0.5 correlated with it.
There is also a base inclination that method ought to be important, so also
retaining it despite its high correlation with size.

## Top Model
```{r}
m.size.all.slp <- glm.nb(hpe ~ size*lnlpe.sc + water_clarity*lnlpe.sc + 
                           substrate_shade*lnlpe.sc + lpe_method*lnlpe.sc +
                           water_clarity + substrate_shade + 
                           lpe_method + size + lnlpe.sc, 
                   na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                   data=cal,  
                   link="log")

png(file = "./figures/01_Model_Diagnostics/01-04-GeneralizedAllData-Residuals.png",
    height = 5, width = 5, units = "in", res = 96)
plot(log(fitted(m.size.all.slp)), residuals(m.size.all.slp, type = "pearson"),
     xlab = "ln(Fitted Values)", ylab = "Pearson Residuals",
     main = "Generalized (1) - Clarity GLM")
abline(h = 0)
dev.off()
# possible outliers
cal$usid[which(abs(residuals(m.size.all.slp, type = "pearson"))>3)]
# Influential Data
plot(cooks.distance(m.size.all.slp))
abline(h = 4/nrow(cal), col = "blue") # Identifies points with High Influence
abline(h = 3*mean(cooks.distance(m.size.all.slp)), col = "red")  # Identifies points with High Influence
# which system-years are possible issues or outliers?
# high leverage
cal$usid[which(cooks.distance(m.size.all.slp) > 3*mean(cooks.distance(m.size.all.slp)))]

summary(m.size.all.slp)

m.coef <- as.data.frame(summary(m.size.all.slp)$coefficients)
m.coef$parameter = row.names(m.coef)
colnames(m.coef)[2] = "std.error"

ggplot(filter(m.coef, parameter != "(Intercept)"),
       aes(x = reorder_within(parameter, Estimate,
                              parameter, mean), 
           y = Estimate)) +
  geom_point() +
  scale_x_reordered() +
  geom_errorbar(aes(ymin = Estimate-std.error, ymax = Estimate+std.error),
                width = 0.5) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  labs(x = "Parameter")

ggsave("./figures/01_Model_Diagnostics/01-04-GeneralizedAllData-Parameters.png",
       width = 5, height = 5, dpi = 96)

m.coef <- dplyr::select(m.coef, parameter, Estimate, std.error) %>% 
  arrange(desc(abs(Estimate))) %>% 
  mutate(cross.zero = ifelse(abs(Estimate) < abs(std.error), "Yes", "No"))
write.csv(m.coef, file = "./tables/01_04-GeneralizedAll-Parameters.csv",
          row.names = F)
```

### Comparing Size Models
```{r comparing global models}
# these are the TMB models
arrange(MuMIn::AICc(m.size.all.slp,
                    m.global, m.global.no.interaction), AICc) %>% 
  mutate(delta = AICc - min(AICc))

# size interaction with slope excluding system stability and large woody debris (m.size.slp)
# is most parsimonious relative to global within 2 AICc.
```

```{r}
m.dredge <- dredge(m.size.all.slp, beta = "none", rank = "AICc")
m.dredge

m.dredge <- as.data.frame(m.dredge %>% filter(delta < 4)) %>% 
  dplyr::select(lnlpe.sc:delta)

write.csv(m.dredge, file = "./tables/01_04-GeneralizedAll.csv", na = "", row.names = F)
```
Top model from original selection (m.size.all.slp).
It is within 2 AICc of the top models.
Not going with most parsimonious as I am interested in the additional explanatory
power of covariates.
Slight risk of an overfit model, but parameter estimates should be more accurate.

## Leave One Out
```{r}
model.test <- data.frame(model=character(),
                 usid=character(), 
                 fit=numeric(), 
                 se.fit=numeric(),
                 calibrated.estimate=numeric()) 
```

```{r}
saveRDS(m.size.all.slp, file = "./r_objects/models/All_Clarity_Model.rds")

specifications_out <- data.frame(model = c("All_Clarity_Model"),
                                 target = c("lpe","lpe","lpe"),
                                 order = c(1,2,3),
                                 type = c("log","center","scale"),
                                 transform = c(NA, mean(cal$lnlpe), sd(cal$lnlpe)))

spec.out <- spec.out %>% 
  filter(model != "All_Clarity_Model") %>% 
  rbind(., specifications_out)

saveRDS(spec.out, file=modout)
```
Need to add a running of the model without the one data point.
```{r}
model = m.size.all.slp
for(i in 1:nrow(cal)){
    m.temp <- glm.nb(as.character(model$call)[2],
                     data=cal[-i, ],
                     na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                     link="log")
    p.temp <- predict(m.temp, type="link", cal[i, ], se.fit=T)
    model.test <- rbind(model.test, data.frame(model = as.character(model$call)[2],
                                 usid = cal$usid[i],
                                 fit = p.temp$fit,
                                 se.fit = p.temp$se.fit,
                        calibrated.estimate = exp(p.temp$fit)))

}
```

```{r}
model.rank <- left_join(model.test, 
                        dplyr::select(cal,
                                      usid, hpe, watershed_group:year, lpe_method)) %>% 
  mutate(error = calibrated.estimate - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error),
         rank.set.level = 1)

```
### Plots
```{r box and whisker of percent error by system}
model.rank %>% 
  ggplot(aes(x = reorder(gazetted_stream_name, abs.per.error, FUN = median), 
             # reorder FUN = median, reorders based on the median
             y = abs.per.error)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_point(position = "jitter", alpha = 0.5,
             aes(colour = lpe_method)) +
  geom_hline(yintercept = 0.5, colour = "grey") +
  coord_flip() +
  labs(title = "All Data Model Performance",
       y = "Absolute percent error", x = "Streams",
       colour = "Survey Method")
```
Easy to pick out that Middle River, Scotch Creek, Nadina, Tachie, Harrison, 
Stellako, and maybe Gluske are the biggest problem children
Either do to the number of > 1 absolute percent errors or generally just the
spread of errors.

### Save Model Ranks
```{r}
saveRDS(model.rank, "r_objects/01_Model_Ranks_from_All_Clarity_Model.rds")
```

### Lower < Training
This section is to investigate out-of-range behaviour of the model.
"Out-of-range" with regards to visual counts being lower than the training data
```{r}
# creating a data set to show behaviour near the origin
dat.origin <- expand_grid(size = unique(cal$size),
                          water_clarity = unique(cal$water_clarity),
                          substrate_shade = unique(cal$substrate_shade),
                          lpe_method = unique(cal$lpe_method),
                          lpe = seq(1,25000, by = 1)) %>% 
  mutate(lnlpe.sc = (log(lpe) - mean(cal$lnlpe)) / sd(cal$lnlpe) )

# create model fits
dat.origin.fit <- predict(m.size.all.slp, type="link", dat.origin, se.fit=T)

# add model fits as well as comparison to default 1.8 method
dat.origin <- dat.origin %>% 
  mutate(fit = dat.origin.fit$fit,
         se.fit = dat.origin.fit$se.fit,
         expanded = exp(fit),
         expanded.lower = exp(fit - se.fit),
         expanded.upper = exp(fit + se.fit),
         default = lpe*1.8,
         dif = expanded - default,
         per.dif = dif / default,
         ratio = expanded / lpe) %>% 
  mutate(substrate_shade = as.factor(substrate_shade),
         substrate_shade = factor(substrate_shade, levels = c("Light",
                                                              "Medium",
                                                              "Dark")),
         size = as.factor(size),
         size = factor(size, levels = c("Small", "Medium", "Large", "X-Large")))

# Figure A
# # All 4 Sizes up to 200
dat.origin %>% 
  filter(lpe_method == "Ground",
         #size %in% c("Medium", "Small"),
         lpe < 200) %>% 
  ggplot(aes(x = lpe,
             y = expanded, colour = water_clarity)) +
  geom_line() +
  geom_line(aes(y = default, colour = "Default")) +
  geom_abline(lty = 2) +
  scale_colour_manual(name = "Water Clarity \n or Default",
                      values = c("grey", "red", "black"),
                      labels = c("Clear", "Default", "Turbid/Tannic")) +
  facet_grid(size~substrate_shade, scales = "free_y") +
  labs(x = "Visual Count", y = "Expanded Estimate")

ggsave("./figures/01_Model_Diagnostics/01-04-GeneralizedAllData-Origin.png",
       width = 6, height = 5, dpi = 150)

# Figure B
# # Small up to 10,000
dat.origin %>% 
  filter(size %in% c("Small"),
         lpe < 10001) %>% 
  ggplot(aes(x = lpe,
             y = expanded, colour = water_clarity)) +
  geom_line() +
  geom_line(aes(y = default, colour = "Default")) +
  geom_abline(lty = 2) +
  facet_grid(lpe_method~substrate_shade) +
  labs(title = "Size: Small")

ggsave("./figures/01_Model_Diagnostics/01-04-GeneralizedAllData-Origin-Small.png",
       width = 7, height = 3.5, dpi = 150)

# Figure C1
# # Small up to 20,000 only Medium Substrate
dat.origin %>% 
  filter(size %in% c("Small"),
         substrate_shade == "Medium") %>% 
  ggplot(aes(x = lpe,
             y = expanded, colour = water_clarity)) +
  geom_line() +
  geom_line(aes(y = default, colour = "Default")) +
  geom_abline(lty = 2) +
  facet_grid(lpe_method~substrate_shade) +
  labs(title = "Size: Small")

ggsave("./figures/01_Model_Diagnostics/01-04-GeneralizedAllData-Origin-Small-ZoomOut.png",
       width = 5, height = 3.5, dpi = 150)

# Figure C2
# # Small up to 500 only Medium Substrate
dat.origin %>% 
  filter(size %in% c("Small"),
         substrate_shade == "Medium",
         lpe < 501) %>% 
  ggplot(aes(x = lpe,
             y = expanded, colour = water_clarity)) +
  geom_line() +
  geom_line(aes(y = default, colour = "Default")) +
  geom_abline(lty = 2) +
  facet_grid(lpe_method~substrate_shade) +
  labs(title = "Size: Small")

ggsave("./figures/01_Model_Diagnostics/01-04-GeneralizedAllData-Origin-Small-ZoomIn.png",
       width = 3.75, height = 3.5, dpi = 150)

# Summarizing the percent difference between the clear and turbid estimates
dat.origin.summary <- dat.origin %>% 
  dplyr::select(size:lpe,expanded) %>% 
  filter(size == "Small") %>% 
  pivot_wider(names_from = water_clarity,
              values_from = expanded) %>% 
  mutate(dif = Pt_Turb_Tannic - Clear,
         per.dif = (dif / Clear)*100)

filter(dat.origin.summary, lpe == 500)

# Figure D
# # Figure of the Percent Difference between Clear and Turbid
dat.origin.summary %>% 
  ggplot(aes(x = lpe, y = per.dif)) +
  geom_line() +
  facet_grid(lpe_method~substrate_shade) +
  labs(title = "Size: Small", 
       y = "Percent Difference Between \n Clear to Turbid Tannic")
ggsave("./figures/01_Model_Diagnostics/01-04-GeneralizedAllData-Origin-Percent-Diff.png",
       width = 7, height = 3.75, dpi = 150)

# Figure E
# # Figure of Percent Difference between Clear and 1.8
  dat.origin %>% 
  filter(size %in% c("Small"),
         substrate_shade == "Medium") %>% 
  ggplot(aes(x = lpe, y = per.dif)) +
  geom_line() +
  facet_grid(lpe_method~water_clarity, scales = "free_y") +
  scale_y_continuous(labels = percent) +
  geom_hline(yintercept = 1, lty = 2, col = "grey") +
  geom_hline(yintercept = 2, lty = 2, col = "grey20") +
  labs(title = "Clarity GLM for Size: Small - Relative to 1.8", 
       y = "Percent Difference Between \n GLM to 1.8 Default")
ggsave("./figures/01_Model_Diagnostics/01-04-GeneralizedAllData-Origin-Percent-Diff-Rel-Default.png",
       width = 5, height = 4, dpi = 150)

dat.origin %>% 
  filter(size %in% c("Small"),
         substrate_shade == "Medium",
         lpe == 500) %>% 
  group_by(size, substrate_shade, lpe_method, water_clarity) %>% 
  summarise(min.per.dif = min(per.dif),
            max.per.dif = max(per.dif))

# Summarizing the percent difference between the ground and aerial estimates
dat.origin.summary.method <- dat.origin %>% 
  dplyr::select(size:lpe,expanded) %>% 
  filter(size == "Small") %>% 
  pivot_wider(names_from = lpe_method,
              values_from = expanded) %>% 
  mutate(dif = Aerial - Ground,
         per.dif = (dif / Ground)*100)

filter(dat.origin.summary.method, lpe == 500)

# Summarizing the percent difference between the clear and turbid estimates AND Method
dat.origin.summary.method.clarity <- dat.origin %>% 
  dplyr::select(size:lpe,expanded) %>% 
  filter(size == "Small") %>% 
  pivot_wider(names_from = c(water_clarity, lpe_method),
              values_from = expanded) %>% 
  mutate(dif = Pt_Turb_Tannic_Aerial - Clear_Ground,
         per.dif = (dif / Clear_Ground)*100)

filter(dat.origin.summary.method.clarity, lpe == 1)
filter(dat.origin,
       lpe == 2230,
       size == "Small",
       substrate_shade == "Light")
```

#### Additional Figures
```{r}
minimums <- readRDS("./r_objects/00-Regression_LPE_Ranges.rds")
minimums$size <- factor(minimums$size, levels = c("Small", "Medium", 
                                                  "Large", "X-Large"))

minimums2 <- minimums %>% 
  rename(lpe = min10) %>% 
  left_join(., 
            dplyr::select(dat.origin,
                         size, lpe_method, lpe, substrate_shade, water_clarity,
                         lnlpe.sc, fit, se.fit, expanded, ratio))

cal %>% 
  filter(lpe < 25000) %>% 
  ggplot(aes(x = lpe, y = hpe/lpe, 
             colour = water_clarity)) +
  geom_hline(yintercept = 1.8, col = "red", alpha = 0.5) +
  geom_point(size = 2) +
  scale_colour_manual(name = "Water Clarity",
                      values = c("gray", 
                                 "black"),
                      labels = c("Clear", "Turbid/Tannic")) +
  ggnewscale::new_scale_colour() +
  geom_point(aes(colour = size)) +
  geom_vline(data = minimums[-nrow(minimums),],
             aes(xintercept = min10,
                 colour = size)) +
  scale_colour_manual(name = "Size",
                      values = size.cols) +
  ggnewscale::new_scale_colour() +
  geom_line(data = filter(dat.origin,
                          size == "Small",
                          substrate_shade == "Medium"),
            aes(x = lpe, y = ratio, colour = water_clarity)) +
    scale_colour_manual(name = "Water Clarity",
                      values = c("gray", 
                                 "black"),
                      labels = c("Clear", "Turbid/Tannic")) +
  facet_wrap(~lpe_method) +
  labs(y = "Expansion Factor", x = "Visual Count") 

ggsave("./figures/LPE_Clarity_Size_Method_Ratios.png",
       height = 4.5, width = 10, dpi = 150)

# < 10000
cal %>% 
  filter(lpe < 10000) %>% 
  ggplot(aes(x = lpe, y = hpe/lpe, 
             colour = water_clarity)) +
  geom_hline(yintercept = 1.8, col = "red", alpha = 0.5) +
  geom_point(size = 2) +
  geom_point(aes(colour = size)) +
  scale_colour_manual(values = c("gray", size.cols[c(2)], 
                                 "black", size.cols[c(3,4)])) +
  facet_wrap(~lpe_method) +
  geom_vline(data = filter(minimums, min10 < 10000),
             aes(xintercept = min10,
                 colour = size)) +
  geom_line(data = filter(dat.origin,
                          size == "Small",
                          substrate_shade == "Medium",
                          lpe < 10000),
            aes(x = lpe, y = ratio)) +
  labs(y = "Expansion Factor", x = "Visual Count") 

ggsave("./figures/LPE_Clarity_Size_Method_Ratios_10000.png",
       height = 4.5, width = 10, dpi = 150)

# X-Large <25000
cal %>% 
  filter(lpe < 25000) %>% 
  ggplot(aes(x = lpe, y = hpe/lpe, 
             colour = water_clarity)) +
  geom_hline(yintercept = 1.8, col = "red", alpha = 0.5) +
  geom_point(size = 2) +
  geom_point(aes(colour = size)) +
  scale_colour_manual(values = c("gray", size.cols[c(1,2)], 
                                 "black", size.cols[c(3,4)])) +
  facet_wrap(~lpe_method) +
  geom_vline(data = filter(minimums, min10 < 10000),
             aes(xintercept = min10,
                 colour = size)) +
  geom_line(data = filter(dat.origin,
                          size == "X-Large",
                          substrate_shade == "Medium",
                          lpe < 25000),
            aes(x = lpe, y = ratio)) +
  ylim(c(0, 42)) +
  labs(y = "Expansion Factor", x = "Visual Count",
       title = "X-Large")

# Large <25000
cal %>% 
  filter(lpe < 25000) %>% 
  ggplot(aes(x = lpe, y = hpe/lpe, 
             colour = water_clarity)) +
  geom_hline(yintercept = 1.8, col = "red", alpha = 0.5) +
  geom_point(size = 2) +
  geom_point(aes(colour = size)) +
  scale_colour_manual(values = c("gray", size.cols[c(1,2)], 
                                 "black", size.cols[c(3,4)])) +
  facet_wrap(~lpe_method) +
  geom_vline(data = filter(minimums, min10 < 10000),
             aes(xintercept = min10,
                 colour = size)) +
  geom_line(data = filter(dat.origin,
                          size == "Large",
                          substrate_shade == "Medium",
                          lpe < 25000),
            aes(x = lpe, y = ratio)) +
  ylim(c(0, 42)) +
  labs(y = "Expansion Factor", x = "Visual Count",
       title = "Large")
```

### Ratios for Low Abundance
```{r}
# creating low abundance subset-based ratio estimators
m.ratios <- cal %>% 
  filter(lpe <= 10000) %>% 
  group_by(water_clarity) %>% 
  summarise(MoR = mean(hpe/lpe),
            MedianOR = median(hpe/lpe),
            ROM = mean(log(hpe)) / mean(log(lpe)),
            RoMedians = median(hpe) / median(lpe),
            n = n())

# adding it to the origin data grid
dat.origin <- left_join(dat.origin, m.ratios)

# calculating the ratio based expanded estimates
dat.origin <- dat.origin %>% 
  mutate(MoR.expanded = MoR*lpe,
         MedianOR.expanded = MedianOR*lpe,
         ROM.expanded = exp(log(lpe)*ROM),
         RoMedians.expanded = lpe*RoMedians)

# selecting the specific expanded results and pivoting them longer for easier plotting
dat.origin2 <- dat.origin %>% 
  dplyr::select(size:lpe, expanded, default, MoR.expanded:RoMedians.expanded) %>% 
  rename(glm.expanded = expanded, default.expanded = default) %>% 
  pivot_longer(cols = glm.expanded:RoMedians.expanded,
               names_to = "Model",
               values_to = "Expanded",
               names_pattern = "(.*)\\.expanded")

# Figure A
# Small comparing different ratio types < 500
dat.origin2 %>% 
  filter(size == "Small",
         substrate_shade == "Medium",
         lpe_method == "Ground",
         lpe < 500,
         Model != "default") %>% 
  ggplot(aes(x = lpe, y = Expanded, colour = water_clarity)) +
  geom_line() +
  geom_line(aes(y = lpe*1.8, colour = "Default"), alpha = 0.8) +
  geom_abline(lty = 2) +
  #scale_colour_manual(values = c("#E69F00",  "#009E73", 
  #               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) +
  facet_grid(~Model) +
  labs(title = "Size: Small")

ggsave("./figures/01_Model_Diagnostics/01-04-GeneralizedAllData-Origin-Ratio-Compare.png",
       width = 9, height = 2.5, dpi = 150)

# Figure B
# Small comparing different ratio types up to 10,000
dat.origin2 %>% 
  filter(size == "Small",
         substrate_shade == "Medium",
         lpe_method == "Ground",
         #lpe < 500,
         Model != "default") %>% 
  ggplot(aes(x = lpe, y = Expanded, colour = water_clarity)) +
  geom_line() +
  geom_line(aes(y = lpe*1.8, colour = "Default"), alpha = 0.8) +
  geom_abline(lty = 2) +
  #scale_colour_manual(values = c("#E69F00",  "#009E73", 
  #               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) +
  facet_grid(~Model) +
  labs(title = "Size: Small")

ggsave("./figures/01_Model_Diagnostics/01-04-GeneralizedAllData-Origin-Ratio-Compare-Full.png",
       width = 10, height = 2.5, dpi = 150)
```

## 1.8 Expansion
```{r}
model.rank2 <- model.rank %>% 
  left_join(., dplyr::select(cal, usid, lpe)) %>% 
    mutate(fit = lpe*1.8,
           error = fit - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error),
         se.fit = sd(error), # just a quick made up place holder for 1.8 expansion se.fit error
         calibrated.estimate = fit,
         model = "hpe ~ lpe*1.8",
         rank.set.level = 1) %>% 
      dplyr::select(-lpe)
```
### Save Model Ranks
```{r}
saveRDS(model.rank2, "r_objects/01_Model_Ranks_from_All_Data_Classic_Expansion.rds")
```

#2
## All Minimal Model
Noted that recovering survey data such as water clarity was not easily done.
Created a model without water clarity and method.
Only Substrate Shade and Size as those are fixed by the system type.
```{r}
m.reduc.g <- glm.nb(hpe ~ lnlpe.sc + size + substrate_shade + 
                      size*lnlpe.sc + substrate_shade*lnlpe.sc, 
    data=cal, 
    na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
    link="log")
summary(m.reduc.g)

png(file = "./figures/01_Model_Diagnostics/01-04-GeneralizedMinimal-Residuals.png",
    height = 5, width = 5, units = "in", res = 96)
plot(log(fitted(m.reduc.g)), residuals(m.reduc.g, type = "pearson"),
     xlab = "ln(Fitted Values)", ylab = "Pearson Residuals",
     main = "Generalized (1) - Minimal GLM")
abline(h = 0)
dev.off()

# possible outliers
cal$usid[which(abs(residuals(m.reduc.g, type = "pearson"))>3)]
# Influential Data
plot(cooks.distance(m.reduc.g))
abline(h = 4/nrow(cal), col = "blue") # Identifies points with High Influence
abline(h = 3*mean(cooks.distance(m.reduc.g)), col = "red")  # Identifies points with High Influence
# which system-years are possible issues or outliers?
# high leverage
cal$usid[which(cooks.distance(m.reduc.g) > 3*mean(cooks.distance(m.reduc.g)))]

m.coef <- as.data.frame(summary(m.reduc.g)$coefficients)
m.coef$parameter = row.names(m.coef)
colnames(m.coef)[2] = "std.error"

ggplot(filter(m.coef, parameter != "(Intercept)"),
       aes(x = reorder_within(parameter, Estimate,
                              parameter, mean), 
           y = Estimate)) +
  geom_point() +
  scale_x_reordered() +
  geom_errorbar(aes(ymin = Estimate-std.error, ymax = Estimate+std.error),
                width = 0.5) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  labs(x = "Parameter")

ggsave("./figures/01_Model_Diagnostics/01-04-GeneralizedMinimal-Parameters.png",
       width = 5, height = 5, dpi = 96)

m.coef <- dplyr::select(m.coef, parameter, Estimate, std.error) %>% 
  arrange(desc(abs(Estimate))) %>% 
  mutate(cross.zero = ifelse(abs(Estimate) < abs(std.error), "Yes", "No"))
write.csv(m.coef, file = "./tables/01_04-GeneralizedMinimal-Parameters.csv",
          row.names = F)

m.dredge <- dredge(m.reduc.g, beta = "none", rank = "AICc")
m.dredge
# global is top
# 
m.dredge <- as.data.frame(m.dredge %>% filter(delta < 5)) %>% 
  dplyr::select(lnlpe.sc:delta)

write.csv(m.dredge, file = "./tables/01_04-GeneralizedMinimal.csv", na = "", row.names = F)
```

## Leave One Out
```{r}
model.test <- data.frame(model=character(),
                 usid=character(), 
                 fit=numeric(), 
                 se.fit=numeric(),
                 calibrated.estimate=numeric()) 
```

```{r}
saveRDS(m.reduc.g, file = "./r_objects/models/All_Minimal_Model.rds")

specifications_out <- data.frame(model = c("All_Minimal_Model"),
                                 target = c("lpe","lpe","lpe"),
                                 order = c(1,2,3),
                                 type = c("log","center","scale"),
                                 transform = c(NA, mean(cal$lnlpe), sd(cal$lnlpe)))

spec.out <- spec.out %>% 
  filter(model != "All_Minimal_Model") %>% 
  rbind(., specifications_out)

saveRDS(spec.out, file=modout)
```
Need to add a running of the model without the one data point.
```{r}
model = m.reduc.g
for(i in 1:nrow(cal)){
    m.temp <- glm.nb(as.character(model$call)[2],
                     data=cal[-i, ],
                     na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                     link="log")
    p.temp <- predict(m.temp, type="link", cal[i, ], se.fit=T)
    model.test <- rbind(model.test, data.frame(model = as.character(model$call)[2],
                                 usid = cal$usid[i],
                                 fit = p.temp$fit,
                                 se.fit = p.temp$se.fit,
                                 calibrated.estimate = exp(p.temp$fit)))

}
```

```{r}
model.rank <- left_join(model.test, 
                        dplyr::select(cal,
                                      usid, hpe, watershed_group:year, lpe_method)) %>% 
  mutate(error = calibrated.estimate - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error),
         rank.set.level = 1)

```
### Plots
```{r box and whisker of percent error by system}
model.rank %>% 
  ggplot(aes(x = reorder(gazetted_stream_name, abs.per.error, FUN = median), 
             # reorder FUN = median, reorders based on the median
             y = abs.per.error)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_point(position = "jitter", alpha = 0.5,
             aes(colour = lpe_method)) +
  geom_hline(yintercept = 0.5, colour = "grey") +
  coord_flip() +
  labs(title = "All Data Model Performance",
       y = "Absolute percent error", x = "Streams",
       colour = "Survey Method")

#ggsave("figures/Model Performance/01_Model_Performance_All_Minimal_Model.png",
#       height = 6, width = 8, dpi = 150)
```

#### Save Model Ranks
```{r}
saveRDS(model.rank, "r_objects/01_Model_Ranks_from_All_Minimal_Model.rds")
```

### Higher > Training
```{r}
# reading in the ROM General Models
rom <- readRDS("./r_objects/models/All_ROM_Gen_Model.rds")
romb <- readRDS("./r_objects/models/All_ROM_GenB_Model.rds")
rom$lpe_method <- c("Aerial", "Ground")

# creating a data set to show behaviour near the origin
dat.abund <- expand_grid(size = unique(cal$size),
                          #water_clarity = unique(cal$water_clarity),
                          substrate_shade = unique(cal$substrate_shade),
                          lpe_method = unique(cal$lpe_method),
                          lpe = seq(75000,500000, by = 1000)) %>% 
  mutate(lnlpe = log(lpe),
         lnlpe.sc = (log(lpe) - mean(cal$lnlpe)) / sd(cal$lnlpe) ) %>% 
  left_join(., rom) %>% 
  mutate(rom.fit = lnlpe*rom,
         rom.expanded = exp(rom.fit),
         romb.fit = lnlpe*romb$rom,
         romb.expanded = exp(romb.fit))

# create model fits
dat.abund.fit <- predict(m.reduc.g, type="link", dat.abund, se.fit=T)

# add model fits as well as comparison to default 1.8 method
dat.abund <- dat.abund %>% 
  mutate(glm.fit = dat.abund.fit$fit,
         se.fit = dat.abund.fit$se.fit,
         glm.expanded = exp(glm.fit),
         expanded.lower = exp(glm.fit - se.fit),
         expanded.upper = exp(glm.fit + se.fit),
         default.expanded = lpe*1.8,
         dif = glm.expanded - default.expanded,
         per.dif = dif / default.expanded) %>% 
  mutate(substrate_shade = as.factor(substrate_shade),
         substrate_shade = factor(substrate_shade, levels = c("Light",
                                                              "Medium",
                                                              "Dark")),
         size = as.factor(size),
         size = factor(size, levels = c("Small", "Medium", "Large", "X-Large")))

# selecting the specific expanded results and pivoting them longer for easier plotting
dat.abund2 <- dat.abund %>% 
  dplyr::select(size:lpe, rom.expanded, romb.expanded, glm.expanded, default.expanded) %>% 
  pivot_longer(cols = rom.expanded:default.expanded,
               names_to = "Model",
               values_to = "Expanded",
               names_pattern = "(.*)\\.expanded") %>% 
  mutate(sub.mod.filter = paste0(substrate_shade, ".", Model),
         Model = as.factor(Model),
         Model = factor(Model, levels = c("glm", "default", "rom", "romb")))

# Stop figures from having scientific notation
options(scipen=999)

# Figure A
# Excluding Small - Compare ROM with GLM - Ground
dat.abund2 %>% 
  filter(lpe_method == "Ground",
         Model != "romb") %>% 
  ggplot(aes(x = lpe, y = Expanded, colour = Model)) +
  geom_line() +
  geom_abline(lty = 2) +
  geom_vline(xintercept = 300000, lty = 3, alpha = 0.5) +
  scale_colour_manual(values = c("orange", "red", "#56B4E9"),
                      labels = c("Minimal GLM",
                                 "Defualt 1.8",
                                 "Ratio of Means \n General Ground")) +
  facet_grid(size~substrate_shade) +
  labs(title = "Ground", x = "Visual Count", y = "Expanded Estimate")

ggsave("./figures/01_Model_Diagnostics/01-04-GeneralizedAllData-HighAbundance-Compare-Ground.png",
       width = 9, height = 7, dpi = 150)

# Figure B
# Excluding Small - Compare ROM with GLM - Aerial
dat.abund2 %>% 
  filter(lpe_method == "Aerial",
         Model != "rom") %>% 
  ggplot(aes(x = lpe, y = Expanded, colour = Model)) +
  geom_line() +
  geom_abline(lty = 2) +
  geom_vline(xintercept = 300000, lty = 3, alpha = 0.5) +
  scale_colour_manual(values = c("orange", "red", "#009E73"),
                      labels = c("Minimal GLM",
                                 "Defualt 1.8",
                                 "Ratio of Means \n General Both")) +
  facet_grid(size~substrate_shade) +
  labs(title = "Aerial", x = "Visual Count", y = "Expanded Estimate")

ggsave("./figures/01_Model_Diagnostics/01-04-GeneralizedAllData-HighAbundance-Compare-Aerial.png",
       width = 9, height = 7, dpi = 150)


# Figure A2 SMALL
# Excluding Small - Compare ROM with GLM - Ground < 200,000
dat.abund2 %>% 
  filter(lpe_method == "Ground",
         size != "Small",
         lpe < 200001) %>% 
  ggplot(aes(x = lpe, y = Expanded, colour = Model)) +
  geom_line() +
  geom_abline(lty = 2) +
  #scale_colour_manual(values = c("#E69F00",  "#009E73", 
  #               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) +
  facet_grid(size~substrate_shade) +
  labs(title = "Ground")

ggsave("./figures/01_Model_Diagnostics/01-04-GeneralizedAllData-HighAbundance-Compare-Ground-Small.png",
       width = 12, height = 7, dpi = 150)

# Figure B2 SMALL
# Excluding Small - Compare ROM with GLM - Aerial  < 200,000
dat.abund2 %>% 
  filter(lpe_method == "Aerial",
         size != "Small",
         lpe < 200001) %>% 
  ggplot(aes(x = lpe, y = Expanded, colour = Model)) +
  geom_line() +
  geom_abline(lty = 2) +
  #scale_colour_manual(values = c("#E69F00",  "#009E73", 
  #               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) +
  facet_grid(size~substrate_shade) +
  labs(title = "Aerial")

ggsave("./figures/01_Model_Diagnostics/01-04-GeneralizedAllData-HighAbundance-Compare-Aerial-Small.png",
       width = 12, height = 7, dpi = 150)
```

# Size Aggregate Models
# 3
# SM
Removing systems that are big outliers such as Harrison and Tachie and Middle
All of these systems are X-LargeL as well

Little River is the only other system in the X-Large category but it only has
four samples, so I am removing all X-Large and Large.

Later determined that Adding Large Data may improve the fit. (?)
Large Data was also quite an outlier relative to the Medium and Small Data though.
Decided on removing Large and X-Large.
```{r}
cal2 <- cal %>% 
  filter(size != "X-Large",
         size != "Large") %>% 
  mutate(lnlpe.sc = (lnlpe-mean(lnlpe))/sd(lnlpe))
```

```{r}
cal2 %>%  
  ggplot(aes(x = lnlpe.sc, y = log(hpe), colour = size)) +
  geom_point() +
  labs(x = "ln(Scaled - Low Precision Estimate)", 
       y = "ln(High Precision Estimate)") +
  scale_colour_manual(values = size.cols[-c(1,4)])
```

## Correlation Coefficients
```{r correlation figure}
cal2.cor <- round2(cor(dplyr::select(cal2, size_recode, stability_recode, water_clarity_recode, 
                 substrate_shade_recode, canopy_cover_recode, lwd_recode,
                 lpe_method_recode, lnlpe.sc)), 3)
cal2.cor

#png(file="figures/Sockeye Calibration Correlation Plot no XL.png",
#    units = "in", width=10, height=5, res = 250)

chart.Correlation(dplyr::select(cal2, size_recode, stability_recode, 
                                water_clarity_recode, substrate_shade_recode, 
                                canopy_cover_recode, lwd_recode, 
                                lpe_method_recode, lnlpe.sc), 
                  histogram = TRUE, method = "pearson")
#dev.off()
```
size is correlated with canopy, lwd, lpe_method, and lpe
stability correlated with clarity and canopy
clarity is somewhat with stability, canopy and lpe_method
shade is good
canopy is with size, stability, lwd, lpe_method, and lpe
lwd is with size, canopy, and lpe

## Global Model
### Covariate Selection
Canopy Cover was deemed to be less helpful due to sudden changes by fire or
other year-to-year variation and that it wouldn't be that influential on
observer efficiency.

Size had the largest number of > 0.6 and only > 0.8 - remove.

stability was correlated with clarity and canopy. Plus 
the issue with recovering clarity in historical records makes it a little moot 
to try. Putting it back in because we can make an effort to recover/give a
good guess at what the clarity may be.

Substrate Shade was good.

LWD was .7 correlated with size, removing.
lpe_method was also correlated with size, but making a SME decision to keep it in.
Keeping it to test the global model.
```{r}
m.global <- glm.nb(hpe ~ size*lnlpe.sc + system_stability*lnlpe.sc + water_clarity*lnlpe.sc + 
                           substrate_shade*lnlpe.sc + #large_woody_debris + 
                           lpe_method*lnlpe.sc, 
                   data=cal2,
                     na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                     link="log")

summary(m.global)

plot(log(fitted(m.global)), residuals(m.global, type = "pearson"))
plot(log(fitted(m.global)), sqrt(abs(residuals(m.global, type = "pearson"))))

m.global.no.interaction <- glm.nb(hpe ~ size + system_stability + water_clarity + 
                           substrate_shade + #large_woody_debris + 
                           lpe_method + lnlpe.sc, 
                           data=cal2,
                     na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                     link="log")


summary(m.global.no.interaction)

plot(log(fitted(m.global.no.interaction)), residuals(m.global.no.interaction, type = "pearson"))
plot(log(fitted(m.global.no.interaction)), sqrt(abs(residuals(m.global.no.interaction, type = "pearson"))))
```

#### Variance Inflation Factors
general definition:
https://online.stat.psu.edu/stat462/node/180/
on interactions and categorical2 variables:
https://statistical2horizons.com/multicollinearity/
```{r}
vif(m.global) # with interaction 
vif(m.global.no.interaction) # without interaction
```

Summary:
lnlpe.sc has the highest VIF with interaction, but because it is directly interacting
with size.
Generally, categorical variables may also have high VIFs due to their nature.

Keeping with the interpretation without the interaction term.

Dropping Size, Canopy Cover, and LWD.
Considered dropping Water_Clarity due to it not being 100% retrievable but keeping it in.
Stability should be put back in

## Reduced Model
```{r size model with change in intercept}
m.cal2.reduced <- glm.nb(hpe ~ lnlpe.sc + water_clarity + system_stability +
                           substrate_shade + lpe_method +
                        water_clarity*lnlpe.sc + system_stability*lnlpe.sc +
                             substrate_shade*lnlpe.sc + lpe_method*lnlpe.sc, 
                   na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                   data=cal2,  
                   link="log")

plot(log(fitted(m.cal2.reduced)), residuals(m.cal2.reduced, type = "pearson"))
plot(log(fitted(m.cal2.reduced)), sqrt(abs(residuals(m.cal2.reduced, type = "pearson"))))
# possible outliers
cal2$usid[which(abs(residuals(m.cal2.reduced, type = "pearson"))>3)]
# Influential Data
plot(cooks.distance(m.cal2.reduced))
abline(h = 4/nrow(cal2), col = "blue") # Identifies points with High Influence
abline(h = 3*mean(cooks.distance(m.cal2.reduced)), col = "red")  # Identifies points with High Influence
# which system-years are possible issues or outliers?
# high leverage
cal2$usid[which(cooks.distance(m.cal2.reduced) > 3*mean(cooks.distance(m.cal2.reduced)))]

summary(m.cal2.reduced)
```

#### Comparing Size Models
```{r comparing global models}
# these are the TMB models
arrange(MuMIn::AICc(m.global, m.global.no.interaction, m.cal2.reduced), AICc) %>% 
  mutate(delta = AICc - min(AICc))

# parsimony is best fit.
# is within 2 AIC of the global, but most parsimonious.
```

```{r}
m.dredge <- dredge(m.cal2.reduced, beta = "none", rank = "AICc")
m.dredge

m.dredge <- as.data.frame(m.dredge %>% filter(delta < 4)) %>% 
  dplyr::select(lnlpe.sc:delta, -df, -logLik) %>% 
  mutate(AICc = round2(AICc, 2),
         delta = round2(delta, 2))

write.csv(m.dredge, file = "./tables/01_04-SizeAggregate-SM.csv", na = "", row.names = F)
```

### Top Model
```{r size model with change in intercept}
m.cal2.candidate <- glm.nb(hpe ~ lnlpe.sc + water_clarity + 
                           substrate_shade + lpe_method +
                        water_clarity*lnlpe.sc +
                             substrate_shade*lnlpe.sc + lpe_method*lnlpe.sc, 
                   na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                   data=cal2,  
                   link="log")

# Pearson residual is more appropriate for a glm that doesn't assume gaussian error distribution
png(file = "./figures/01_Model_Diagnostics/01-04-SizeAggregate-SM-Residuals.png",
    height = 5, width = 5, units = "in", res = 96)
plot(log(fitted(m.cal2.candidate)), residuals(m.cal2.candidate, type = "pearson"),
     xlab = "ln(Fitted Values)", ylab = "Pearson Residuals",
     main = "Size Aggregate (2) - Small-Medium Model")
abline(h = 0)
dev.off()

# possible outliers
cal2$usid[which(abs(residuals(m.cal2.candidate, type = "pearson"))>3)]
# Influential Data
plot(cooks.distance(m.cal2.candidate))
abline(h = 4/nrow(cal2), col = "blue") # Identifies points with High Influence
abline(h = 3*mean(cooks.distance(m.cal2.candidate)), col = "red")  # Identifies points with High Influence
# which system-years are possible issues or outliers?
# high leverage
cal2$usid[which(cooks.distance(m.cal2.candidate) > 3*mean(cooks.distance(m.cal2.candidate)))]
cal2$hpe[which(cooks.distance(m.cal2.candidate) > 3*mean(cooks.distance(m.cal2.candidate)))]

m.coef <- as.data.frame(summary(m.cal2.candidate)$coefficients)
m.coef$parameter = row.names(m.coef)
colnames(m.coef)[2] = "std.error"

ggplot(filter(m.coef, parameter != "(Intercept)"),
       aes(x = reorder_within(parameter, Estimate,
                              parameter, mean), 
           y = Estimate)) +
  geom_point() +
  scale_x_reordered() +
  geom_errorbar(aes(ymin = Estimate-std.error, ymax = Estimate+std.error),
                width = 0.5) +
  geom_hline(yintercept = 0) +
  coord_flip() + 
  labs(x = "Parameter")

ggsave("./figures/01_Model_Diagnostics/01-04-SizeAggregate-SM-Parameters.png",
       width = 5, height = 5, dpi = 96)

m.coef <- dplyr::select(m.coef, parameter, Estimate, std.error) %>% 
  arrange(desc(abs(Estimate))) %>% 
  mutate(cross.zero = ifelse(abs(Estimate) < abs(std.error), "Yes", "No"),
         Estimate = round2(Estimate, 2),
         std.error = round2(std.error, 2))
write.csv(m.coef, file = "./tables/01_04-SizeAggregate-SM-Parameters.csv",
          row.names = F)
```

## Leave One Out
```{r}
model.test <- data.frame(model=character(),
                 usid=character(), 
                 fit=numeric(), 
                 se.fit=numeric(),
                 calibrated.estimate=numeric()) 
```

```{r}
saveRDS(m.cal2.candidate, file = "./r_objects/models/SM_Data_Model.rds")

specifications_out <- data.frame(model = c("SM_Data_Model"),
                                 target = c("lpe","lpe","lpe"),
                                 order = c(1,2,3),
                                 type = c("log","center","scale"),
                                 transform = c(NA, mean(cal2$lnlpe), sd(cal2$lnlpe))) # note that in the future, may want to re-center and scale using the SM Data instead of the whole dataset

spec.out <- spec.out %>% 
  filter(model != "SM_Data_Model") %>% 
  rbind(., specifications_out)

saveRDS(spec.out, file=modout)
```
Need to add a running of the model without the one data point.
```{r}
model = m.cal2.candidate
  for(i in 1:nrow(cal2)){
    m.temp <- glm.nb(as.character(model$call)[2],
                     data=cal2[-i, ],
                     na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                     link="log")
    p.temp <- predict(m.temp, type="link", cal2[i, ], se.fit=T)
    model.test <- rbind(model.test, data.frame(model = as.character(model$call)[2],
                                 usid = cal2$usid[i],
                                 fit = p.temp$fit,
                                 se.fit = p.temp$se.fit,
                        calibrated.estimate = exp(p.temp$fit)))
}
```

```{r}
model.rank <- left_join(model.test, 
                        dplyr::select(cal2,
                                      usid, hpe, watershed_group:year, lpe_method)) %>% 
  mutate(error = calibrated.estimate - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error),
         rank.set.level = 2)

```
### Plots
```{r box and whisker of percent error by system}
model.rank %>% 
  ggplot(aes(x = reorder(gazetted_stream_name, abs.per.error, FUN = median), 
             # reorder FUN = median, reorders based on the median
             y = abs.per.error)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_point(position = "jitter", alpha = 0.5,
             aes(colour = lpe_method)) +
  geom_hline(yintercept = 0.5, colour = "grey") +
  coord_flip() +
  labs(title = "Data excluding Large+ Model Performance",
       y = "Absolute percent error", x = "Streams",
       colour = "Survey Method")
```

## Save Model Ranks
```{r}
saveRDS(model.rank, "r_objects/01_Model_Ranks_from_SM_Data_Model.rds")
```

# 4
# Analysis - L Only
```{r}
cal %>%  filter(size %in% c("X-Large", "Large")) %>% 
  ggplot(aes(x = lnlpe.sc, y = log(hpe), colour = gazetted_stream_name)) +
  geom_point() +
  labs(x = "ln(Scaled - Low Precision Estimate)", 
       y = "ln(High Precision Estimate)") 

cal %>%  filter(size %in% c("X-Large", "Large")) %>% 
  count(lpe_method, gazetted_stream_name)
```
While reviewing Harrison data more closely, it is notable that the ground
surveys are not the entire river. The counts are also much lower than the
aerial counts, to a degree that even a "ground" correction might not save them.
Particularly for making an XL-only relationship, the only other ground was 1
Little River survey, which was nearly identical to the aerial count done at the
time.

Removed Harrison Ground Surveys in the original database.
```{r}
cal3 <- cal %>% 
  filter(size %in% c("X-Large", "Large")) %>% 
  # using cloud_cover, so recoding it %>% 
  mutate(cloud_cover_recode = case_when( # changing cover to a number
      cloud_cover=="0"~1, 
      cloud_cover=="25"~2, 
      cloud_cover=="50"~3,
      cloud_cover=="75"~4,
      cloud_cover=="100"~5),
      lnlpe.sc = (lnlpe - mean(lnlpe))/sd(lnlpe))
```

```{r}
cal3 %>%  
  ggplot(aes(x = lnlpe.sc, y = log(hpe), colour = water_clarity)) +
  geom_point() +
  labs(x = "ln(Scaled - Low Precision Estimate)", 
       y = "ln(High Precision Estimate)") +
  geom_smooth(method = "lm")

summary(cal3)
```
Only 2 with different system_stability, remove.
canopy only has 2 samples in High.
brightness has 3 NAs
cloud_cover has 3 NAs
fish_vis has 2 NAs

## Correlation Coefficients
```{r correlation figure}
cal3.cor <- round2(cor(dplyr::select(cal3, size_recode, water_clarity_recode, 
                                    substrate_shade_recode, 
                                    lpe_method_recode, lnlpe.sc)), 3)
cal3.cor

chart.Correlation(dplyr::select(cal3, size_recode, water_clarity_recode, 
                                substrate_shade_recode, 
                                lpe_method_recode, lnlpe.sc), 
                  histogram = TRUE, method = "pearson")
```
Note that water clarity is mostly static within a system but there are some 
Harrison surveys where it is not.

size and method are correlated. Otherwise pretty ok.

## Global Model
### Covariate Selection
Will keep both Water Clarity and Substrate Shade, but expect to drop one of them
```{r}
m.global <- glm.nb(hpe ~ size*lnlpe.sc + #system_stability + 
                     water_clarity*lnlpe.sc + 
                           substrate_shade*lnlpe.sc + #large_woody_debris + 
                           lpe_method*lnlpe.sc, 
                   data=cal3,
                     na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                     link="log")

summary(m.global)

plot(log(fitted(m.global)), residuals(m.global, type = "pearson"))

m.global.no.interaction <- glm.nb(hpe ~ size + #system_stability + 
                                    water_clarity + 
                           substrate_shade + #large_woody_debris + 
                           lpe_method + lnlpe.sc, 
                           data=cal3,
                     na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                     link="log")


summary(m.global.no.interaction)

plot(log(fitted(m.global.no.interaction)), residuals(m.global.no.interaction, type = "pearson"))
plot(log(fitted(m.global.no.interaction)), sqrt(abs(residuals(m.global.no.interaction, type = "pearson"))))
```

#### Variance Inflation Factors
general definition:
https://online.stat.psu.edu/stat462/node/180/
```{r}
vif(m.global)
vif(m.global.no.interaction)
```

Summary:
All have relatively low VIF, so keeping them all.

```{r}
m.dredge <- dredge(m.global, beta = "none", rank = "AICc")
m.dredge

m.dredge <- as.data.frame(m.dredge %>% filter(delta < 3)) %>% 
  dplyr::select(lnlpe.sc:delta, -df, -logLik) %>% 
  mutate(AICc = round2(AICc, 2),
         delta = round2(delta, 2))

write.csv(m.dredge, file = "./tables/01_04-SizeAggregate-LXL.csv", na = "", row.names = F)
```
Top 2 models both have lpe and one of either size or water clarity.
3rd model has all three, lpe, size, and clarity.

No interaction terms nor method are included within < 2 AICc.

Several models within 4 AIC.

Going with the model that has size and water clarity.
Not most Parsimonious, but would be most similar to a model averaged of < 2 AICc.

## Top Model
```{r}
m.L.top <- glm.nb(hpe ~ lnlpe.sc + water_clarity + size, 
    data=cal3,  
    na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
    link="log")

summary(m.L.top)

# Pearson residual is more appropriate for a glm that doesn't assume gaussian error distribution
png(file = "./figures/01_Model_Diagnostics/01-04-SizeAggregate-LXL-Residuals.png",
    height = 5, width = 5, units = "in", res = 96)
plot(log(fitted(m.L.top)), residuals(m.L.top, type = "pearson"),
     xlab = "ln(Fitted Values)", ylab = "Pearson Residuals",
     main = "Size Aggregate (2) - Large-X-Large Model")
abline(h = 0)
dev.off()

# possible outliers
cal3$usid[which(abs(residuals(m.L.top, type = "pearson"))>1.4)]
# Influential Data
plot(cooks.distance(m.L.top))
abline(h = 4/nrow(cal3), col = "blue") # Identifies points with High Influence
abline(h = 3*mean(cooks.distance(m.L.top)), col = "red")  # Identifies points with High Influence
# which system-years are possible issues or outliers?
# high leverage
cal3$usid[which(cooks.distance(m.L.top) > 3*mean(cooks.distance(m.L.top)))]
```

Bad Residual pattern. Reject model.

## Aerial Only
```{r}
cal4 <- cal3 %>% 
  filter(lpe_method == "Aerial") %>% 
  mutate(lnlpe.sc = (lnlpe - mean(lnlpe))/sd(lnlpe))

cal4 %>% count(water_clarity, size)
cal3 %>% count(water_clarity, size)
```

```{r}
m.LA.global <- glm.nb(hpe ~ lnlpe.sc + water_clarity*lnlpe.sc + size*lnlpe.sc, 
    data=cal4,  
    na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
    link="log")

summary(m.LA.global)

# Pearson residual is more appropriate for a glm that doesn't assume gaussian error distribution
plot(log(fitted(m.LA.global)), residuals(m.LA.global, type = "pearson"),
     xlab = "ln(Fitted Values)", ylab = "Pearson Residuals",
     main = "Size Aggregate (2) - LXL Model")
abline(h = 0)
```

```{r}
m.dredge <- dredge(m.LA.global, beta = "none", rank = "AICc")
m.dredge

m.dredge <- as.data.frame(m.dredge %>% filter(delta < 3.01)) %>% 
  dplyr::select(lnlpe.sc:delta, -df, -logLik) %>% 
  mutate(AICc = round2(AICc, 2),
         delta = round2(delta, 2))

write.csv(m.dredge, file = "./tables/01_04-SizeAggregate-LXL-Aerial.csv", na = "", row.names = F)
```
### Top Model
```{r}
m.LA.top <- glm.nb(hpe ~ lnlpe.sc + size*lnlpe.sc, 
    data=cal4,  
    na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
    link="log")

summary(m.LA.top)

# Pearson residual is more appropriate for a glm that doesn't assume gaussian error distribution
png(file = "./figures/01_Model_Diagnostics/01-04-SizeAggregate-LXL-Aerial-Residuals.png",
    height = 5, width = 5, units = "in", res = 96)
plot(log(fitted(m.LA.top)), residuals(m.LA.top, type = "pearson"),
     xlab = "ln(Fitted Values)", ylab = "Pearson Residuals",
     main = "Size Aggregate (2) - LXL Model")
abline(h = 0)
dev.off()

# possible outliers
cal3$usid[which(abs(residuals(m.LA.top, type = "pearson"))>3)]
# Influential Data
plot(cooks.distance(m.LA.top))
abline(h = 4/nrow(cal3), col = "blue") # Identifies points with High Influence
abline(h = 3*mean(cooks.distance(m.LA.top)), col = "red")  # Identifies points with High Influence
# which system-years are possible issues or outliers?
# high leverage
cal3$usid[which(cooks.distance(m.LA.top) > 3*mean(cooks.distance(m.LA.top)))]

m.coef <- as.data.frame(summary(m.LA.top)$coefficients)
m.coef$parameter = row.names(m.coef)
colnames(m.coef)[2] = "std.error"

ggplot(filter(m.coef, parameter != "(Intercept)"),
       aes(x = reorder_within(parameter, Estimate,
                              parameter, mean), 
           y = Estimate)) +
  geom_point() +
  scale_x_reordered() +
  geom_errorbar(aes(ymin = Estimate-std.error, ymax = Estimate+std.error),
                width = 0.5) +
  geom_hline(yintercept = 0) +
  coord_flip()

ggsave("./figures/01_Model_Diagnostics/01-04-SizeAggregate-LXL-Aerial-Parameters.png",
       width = 5, height = 5, dpi = 96)

m.coef <- dplyr::select(m.coef, parameter, Estimate, std.error) %>% 
  arrange(desc(abs(Estimate))) %>% 
  mutate(cross.zero = ifelse(abs(Estimate) < abs(std.error), "Yes", "No"),
         Estimate = round2(Estimate, 2),
         std.error = round2(std.error, 2))
write.csv(m.coef, file = "./tables/01_04-SizeAggregate-LXL-Aerial-Parameters.csv",
          row.names = F)
```

## Leave One Out
```{r}
model.test <- data.frame(model=character(),
                 usid=character(), 
                 fit=numeric(), 
                 se.fit=numeric(),
                 calibrated.estimate=numeric()) 
```

```{r}
saveRDS(m.LA.top, file = "./r_objects/models/LXL_Data_Model.rds")

specifications_out <- data.frame(model = c("LXL_Data_Model"),
                                 target = c("lpe","lpe","lpe"),
                                 order = c(1,2,3),
                                 type = c("log","center","scale"),
                                 transform = c(NA, mean(cal4$lnlpe), sd(cal4$lnlpe))) # note that in the future, may want to re-center and scale using the XLXL Data instead of the whole dataset

spec.out <- spec.out %>% 
  filter(model != "LXL_Data_Model") %>% 
  rbind(., specifications_out)

saveRDS(spec.out, file=modout)
```
Need to add a running of the model without the one data point.
```{r}
model = m.LA.top
for(i in 1:nrow(cal4)){
    m.temp <- glm.nb(as.character(model$call)[2],
                     data=cal4[-i, ],
                     na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                     link="log")
    p.temp <- predict(m.temp, type="link", cal4[i, ], se.fit=T)
    model.test <- rbind(model.test, data.frame(model = as.character(model$call)[2],
                                 usid = cal4$usid[i],
                                 fit = p.temp$fit,
                                 se.fit = p.temp$se.fit,
                        calibrated.estimate = exp(p.temp$fit)))
}
```

```{r}
model.rank <- left_join(model.test, 
                        dplyr::select(cal4,
                                      usid, hpe, watershed_group:year, lpe_method)) %>% 
  mutate(error = calibrated.estimate - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error),
         rank.set.level = 2)

```

# Save Model Ranks
```{r}
saveRDS(model.rank, "r_objects/01_Model_Ranks_from_LXL_Data_Model.rds")
```

# 5
# Analysis - System Type - Specific
```{r}
(cal.fil.5 <- count(cal, gazetted_stream_name, lpe_method) %>% 
  filter(n>5) %>% 
  mutate(filter5 = paste0(gazetted_stream_name, "_", lpe_method)))

cal5 <- cal %>% 
  mutate(filter5 = paste0(gazetted_stream_name, "_", lpe_method)) %>% 
  filter(filter5 %in% cal.fil.5$filter5)

cal5 %>% 
  ggplot(aes(x = lnlpe, y = log(hpe),
             colour = filter5)) +
  geom_point()
```

## Individual Model
```{r}
p.list <- list() # list for plots

for(k in 1:nrow(cal.fil.5)){
  print(cal.fil.5$filter5[k])
  # make temp data with just one system
  cal5.temp <- filter(cal5, filter5 == cal.fil.5$filter5[k])
  # make temp model with just one system
  m.temp <- glm.nb(hpe ~ lnlpe, 
    data=cal5.temp,  
    na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
    link="log")
  
  print("Outliers")
  print(cal5.temp$usid[which(abs(residuals(m.temp, type = "pearson"))>3)])
  print("High-leverage")
  print(cal5.temp$usid[which(cooks.distance(m.temp) > 3*mean(cooks.distance(m.temp)))])
  
        # Pearson residual is more appropriate for a glm that doesn't assume gaussian error distribution
      png(file = paste0("./figures/01_Model_Diagnostics/01-04-Individual-", 
                        cal5.temp$gazetted_stream_name[1], "-", 
                        cal5.temp$lpe_method[1],"-Residuals.png"),
    height = 5, width = 5, units = "in", res = 96)
     plot(log(fitted(m.temp)), residuals(m.temp, type = "pearson"),
           xlab = "ln(Fitted Values)", ylab = "Pearson Residuals",
           main = paste0("Individual (5) ", cal5.temp$gazetted_stream_name[1], 
                         "-", cal5.temp$lpe_method[1], " Model"))
      abline(h = 0)
      dev.off()
  
      p.temp <- broom::augment(m.temp)
      p.list[[k]] <- ggplot(p.temp, aes(x = .fitted, y = .std.resid)) + 
        geom_point(pch = 1, size = 2) +
        geom_hline(yintercept = 0) +
        labs(x = "ln(Fitted Values)", y = "Pearson Residuals",
           title = paste0(cal5.temp$gazetted_stream_name[1], 
                         "-", cal5.temp$lpe_method[1])) + 
        theme(plot.title = element_text(size=11))
      
  # model test dataframe
  model.test <- data.frame(model=character(),
                 usid=character(), 
                 fit=numeric(), 
                 se.fit=numeric(),
                 calibrated.estimate=numeric())

  saveRDS(m.temp, file = paste0("./r_objects/models/", cal.fil.5$filter5[k], "-Individual_Data_Model.rds"))
  
    model = m.temp
    
    for(i in 1:nrow(cal5.temp)){
      m.temp <- glm.nb(as.character(model$call)[2],
                     data=cal5.temp[-i, ],
                     na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                     link="log")
      
      p.temp <- predict(m.temp, type="link", cal5.temp[i, ], se.fit=T)
      model.test <- rbind(model.test, data.frame(model = as.character(model$call)[2],
                                 usid = cal5.temp$usid[i],
                                 fit = p.temp$fit,
                                 se.fit = p.temp$se.fit,
                        calibrated.estimate = exp(p.temp$fit)))
  }
  model.rank.temp <- left_join(model.test, 
                        dplyr::select(cal5.temp,
                                      usid, hpe, watershed_group:year, lpe_method)) %>% 
    mutate(error = calibrated.estimate - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error),
         rank.set.level = 5)
  
  if(k == 1) model.rank = model.rank.temp
  if(k != 1) model.rank = rbind(model.rank, model.rank.temp)
}

wrap_plots(p.list, ncol = 3) +
  plot_layout(guides = "collect",
              axes = "collect")

ggsave("./figures/01_Model_Diagnostics/01-04-Individual-Data_Model-ALL.png",
       height = 9, width = 8, dpi = 96)
```

# Save Model Ranks
```{r}
saveRDS(model.rank, "r_objects/01_Model_Ranks_from_Individual_Data_Model.rds")
```

```{r}
specifications_out <- data.frame(model = c("Individual_Data_Model"),
                                 target = c("lpe"),
                                 order = c(1),
                                 type = c("log"),
                                 transform = c(NA)) 

spec.out <- spec.out %>% 
  filter(model != "Individual_Data_Model") %>% 
  rbind(., specifications_out)

saveRDS(spec.out, file=modout)
```

```{r}

```

