---
title: "01-06-Discharge-Specific"
author: "Michael Arbeider"
date: "2026-03-06"
---

# Set-up
## Packages
https://www.rdocumentation.org/packages/tidyhydat/versions/0.6.1/topics/hy_daily_flows
```{r}
rm(list = ls())
library(tidyverse) # most %>% 
theme_set(theme_bw()) # ggplot
library(MASS) # boxcox - masked from ‘package:dplyr’: select()
library(lme4) # generalized linear regressions
# methods(class = "lmerMod") # shows what functionality is available
library(MuMIn) # dredge
library(PerformanceAnalytics) # correlation figure
library(glmmTMB) # generalized linear regressions
library(car)       # for vif()
library(patchwork) # for multi plot ggplot graphs
library(scales) # for label_comma()
library(tidyhydat) # hy_daily for hydrometric data
source("00-Functions.R") # scale_x_reordered, rank.summary(), round2()
```

## Data
```{r}
modout <- "./r_objects/models/model_specifications.rds" # for saving specifications later
spec.out <- readRDS(modout)
cal <-readRDS("r_objects/sockeye_calibration_data.rds")   

glimpse(cal)

cal <- cal %>% filter(!is.na(water_discharge))
# considering removing systems with less than 5 to be able to use a random effect of system
count(cal, gazetted_stream_name)
count(cal, lpe_method) # stored as a character so the summary didn't catch the count between levels
summary(cal)
```

Variables with > 5 n per level:
system_stability
size (if Small is rolled into Medium)
water_clarity (if Turbid is rolled into Pt. Turbid/Tannic)
substrate_shade
canopy_cover (excluding High which has 0)
large_woody_debris (if High is treated as Medium)
lpe_method


### Plot by System
```{r}
cal %>% 
  ggplot(aes(x = lnlpe.sc, y = log(hpe), 
             colour = gazetted_stream_name,
             shape = size)) +
  geom_point() +
  labs(x = "ln(Scaled - Low Precision Estimate)", 
       y = "ln(High Precision Estimate)")
```

The smallest observation may act as a high leverage point.

# Stellako Specific
```{r}
stel <- cal %>% filter(gazetted_stream_name == "Stellako River")
summary(stel)
count(stel, lpe_method)
# most of these are paired counts between aerial and ground.

stel %>% 
  ggplot(aes(x = lnlpe.sc, y = log(hpe), 
             shape = lpe_method)) +
  geom_point() +
  labs(x = "ln(Scaled - Low Precision Estimate)", 
       y = "ln(High Precision Estimate)")
```

Variables with > 5 n per level (and no NAs):
water_discharge

## Test for Normality
High precision estimates
```{r}
hist(stel$hpe)
shapiro.test(stel$hpe) # p < 0.01-0.05 means that it is NOT normal
b <- boxcox(lm(stel$hpe ~ 1))
b$x[which.max(b$y)] # lambda
hist(log(stel$hpe))
shapiro.test(log(stel$hpe)) # worse
hist(sqrt(stel$hpe))
shapiro.test(sqrt(stel$hpe)) # not as good as untransformed but normal
hist((stel$hpe)^(1/3))
shapiro.test((stel$hpe)^(1/3)) # worst
```
low precision estimates
```{r}  
hist(stel$lpe)
shapiro.test(stel$lpe) # p < 0.01-0.05 means that it is NOT normal
b <- boxcox(lm(stel$lpe ~ 1))
b$x[which.max(b$y)] # lambda
hist(log(stel$lpe))
shapiro.test(log(stel$lpe)) # not very acceptable
hist(sqrt(stel$lpe))
shapiro.test(sqrt(stel$lpe)) # bestish
hist((stel$lpe)^(1/3))
shapiro.test((stel$lpe)^(1/3)) # ok
```

### Transformation Plots
LPE and HPE
```{r}
p.st1 <- stel %>% 
  ggplot(aes(x = lpe, y = hpe,
             colour = water_discharge)) +
  geom_point()

p.st2 <- stel %>% 
  ggplot(aes(x = log(lpe), y = log(hpe),
             colour = water_discharge)) +
  geom_point()

p.st3 <- stel %>% 
  ggplot(aes(x = sqrt(lpe), y = hpe,
             colour = water_discharge)) +
  geom_point()

p.st4 <- stel %>% 
  ggplot(aes(x = sqrt(lpe), y = log(hpe),
             colour = water_discharge)) +
  geom_point()

(p.st1 | p.st2) / (p.st3 | p.st4) + plot_layout(guides = 'collect')
```
None look like they would necessarily make the distribution of data "better"
since the range of data is relatively small compared to the larger data sets.

May be "simplest" to use negative binomial and no transformations on hpe as also suggested from shapiro.
Transform lpe with a sqrt as suggested from shapiro.

Later determined that using the identity link results in negative error bars.
So using the log-link.

## Center and Scale
Scaling flow by system-specific and time period specific recent average.
```{r}
stel.flow <- hy_daily_flows(station_number = "08JB002")
stel.flow <- stel.flow %>% 
  mutate(Year = format(as.Date(Date, format="%Y/%m/%d"),"%Y"),
         JDay = lubridate::yday(Date)) %>% 
  group_by(Year) %>% 
  mutate(records = n()) %>% 
  ungroup() %>% 
  mutate(complete = records >= 364) %>% 
  filter(complete == TRUE)

stel.flow.summary <- stel.flow %>% 
  filter(Year >= 2024-20) %>% 
  group_by(JDay) %>% 
  summarise(average = mean(Value),
            p10 = quantile(Value, 0.10),
            p25 = quantile(Value, 0.25),
            p50 = quantile(Value, 0.50),
            p75 = quantile(Value, 0.75),
            p90 = quantile(Value, 0.90),
            max = max(Value),
            min = min(Value))

ggplot(stel.flow.summary,
       aes(x = JDay, y = average)) +
  geom_line() +
  geom_line(aes(y = p10), col = "purple") +
  geom_line(aes(y = p90), col = "purple") +
  geom_line(aes(y = min), col = "red") +
  geom_line(aes(y = max), col = "red") +
  geom_vline(xintercept = c(245, 281)) + # Sep 1 - Oct 7 - range of Summer Peak Spawn dates from 2023 Prelim (excluding Harrison and Widgeon)
  geom_point(data = stel,
             aes(x = lubridate::yday(lpe_date),
                 y = water_discharge),
             colour = "blue",
             alpha = 0.7)

stel.mean = mean(stel.flow.summary$average[stel.flow.summary$JDay %in% c(245:281)])
stel.sd = sd(stel.flow.summary$average[stel.flow.summary$JDay %in% c(245:281)])
mean(stel$water_discharge)
sd(stel$water_discharge)

stel <- stel %>% 
  mutate(root.lpe = sqrt(lpe),
         root.lpe.sc = scale(root.lpe, center=T, scale=T)[,1],
         discharge.sc = (water_discharge-stel.mean)/stel.sd,
         lnlpe.sc = scale(lnlpe, center = T, scale = T)[,1])

hist(stel$discharge.sc)
```


## Correlation Coefficients
```{r correlation figure}
summary(stel)

stel.cor <- round(cor(dplyr::select(stel,
                 lpe_method_recode, root.lpe.sc, lnlpe.sc,
                 discharge.sc)), 3)
stel.cor

#png(file="figures/Discharge - Stellako - calibration Correlation Plot.png",
#    units = "in", width=10, height=5, res = 250)

chart.Correlation(dplyr::select(stel, 
                 lpe_method_recode, root.lpe.sc, lnlpe.sc,
                 discharge.sc), 
                  histogram = TRUE, method = "pearson")
#dev.off()
```
root.lpe is a little correlated with discharge.

## Model
### Compare lnlpe vs root.lpe
Above, just based on distribution, it may have been best to select square-root
transformed lpe, but I wanted to test it in the model framework as well.
```{r}
m.stel.root <- glm.nb(hpe ~ root.lpe.sc, data = stel, 
                      na.action = "na.fail", link = "log")
m.stel.ln <- glm.nb(hpe ~ lnlpe.sc, data = stel, 
                      na.action = "na.fail", link = "log")

arrange(MuMIn::AICc(m.stel.root, m.stel.ln), AICc) %>% 
  mutate(delta = AICc - min(AICc))
```
using the natural-log is better by >4 AICc.

### Testing Quadratic as Global Model
Notes with regards to quadratics
https://stats.stackexchange.com/questions/379841/in-linear-regression-why-should-we-include-quadratic-terms-when-we-are-only-int
https://stats.stackexchange.com/questions/11009/including-the-interaction-but-not-the-main-effects-in-a-model
https://quantifyinghealth.com/quadratic-term-in-regression/

```{r}
m.stel <- glm.nb(hpe ~ lnlpe.sc:lpe_method + lpe_method:I(lnlpe.sc^2) +
                           lnlpe.sc:discharge.sc + discharge.sc:I(lnlpe.sc^2) +
                   lnlpe.sc + I(lnlpe.sc^2) + lpe_method + discharge.sc, 
    data=stel, 
    na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
    link = "log") 

summary(m.stel)
# Pearson residual is more appropriate for a glm that doesn't assume Gaussian error distribution
plot((fitted(m.stel)), residuals(m.stel, type = "pearson"))

m.stel.coef <- as.data.frame(summary(m.stel)$coefficients)
m.stel.coef$parameter = row.names(m.stel.coef)
colnames(m.stel.coef)[2] = "std.error"

ggplot(m.stel.coef[-1,],
       aes(x = parameter, y = Estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = Estimate-std.error, ymax = Estimate+std.error)) +
  geom_hline(yintercept = 0) +
  coord_flip()
```

#### Dredge
```{r}
m.dredge <- dredge(m.stel, beta = "sd", rank = "AICc")
m.dredge

m.dredge <- as.data.frame(m.dredge %>% filter(delta < 4)) %>% 
  dplyr::select(discharge.sc:delta)

write.csv(m.dredge, file = "./tables/01_06-Stellako-Discharge.csv", na = "", row.names = F)
```

3 witin 4 AICc.
All contain discharge.sc, lnl.sc, lnl.sc^2 and 
the interaction between discharge and lpe^2

Most Parsimonious is also the top.
2 out of the other four contain lpe_method. One contains the interaction 
between discharge and lpe and one contains interaction with lpe_method and lpe^2.

##### Top
```{r}
m.stel.top <- glm.nb(hpe ~ lnlpe.sc + discharge.sc + I(lnlpe.sc^2) +
                       discharge.sc:I(lnlpe.sc^2), 
    data=stel, 
    na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
    link="log")

summary(m.stel.top)
# Pearson residual is more appropriate for a glm that doesn't assume Gaussian error distribution
png(file = "./figures/01_Model_Diagnostics/01-06-Stellako-Discharge-Residuals.png",
    height = 5, width = 5, units = "in", res = 96)
plot(log(fitted(m.stel.top)), residuals(m.stel.top, type = "pearson"),
     xlab = "ln(Fitted Values)", ylab = "Pearson Residuals",
     main = "Specialized (6) - Stellako Discharge Model")
abline(h = 0)
dev.off()

# possible outliers
stel$usid[which(abs(residuals(m.stel.top, type = "pearson"))>3)]
# Influential Data
plot(cooks.distance(m.stel.top))
abline(h = 4/nrow(stel), col = "blue") # Identifies points with High Influence
abline(h = 3*mean(cooks.distance(m.stel.top)), col = "red")  # Identifies points with High Influence
# which system-years are possible issues or outliers?
# high leverage
stel$usid[which(cooks.distance(m.stel.top) > 3*mean(cooks.distance(m.stel.top)))]

m.stel.top.coef <- as.data.frame(summary(m.stel.top)$coefficients)
m.stel.top.coef$parameter = row.names(m.stel.top.coef)
colnames(m.stel.top.coef)[2] = "std.error"

ggplot(m.stel.top.coef[-1,],
       aes(x = reorder_within(parameter, Estimate,
                              parameter, mean), 
           y = Estimate)) +
  geom_point() +
  scale_x_reordered() +
  geom_errorbar(aes(ymin = Estimate-std.error, ymax = Estimate+std.error)) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  labs(x = "Parameter")

ggsave("./figures/01_Model_Diagnostics/01-06-Stellako-Discharge-Parameters.png",
       width = 5, height = 5, dpi = 96)

m.coef <- dplyr::select(m.stel.top.coef, parameter, Estimate, std.error) %>% 
  arrange(desc(abs(Estimate))) %>% 
  mutate(cross.zero = ifelse(abs(Estimate) < abs(std.error), "Yes", "No"))
write.csv(m.coef, file = "./tables/01_06-Steallko-Discharge-Parameters.csv",
          row.names = F)
```

Double check
```{r}
arrange(MuMIn::AICc(m.stel, m.stel.top, m.stel.ln), AICc) %>% 
  mutate(delta = AICc - min(AICc))
```

## Leave One Out
```{r}
model.test <- data.frame(model=character(),
                 usid=character(), 
                 fit=numeric(), 
                 se.fit=numeric(),
                 calibrated.estimate=numeric()) 
```

```{r}
models <- list(m.stel.top)

saveRDS(m.stel.top, file = "./r_objects/models/Stellako_Discharge_Model.rds")

specifications_out <- data.frame(model = "Stellako_Discharge_Model",
                                 target = c("lpe","lpe","lpe",
                                            "discharge","discharge"),
                                 order = c(1,2,3,
                                           1,2),
                                 type = c("log","center","scale",
                                          "center", "scale"),
                                 transform = c(NA, mean(stel$lnlpe), sd(stel$lnlpe),
                                               stel.mean, stel.sd))


spec.out <- spec.out %>% 
  filter(model != "Stellako_Discharge_Model") %>% 
  rbind(., specifications_out)

saveRDS(spec.out, file=modout)
```
Need to add a running of the model without the one data point.
```{r}
for(j in 1:length(models)){
  model = models[[j]]
  for(i in 1:nrow(stel)){
    m.temp <- glm.nb(as.character(model$call)[2],
                     data=stel[-i, ],
                     na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                     link="log")
    p.temp <- predict(m.temp, type="link", stel[i, ], se.fit=T)
    model.test <- rbind(model.test, data.frame(model = as.character(model$call)[2],
                                 usid = stel$usid[i],
                                 fit = p.temp$fit,
                                 se.fit = p.temp$se.fit,
                                 calibrated.estimate = exp(p.temp$fit)))
  }
}
```

```{r}
model.rank <- left_join(model.test, 
                        dplyr::select(stel,
                                      usid, hpe, watershed_group:year, lpe_method)) %>% 
  mutate(error = calibrated.estimate - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error),
         rank.set.level = 6)

```
### Plots
```{r plot of error}
model.rank %>% 
  ggplot(aes(x = hpe, y = error)) +
  geom_point() +
  scale_y_continuous(labels = label_comma())
```

```{r plot of percent error}
model.rank %>% 
  ggplot(aes(x = hpe, y = per.error)) +
  geom_point()
```

```{r plot of absolute percent error}
model.rank %>% 
  ggplot(aes(x = hpe, y = abs.per.error)) +
  geom_point()
```

```{r plot of absolute percent error by system}
set.seed(234)
model.rank %>% 
  ggplot(aes(x = reorder(gazetted_stream_name, abs.per.error, FUN = median), 
             # reorder FUN = median, reorders based on the median
             y = abs.per.error,
             colour = model)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_hline(yintercept = 0.5, colour = "grey") +
  geom_point(position = "jitter", alpha = 0.5) +
  coord_flip() +
  labs(title = "Stellako Discharge Model Performance",
       y = "Absolute percent error", x = "Streams",
       colour = "Model Types") +
  theme(legend.position = "top")

#ggsave("figures/Model Performance/03_Model_Performance_Stellako_Discharge.png",
#       height = 6, width = 8, dpi = 150)
```

```{r}
filter(model.rank,
       abs.per.error > 0.4)
```

## Save Model Ranks
```{r}
saveRDS(model.rank, 
        "r_objects/03_Model_Ranks_from_Stellako_Discharge_Model.rds")
```


# All Systems
```{r}
summary(cal)
count(cal, system_stability)
```

Variables with > 5 n per level (and no NAs):
water_discharge
lpe_method
system_stability
size only has 2 v.smalls, one of which is quite an outlier numerically (Dust 2002)
water_clarity
substrate_shade
canopy_cover
LWD only 1 high, won't include.

Removing Small System
```{r}
count(cal, size)

cal %>% filter(hpe == min(hpe))

cal <- cal %>% 
  filter(size != "Small")
```

Also discovered that Middle River's discharge is from an unknown location.
Removing it.

```{r}
cal <- cal %>% 
  filter(gazetted_stream_name != "Middle River")
```


## Test for Normality
High precision estimates
```{r}
hist(cal$hpe)
shapiro.test(cal$hpe) # p < 0.01-0.05 means that it is NOT normal
b <- boxcox(lm(cal$hpe ~ 1))
b$x[which.max(b$y)] # lambda
hist(log(cal$hpe))
shapiro.test(log(cal$hpe)) # good
hist(sqrt(cal$hpe))
shapiro.test(sqrt(cal$hpe)) # bad
hist((cal$hpe)^(1/3))
shapiro.test((cal$hpe)^(1/3)) # best but log looks fine, so will use that with negative binom.
```
low precision estimates
```{r}  
hist(cal$lpe)
shapiro.test(cal$lpe) # p < 0.01-0.05 means that it is NOT normal
b <- boxcox(lm(cal$lpe ~ 1))
b$x[which.max(b$y)] # lambda
hist(log(cal$lpe))
shapiro.test(log(cal$lpe)) # Good
hist(sqrt(cal$lpe))
shapiro.test(sqrt(cal$lpe)) # bad
hist((cal$lpe)^(1/3))
shapiro.test((cal$lpe)^(1/3)) # not good
```

Discharge should be scaled within a system. 
Thinking about it as, does the relative change in discharge within a system 
explain the residuals of the standard lpe vs hpe relationship?

Plan to centre by system specific recent average and scale by recent system deviation.
## Center and Scale
```{r}
cal <- cal %>% 
  mutate(ln.lpe = log(lpe),
         ln.lpe.sc = scale(ln.lpe, center=T, scale=T)[,1])

# environment canada stations where flow data comes from
ec.stations <- unique(dplyr::select(cal, ec_station_number, gazetted_stream_name, 
                        run_timing_group)) %>% 
  mutate(start = case_when(
    run_timing_group == "Summer" ~ 245, # Sep 1
    run_timing_group == "Early Summer" ~ 245, # Sep 1 (excludes Chilliwack and Upper Barriere)
    run_timing_group == "Late" ~ 275 # Oct 1
  ),
  end = case_when(
    run_timing_group == "Summer" ~ 281, # Oct 7
    run_timing_group == "Early Summer" ~ 272, # Sep 28 (excludes Chilliwack and Upper Barriere)
    run_timing_group == "Late" ~ 299 # Oct 25 just including Shuswap Complex
  ))

# found in comments that Harrison's Discharge is from station 08MG013, while the Water Level is from the one in the database
ec.stations$ec_station_number[ec.stations$gazetted_stream_name == "Harrison River"] <- "08MG013"
# Harrison is also "wack" as a "Summer" timing group. So manually changing its timing
ec.stations$start[ec.stations$gazetted_stream_name == "Harrison River"] <- 313 # Nov 8
ec.stations$end[ec.stations$gazetted_stream_name == "Harrison River"] <- 321 # Nov 16

cal.flow <- hy_daily_flows(station_number = ec.stations$ec_station_number,
                           start_date = "2004-01-01")
cal.flow <- cal.flow %>% 
  mutate(Year = format(as.Date(Date, format="%Y/%m/%d"),"%Y"), # pull out year
         JDay = lubridate::yday(Date)) %>%  # pull out day of year
  filter(!is.na(Value)) %>% # remove missing data rows
  group_by(STATION_NUMBER, Year) %>% # grouping by station number and year
  mutate(records = n()) %>% 
  ungroup() %>% 
  mutate(complete = records >= 364) # flag incomplete years
# although, it should only matter for the periods when we are taking the average.

cal.flow.summary <- cal.flow %>% 
  # filter(Year >= 2024-20) %>% # doing this step with start_date in hy_daily_flows now
  group_by(STATION_NUMBER, JDay) %>% 
  summarise(average = mean(Value),
            n = n(),
            p10 = quantile(Value, 0.10),
            p25 = quantile(Value, 0.25),
            p50 = quantile(Value, 0.50),
            p75 = quantile(Value, 0.75),
            p90 = quantile(Value, 0.90))

ggplot(cal.flow.summary,
       aes(x = JDay, y = average)) +
  geom_line() +
  geom_line(aes(y = p10), col = "purple") +
  geom_line(aes(y = p90), col = "purple") +
  # geom_vline(xintercept = c(245, 281)) 
  facet_wrap(~STATION_NUMBER, scales = "free_y")

ec.stations$mean <- NA
ec.stations$sd <- NA

for(i in 1:nrow(ec.stations)){
  temp <- cal.flow.summary %>% 
    filter(STATION_NUMBER == ec.stations$ec_station_number[i],
           JDay >= ec.stations$start[i],
           JDay <= ec.stations$end[i])
  ec.stations$mean[i] = mean(temp$average)
  ec.stations$sd[i] = sd(temp$average)
}

cal <- cal %>% 
  left_join(., dplyr::select(ec.stations, !ec_station_number)) %>% 
  mutate(discharge.sc = (water_discharge - mean) / sd)

hist(cal$discharge.sc) # there is one big outlier - need to confirm if it is "real"
filter(cal, discharge.sc == max(discharge.sc)) # from Harrison
hist(cal$discharge.sc[cal$gazetted_stream_name == "Harrison River"])
hist(cal$discharge.sc[cal$discharge.sc < max(cal$discharge.sc)]) # there is one big outlier - need to confirm if it is "real"

# from the images below, it looks legitimate. Just big swings in Harrison data.
filter(cal.flow,
       STATION_NUMBER == ec.stations$ec_station_number[ec.stations$gazetted_stream_name == "Harrison River"],
       Year >= 2015, Year <= 2017) %>% 
  ggplot(aes(x = Date, y = Value)) +
  geom_line()

filter(cal.flow.summary,
       STATION_NUMBER == ec.stations$ec_station_number[ec.stations$gazetted_stream_name == "Harrison River"]) %>% 
  ggplot(aes(x = JDay, y = average)) +
  geom_line() +
  geom_line(aes(y = p90), col = "red") +
  geom_line(aes(y = p10), col = "red") +
  geom_vline(xintercept = c(313, 321))

ggplot(cal,
       aes(x = lnlpe.sc, y = log(hpe),
           colour = discharge.sc)) +
  geom_point()
```


## Correlation Coefficients
```{r correlation figure}
cal.cor <- round(cor(dplyr::select(cal, size_recode, stability_recode, water_clarity_recode, 
                 substrate_shade_recode, canopy_cover_recode,
                 lpe_method_recode, ln.lpe.sc, discharge.sc)), 3)
cal.cor

#png(file="figures/Discharge - Sockeye Calibration Correlation Plot.png",
#    units = "in", width=10, height=5, res = 250)

chart.Correlation(dplyr::select(cal, size_recode, stability_recode, water_clarity_recode, 
                 substrate_shade_recode, canopy_cover_recode,
                 lpe_method_recode, ln.lpe.sc, discharge.sc), 
                  histogram = TRUE, method = "pearson")
#dev.off()
```
Number of Variables with strong correlations
size has 3 > 0.5
Stability has 2 > 0.69
Water clarity has 5 > 0.4
substrade_shade is good
canopy_cover has 1 > 0.7 and 2 > 0.5
lpe_method has 2 > 0.4
discharge is not correlated with anything!

Prioritizing Size since it is a known factor from prior analysis.
That would remove Stability, Water_Clarity, and Canopy_Cover.
Keep Substrate Shade and LPE Method along with Size and Discharge.

## Model
```{r}
count(cal, gazetted_stream_name) # stellako a major driver
count(cal, size)
```
Not enough samples per river to make river a random effect.

###Discharge
```{r}
m.cal <- glm.nb(hpe ~ size + ln.lpe.sc + size:ln.lpe.sc + 
                  discharge.sc + 
                  lpe_method + 
                  lpe_method*ln.lpe.sc +
                  discharge.sc*ln.lpe.sc +
                  substrate_shade, 
    data=cal, 
    na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
    link="log")

summary(m.cal)
# Pearson residual is more appropriate for a glm that doesn't assume gaussian error distribution
plot((fitted(m.cal)), residuals(m.cal, type = "pearson"))

m.cal.coef <- as.data.frame(summary(m.cal)$coefficients)
m.cal.coef$parameter = row.names(m.cal.coef)
colnames(m.cal.coef)[2] = "std.error"

ggplot(m.cal.coef[-1,],
       aes(x = parameter, y = Estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = Estimate-std.error, ymax = Estimate+std.error)) +
  geom_hline(yintercept = 0) +
  coord_flip()
```
Discharge doesn't look like it has much of an effect.

#### Dredge
```{r}
m.dredge <- dredge(m.cal, beta = "sd", rank = "AICc")
m.dredge

m.dredge <- as.data.frame(m.dredge %>% filter(delta < 4)) %>% 
  dplyr::select(discharge.sc:delta)

write.csv(m.dredge, file = "./tables/01_06-Discharge-All.csv", na = "", row.names = F)
```

Dredge reveals that the top model and generally more parsimonious model does
not have discharge, nor lpe_method, nor either of the interactions.

Discharge only has about 0.146 total weight within the top 4 AICc. 

###Discharge No Stellako
Trying with removing Stellako
```{r}
cal5 <- filter(cal, gazetted_stream_name != "Stellako River")
cal5 <- cal5 %>% 
  mutate(ln.lpe = log(lpe),
         ln.lpe.sc = scale(ln.lpe, center=T, scale=T)[,1])

m.cal5 <- glm.nb(hpe ~ size + ln.lpe.sc + size*ln.lpe.sc +
                   discharge.sc + 
                   discharge.sc*ln.lpe.sc +
                  lpe_method + 
                  substrate_shade, 
    data=cal5, 
    na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
    link="log")

summary(m.cal5)
# Pearson residual is more appropriate for a glm that doesn't assume gaussian error distribution
plot((fitted(m.cal5)), residuals(m.cal5, type = "pearson"))

m.cal5.coef <- as.data.frame(summary(m.cal5)$coefficients)
m.cal5.coef$parameter = row.names(m.cal5.coef)
colnames(m.cal5.coef)[2] = "std.error"

ggplot(m.cal5.coef[-1,],
       aes(x = parameter, y = Estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = Estimate-std.error, ymax = Estimate+std.error)) +
  geom_hline(yintercept = 0) +
  coord_flip()

m.dredge5 <- dredge(m.cal5, beta = "sd", rank = "AICc")
m.dredge5
```
Discharge is not significant at all in this formulation.
Decision to stick with "most data rich" and keep Stellako

#### Select Top
"TOP" in this case, is just the "top" model that also includes Discharge within
4 AICc of the top-ranked model.

Since there is no interaction between discharge and lpe, I am having a hard-time
convincing myself that there is a biological reason / connection for the
slope only model to work...

I guess we can think of it as it is soaking up additional uncertainty after lpe
is describing most of the relationship?
```{r}
m.cal.top <- glm.nb(hpe ~ size + ln.lpe.sc + substrate_shade + discharge.sc, 
    data=cal, 
    na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
    link = "log")

summary(m.cal.top)
# Pearson residual is more appropriate for a glm that doesn't assume gaussian error distribution
png(file = "./figures/01_Model_Diagnostics/01-06-Discharge-All-Residuals.png",
    height = 5, width = 5, units = "in", res = 96)
plot(log(fitted(m.cal.top)), residuals(m.cal.top, type = "pearson"),
     xlab = "ln(Fitted Values)", ylab = "Pearson Residuals",
     main = "Specialized (6) - Discharge All Systems Model")
abline(h = 0)
dev.off()

# possible outliers
cal$usid[which(abs(residuals(m.cal.top, type = "pearson"))>3)]
# Influential Data
plot(cooks.distance(m.cal.top))
abline(h = 4/nrow(cal), col = "blue") # Identifies points with High Influence
abline(h = 3*mean(cooks.distance(m.cal.top)), col = "red")  # Identifies points with High Influence
# which system-years are possible issues or outliers?
# high leverage
cal$usid[which(cooks.distance(m.cal.top) > 3*mean(cooks.distance(m.cal.top)))]


m.cal.top.coef <- as.data.frame(summary(m.cal.top)$coefficients)
m.cal.top.coef$parameter = row.names(m.cal.top.coef)
colnames(m.cal.top.coef)[2] = "std.error"

ggplot(m.cal.top.coef[-1,],
       aes(x = reorder_within(parameter, Estimate,
                              parameter, mean), 
           y = Estimate)) +
  geom_point() +
  scale_x_reordered() +
  geom_errorbar(aes(ymin = Estimate-std.error, ymax = Estimate+std.error)) +
  geom_hline(yintercept = 0) +
  coord_flip() +
  labs(x = "Parameter")

ggsave("./figures/01_Model_Diagnostics/01-06-Discharge-All-Parameters.png",
       width = 5, height = 5, dpi = 96)

m.coef <- dplyr::select(m.cal.top.coef, parameter, Estimate, std.error) %>% 
  arrange(desc(abs(Estimate))) %>% 
  mutate(cross.zero = ifelse(abs(Estimate) < abs(std.error), "Yes", "No"))
write.csv(m.coef, file = "./tables/01_06-Discharge-All-Parameters.csv",
          row.names = F)
```

## Leave ln Out
```{r}
model.test <- data.frame(model=character(),
                 usid=character(), 
                 fit=numeric(), 
                 se.fit=numeric(),
                 calibrated.estimate=numeric()) 
```

```{r}
models <- list(m.cal.top)

saveRDS(m.cal.top, file = "./r_objects/models/All_Systems_Discharge_Model.rds")

specifications_out <- data.frame(model = "All_Systems_Discharge_Model",
                                 target = c("lpe","lpe","lpe",
                                            "discharge","discharge","discharge"),
                                 order = c(1,2,3,
                                           1,2,3),
                                 type = c("log","center","scale",
                                          "system-specific","center","scale"),
                                 transform = c(NA, mean(cal$ln.lpe), sd(cal$ln.lpe),
                                               NA, NA, NA))

ec.stations.specs <- dplyr::select(ec.stations, gazetted_stream_name, run_timing_group, mean, sd) %>% 
  pivot_longer(., cols = c("mean", "sd"), names_to = "type", values_to = "transform" ) %>% 
  mutate(model = "All_Systems_Discharge_Model",
         target = paste(gazetted_stream_name, run_timing_group, sep = "-"),
         order = rep(c(1,2), length.out = nrow(.)),
         ) %>% 
  dplyr::select(model, target, order, type, transform)

specifications_out <- rbind(specifications_out, ec.stations.specs)

spec.out <- spec.out %>% 
  filter(model != "All_Systems_Discharge_Model") %>% 
  rbind(., specifications_out)

saveRDS(spec.out, file=modout)
```
Need to add a running of the model without the ln data point.
```{r}
for(j in 1:length(models)){
  model = models[[j]]
  for(i in 1:nrow(cal)){
    m.temp <- glm.nb(as.character(model$call)[2],
                     data=cal[-i, ],
                     na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                     link="log")
    p.temp <- predict(m.temp, type="link", cal[i, ], se.fit=T)
    model.test <- rbind(model.test, data.frame(model = as.character(model$call)[2],
                                 usid = cal$usid[i],
                                 fit = p.temp$fit,
                                 se.fit = p.temp$se.fit,
                                 calibrated.estimate = exp(p.temp$fit)))
  }
}
```

```{r}
model.rank <- left_join(model.test, 
                        dplyr::select(cal,
                                      usid, hpe, watershed_group:year, lpe_method)) %>% 
  mutate(error = calibrated.estimate - hpe,
         squared.error = error^2,
         abs.error = abs(error),
         per.error = error/hpe,
         abs.per.error = abs(per.error),
         rank.set.level = 6)

```
### Plots
```{r plot of error}
model.rank %>% 
  ggplot(aes(x = hpe, y = error)) +
  geom_point() +
  scale_y_continuous(labels = label_comma())
```

```{r plot of percent error}
model.rank %>% 
  ggplot(aes(x = hpe, y = per.error)) +
  geom_point()
```

```{r plot of absolute percent error}
model.rank %>% 
  ggplot(aes(x = hpe, y = abs.per.error)) +
  geom_point()
```

```{r plot of absolute percent error by system}
set.seed(234)
model.rank %>% 
  ggplot(aes(x = reorder(gazetted_stream_name, abs.per.error, FUN = median), 
             # reorder FUN = median, reorders based on the median
             y = abs.per.error,
             colour = model)) +
  geom_boxplot(alpha = 0.5, outlier.shape = NA) +
  geom_hline(yintercept = 0.5, colour = "grey") +
  geom_point(position = "jitter", alpha = 0.5) +
  coord_flip() +
  labs(title = "All Systems Dishcharge Model Performance",
       y = "Absolute percent error", x = "Streams",
       colour = "Model Types")

#ggsave("figures/Model Performance/03_Model_Performance_All_Systems_Discharge.png",
#       height = 6, width = 8, dpi = 150)
```

```{r}
filter(model.rank,
       abs.per.error > 0.4)
```

# Save Model Ranks
```{r}
saveRDS(model.rank, 
        "r_objects/03_Model_Ranks_from_All_Systems_Discharge_Model.rds")
```

```{r}
count(model.rank, gazetted_stream_name)
```

