---
title: "01-07-Tachie-Harrison-Specific"
author: "Michael Arbeider"
date: "2026-03-06"
---

# Set-up
## Packages
```{r}
rm(list = ls())
library(tidyverse) # most %>% 
theme_set(theme_bw()) # ggplot
library(MASS) # boxcox
library(lme4) # generalized linear regressions
library(MuMIn) # dredge
library(PerformanceAnalytics) # correlation figure
library(car)       # for vif()
library(patchwork) # for multi plot ggplot graphs
library(tidyhydat) # hy_daily for hydrometric data
```

## Functions
```{r}
source("00-Functions.R") # scale_x_reordered, rank.summary(), round2()
```

## Data
```{r}
cal <-readRDS("r_objects/sockeye_calibration_data.rds")   

glimpse(cal)
```

## Specifications FILE
```{r}
modout <- "./r_objects/models/model_specifications.rds"
spec.out <- readRDS(modout)
```

## Base Plots
https://stackoverflow.com/questions/57153428/r-plot-color-combinations-that-are-colorblind-accessible
```{r colour pallet}
size.cols = c("#E69F00", # Large
              "#56B4E9", # Medium (including the smalls)
              "#D55E00", # V. Small
              "#CC79A7") # X-Large
```

# Harrison Specific
```{r}
har <- cal %>% filter(gazetted_stream_name == "Harrison River")
```

```{r}
har.flow <- hy_daily_flows(station_number = "08MG013") # from ec_comments

har.flow <- har.flow %>% 
  mutate(Year = as.numeric(format(as.Date(Date, format="%Y/%m/%d"),"%Y")), # change to date
         JDay = lubridate::yday(Date)) %>% # create a Julian Day
  filter(!is.na(Value)) %>% # remove missing values
  group_by(Year) %>% # want t count per year
  mutate(records = n()) %>% 
  ungroup() %>% # ungroup
  mutate(complete = records >= 364) %>% # is the total days in the year >= a year?
  filter(complete == TRUE) # only keep the full years.

# average flow per Julian Day for last 20 years
har.flow.summary <- har.flow %>% 
  filter(Year >= max(har.flow$Year)-20) %>% 
  group_by(JDay) %>% 
  summarise(average = mean(Value),
            p10 = quantile(Value, 0.10),
            p25 = quantile(Value, 0.25),
            p50 = quantile(Value, 0.50),
            p75 = quantile(Value, 0.75),
            p90 = quantile(Value, 0.90))
```

```{r}
har$water_discharge <- har.flow$Value[har.flow$Date %in% as.Date(har$lpe_date)]

# Nov  8 - Nov 16 - range of Summer Peak Spawn dates from 2023 Prelim for Harrison
har.mean = mean(har.flow.summary$average[har.flow.summary$JDay %in% c(313:351)])
har.sd = sd(har.flow.summary$average[har.flow.summary$JDay %in% c(313:351)])

mean(har$water_discharge)
sd(har$water_discharge)

har <- har %>% 
  mutate(lpe.sc = scale(lpe, center=T, scale=T)[,1],
         lnlpe = log(lpe),
         lnlpe.sc = scale(lnlpe, center=T, scale=T)[,1],
         discharge.sc = (water_discharge-har.mean)/har.sd)
```

```{r}
ggplot(har.flow.summary,
       aes(x = JDay, y = average)) +
  geom_line(colour = "blue") +
  geom_line(aes(y = p10)) +
  geom_line(aes(y = p90)) +
  geom_vline(xintercept = c(313, 351)) + # Nov  8 - Nov 16 - range of Summer Peak Spawn dates from 2023 Prelim for Harrison
  geom_point(data = har,
             aes(x = lubridate::yday(lpe_date),
                 y = water_discharge))
```

```{r}
ggplot(har,
       aes(x = lnlpe.sc, y = log(hpe),
           colour = discharge.sc)) +
  geom_point()
```

## Global Model
### Covariate Selection
#### Correlation Coefficients
Selecting complete cases and excluding ones that are uniform (e.g. substrate shade)
```{r correlation figure}
har.cor <- round2(cor(dplyr::select(har, water_clarity_recode, 
                 bankfull_recode, brightness_recode, cc_recode, fish_vis_recode,
                 water_vis_recode, discharge.sc, lnlpe.sc)), 3)
har.cor

chart.Correlation(dplyr::select(har, water_clarity_recode, 
                 bankfull_recode, brightness_recode, cc_recode, fish_vis_recode,
                 water_vis_recode, discharge.sc, lnlpe.sc), 
                  histogram = TRUE, method = "pearson")
```
Only Brightness and Cloud Cover have a very high correlation 0.98
A handful have a correlation of >0.5 but < 0.6.
fish_vis has 2 > 0.5, water clarity and lpe
water_clarity is also 0.5 with lpe

But there are only 10 data points, so need to pare down to start even.
Global Model did not work with 7 covariates.
Primary interests from an observer interaction perspective are:
lpe
dishcharge
water_clarity
brightness

dropping:
bankfull
fish_vis and cloud cover and water visibility

fish_vis is a little more subjective than brightness.
flip-flopping between which to test between the two.
brightness was not correlated with lpe, so chosing it for now.
```{r}
m.global <- glm.nb(hpe ~ lnlpe.sc*discharge.sc + lnlpe.sc*water_clarity  + 
                     lnlpe.sc*brightness, 
                   na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                   data=har,
                   link="log")
# glm.nb default negative binomial family is neg binom 2

# Model Checking
# plot(log(fitted(m.global)), residuals(m.global, type = "response")) # uninformative
# https://www.sfu.ca/sasdoc/sashtml/insight/chap39/sect56.htm
# Pearson residual is more appropriate for a glm that doesn't assume gaussian error distribution
plot(log(fitted(m.global)), residuals(m.global, type = "pearson"))
# there are a couple high residuals
plot(log(fitted(m.global)), sqrt(abs(residuals(m.global, type = "pearson"))))
# Influential Data
plot(cooks.distance(m.global))
abline(h = 4/nrow(har), col = "blue") # Identifies points with High Influence
abline(h = 3*mean(cooks.distance(m.global)), col = "red")  # Identifies points with High Influence
# which system-years are possible issues or outliers?
har$usid[which(cooks.distance(m.global) == max(cooks.distance(m.global)))]
har$usid[which(cooks.distance(m.global) > 3*mean(cooks.distance(m.global)))]
# Model Summary
summary(m.global)
```

#### Variance Inflation Factors
general definition:
https://online.stat.psu.edu/stat462/node/180/
on interactions and categorical variables:
https://statisticalhorizons.com/multicollinearity/
```{r}
vif(m.global)
```

Summary:
nothing stands out RE VIF
### Dredge
```{r}
m.dredge.global <- dredge(m.global, beta = "none", rank = "AICc")
m.dredge.global
```

Only 1 model under 4 detla AICc
Only Includes lnlpe.sc

That is the same model as the "Individual" model.
Thus, I will not repeat it.

# Tachie Specific
```{r}
tach <- cal %>% filter(gazetted_stream_name == "Tachie River")
```

```{r}
tach.flow <- hy_daily_flows(station_number = tach$ec_station_number[3]) # from ec_comments

tach.flow <- tach.flow %>% 
  mutate(Year = as.numeric(format(as.Date(Date, format="%Y/%m/%d"),"%Y")), # change to date
         JDay = lubridate::yday(Date)) %>% # create a Julian Day
  filter(!is.na(Value)) %>% # remove missing values
  group_by(Year) %>% # want t count per year
  mutate(records = n()) %>% 
  ungroup() %>% # ungroup
  mutate(complete = records >= 364) %>% # is the total days in the year >= a year?
  filter(complete == TRUE) # only keep the full years.

# average flow per Julian Day for last 20 years
tach.flow.summary <- tach.flow %>% 
  filter(Year >= max(tach.flow$Year)-20) %>% 
  group_by(JDay) %>% 
  summarise(average = mean(Value),
            p10 = quantile(Value, 0.10),
            p25 = quantile(Value, 0.25),
            p50 = quantile(Value, 0.50),
            p75 = quantile(Value, 0.75),
            p90 = quantile(Value, 0.90))
```

```{r}
tach$water_discharge <- tach.flow$Value[tach.flow$Date %in% as.Date(tach$lpe_date)]

# Sep  7 - Sep 30 - range of Summer Peak Spawn dates from 2023 Prelim for Tachie
pstrt <- 251
pend <- 274
tach.mean = mean(tach.flow.summary$average[tach.flow.summary$JDay %in% c(pstrt:pend)])
tach.sd = sd(tach.flow.summary$average[tach.flow.summary$JDay %in% c(pstrt:pend)])

mean(tach$water_discharge)
sd(tach$water_discharge)

tach <- tach %>% 
  mutate(lpe.sc = scale(lpe, center=T, scale=T)[,1],
         lnlpe = log(lpe),
         lnlpe.sc = scale(lnlpe, center=T, scale=T)[,1],
         discharge.sc = (water_discharge-tach.mean)/tach.sd)
```

```{r}
ggplot(tach.flow.summary,
       aes(x = JDay, y = average)) +
  geom_line(colour = "blue") +
  geom_line(aes(y = p10)) +
  geom_line(aes(y = p90)) +
  geom_vline(xintercept = c(pstrt, pend)) +
  geom_point(data = tach,
             aes(x = lubridate::yday(lpe_date),
                 y = water_discharge))
```

```{r}
ggplot(tach,
       aes(x = lnlpe, y = log(hpe),
           colour = discharge.sc)) +
  geom_point()
```

## Global Model
### Covariate Selection
#### Correlation Coefficients
Selecting complete cases and excluding ones that are uniform (e.g. substrate shade)
```{r correlation figure}
tach.cor <- round2(cor(dplyr::select(tach, 
                 brightness_recode, cc_recode, fish_vis_recode,
                 discharge.sc, lnlpe.sc)), 3)
tach.cor

chart.Correlation(dplyr::select(tach, 
                 brightness_recode, cc_recode, fish_vis_recode,
                 discharge.sc, lnlpe.sc), 
                  histogram = TRUE, method = "pearson")
```
Only Brightness and Cloud Cover have a very high correlation 0.98
Both Brightness and Cloud Cover are correlated (randomly?) with discharge and cc with fish vis
Fish vis and discharge are not correlated.

Reducing to only include:
Fish Vis
Discharge
Lpe
```{r}
m.global <- glm.nb(hpe ~ lnlpe.sc*discharge.sc + lnlpe.sc*fish_vis, 
                   na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                   data=tach,
                   link="log")
# glm.nb default negative binomial family is neg binom 2

# Model Checking
# plot(log(fitted(m.global)), residuals(m.global, type = "response")) # uninformative
# https://www.sfu.ca/sasdoc/sashtml/insight/chap39/sect56.htm
# Pearson residual is more appropriate for a glm that doesn't assume gaussian error distribution
plot(log(fitted(m.global)), residuals(m.global, type = "pearson"))
# there are a couple high residuals
plot(log(fitted(m.global)), sqrt(abs(residuals(m.global, type = "pearson"))))
# Influential Data
plot(cooks.distance(m.global))
abline(h = 4/nrow(tach), col = "blue") # Identifies points with High Influence
abline(h = 3*mean(cooks.distance(m.global)), col = "red")  # Identifies points with High Influence
# which system-years are possible issues or outliers?
tach$usid[which(cooks.distance(m.global) == max(cooks.distance(m.global)))]
tach$usid[which(cooks.distance(m.global) > 3*mean(cooks.distance(m.global)))]
# Model Summary
summary(m.global)

m.global.no.interaction <- glm.nb(hpe ~ lnlpe.sc + discharge.sc + fish_vis, 
                   na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                   data=tach,
                   link="log")
# glm.nb default negative binomial family is neg binom 2

# Plotting the Coefficients
m.global.coef <- as.data.frame(summary(m.global)$coefficients)
m.global.coef$parameter = row.names(m.global.coef)
colnames(m.global.coef)[2] = "std.error"

ggplot(m.global.coef,
       aes(x = parameter, y = Estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = Estimate-std.error, ymax = Estimate+std.error)) +
  geom_hline(yintercept = 0) +
  coord_flip()
```

#### Variance Inflation Factors
general definition:
https://online.stat.psu.edu/stat462/node/180/
on interactions and categorical variables:
https://statisticalhorizons.com/multicollinearity/
```{r}
vif(m.global)
vif(m.global.no.interaction)
```

Summary:
nothing stands out RE VIF

### Dredge
```{r}
m.dredge.global <- dredge(m.global, beta = "none", rank = "AICc")
m.dredge.global
```

Only 1 model under 4 detla AICc
Only Includes lnlpe.sc

That is the same model as the "Individual" model.
Thus, it will not be saved as a new model to try.

## Ratio Model
```{r}
ggplot(tach,
       aes(y = log(hpe)/lnlpe, x = discharge.sc,
           colour = fish_vis,
           shape = brightness)) +
  geom_point()

ggplot(tach,
       aes(y = hpe/lpe, x = discharge.sc,
           colour = fish_vis,
           shape = brightness)) +
  geom_point()

ggplot(tach,
       aes(y = hpe/lpe, x = log(lpe),
           colour = fish_vis,
           shape = brightness)) +
  geom_point()

ggplot(tach,
       aes(y = log(hpe)/lnlpe, x = log(lpe),
           colour = fish_vis,
           shape = brightness)) +
  geom_point()
```



```{r}
m.global <- lm(log(hpe)/log(lpe) ~ discharge.sc*fish_vis + discharge.sc*brightness, 
                   na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                   data=tach)
# glm.nb default negative binomial family is neg binom 2

# Model Checking
# plot(log(fitted(m.global)), residuals(m.global, type = "response")) # uninformative
# https://www.sfu.ca/sasdoc/sashtml/insight/chap39/sect56.htm
# Pearson residual is more appropriate for a glm that doesn't assume gaussian error distribution
plot(log(fitted(m.global)), residuals(m.global, type = "pearson"))
# there are a couple high residuals
plot(log(fitted(m.global)), sqrt(abs(residuals(m.global, type = "pearson"))))
# Influential Data
plot(cooks.distance(m.global))
abline(h = 4/nrow(tach), col = "blue") # Identifies points with High Influence
abline(h = 3*mean(cooks.distance(m.global)), col = "red")  # Identifies points with High Influence
# which system-years are possible issues or outliers?
which(cooks.distance(m.global) == max(cooks.distance(m.global)))
tach$usid[which(cooks.distance(m.global) > 3*mean(cooks.distance(m.global)))]
# Model Summary
summary(m.global)

m.global.no.interaction <- lm(log(hpe)/log(lpe) ~ discharge.sc + fish_vis + brightness, 
                   na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                   data=tach)
# glm.nb default negative binomial family is neg binom 2

# Plotting the Coefficients
m.global.coef <- as.data.frame(summary(m.global)$coefficients)
m.global.coef$parameter = row.names(m.global.coef)
colnames(m.global.coef)[2] = "std.error"

ggplot(m.global.coef,
       aes(x = parameter, y = Estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = Estimate-std.error, ymax = Estimate+std.error)) +
  geom_hline(yintercept = 0) +
  coord_flip()
```

#### Variance Inflation Factors
general definition:
https://online.stat.psu.edu/stat462/node/180/
on interactions and categorical variables:
https://statisticalhorizons.com/multicollinearity/
```{r}
vif(m.global)
vif(m.global.no.interaction)
```

Summary:
nothing stands out RE VIF

### Dredge
```{r}
m.dredge.global <- dredge(m.global, beta = "none", rank = "AICc")
m.dredge.global
```

Most parsimonious is Intercept Only model.
One Model within 4 AICc, includes discharge.
### Model Averaged
```{r}
ma.global = model.avg(m.dredge.global, subset = delta < 4, fit = TRUE)
```
May test if this is better than individual.

# Harrison + Tachie
```{r}
combo <- rbind(har, tach)
```

```{r}
ggplot(combo,
       aes(x = lnlpe.sc, y = log(hpe),
           colour = discharge.sc)) +
  geom_point() +
  geom_smooth(method = "lm")

ggplot(combo,
       aes(x = lpe, y = hpe,
           colour = discharge.sc)) +
  geom_point() +
  geom_smooth(method = "lm")

hist(combo$discharge.sc)
```

## Global Model
### Covariate Selection
#### Correlation Coefficients
Selecting complete cases and excluding ones that are uniform (e.g. substrate shade)
```{r correlation figure}
combo.cor <- round2(cor(dplyr::select(combo, water_clarity_recode,
                 brightness_recode, cc_recode, fish_vis_recode,
                 discharge.sc, lnlpe.sc)), 3)
combo.cor

chart.Correlation(dplyr::select(combo, water_clarity_recode,
                 brightness_recode, cc_recode, fish_vis_recode,
                 discharge.sc, lnlpe.sc), 
                  histogram = TRUE, method = "pearson")
```
Only Brightness and Cloud Cover have a very high correlation 0.98


```{r}
m.global.vis <- glm.nb(hpe ~ lnlpe.sc*discharge.sc + lnlpe.sc*fish_vis, 
                   na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                   data=tach,
                   link="log")
# glm.nb default negative binomial family is neg binom 2

# Model Checking
# plot(log(fitted(m.global)), residuals(m.global, type = "response")) # uninformative
# https://www.sfu.ca/sasdoc/sashtml/insight/chap39/sect56.htm
# Pearson residual is more appropriate for a glm that doesn't assume gaussian error distribution
plot(log(fitted(m.global.vis)), residuals(m.global.vis, type = "pearson"))
# there are a couple high residuals
plot(log(fitted(m.global.vis)), sqrt(abs(residuals(m.global.vis, type = "pearson"))))
# Influential Data
plot(cooks.distance(m.global.vis))
abline(h = 4/nrow(combo), col = "blue") # Identifies points with High Influence
abline(h = 3*mean(cooks.distance(m.global.vis)), col = "red")  # Identifies points with High Influence
# which system-years are possible issues or outliers?
which(cooks.distance(m.global.vis) == max(cooks.distance(m.global.vis)))
combo$usid[which(cooks.distance(m.global.vis) > 3*mean(cooks.distance(m.global.vis)))]
# Model Summary
summary(m.global.vis)

m.global.bright <- glm.nb(hpe ~ lnlpe.sc*discharge.sc + 
                     lnlpe.sc*brightness, 
                   na.action="na.fail", # https://stackoverflow.com/questions/25281739/dredge-function-error-r-package-mumln
                   data=tach,
                   link="log")
# glm.nb default negative binomial family is neg binom 2

# Pearson residual is more appropriate for a glm that doesn't assume gaussian error distribution
plot(log(fitted(m.global.bright)), residuals(m.global.bright, type = "pearson"))
# there are a couple high residuals
plot(log(fitted(m.global.bright)), sqrt(abs(residuals(m.global.bright, type = "pearson"))))
# Influential Data
plot(cooks.distance(m.global.bright))
abline(h = 4/nrow(combo), col = "blue") # Identifies points with High Influence
abline(h = 3*mean(cooks.distance(m.global.bright)), col = "red")  # Identifies points with High Influence
# which system-years are possible issues or outliers?
which(cooks.distance(m.global.bright) == max(cooks.distance(m.global.bright)))
combo$usid[which(cooks.distance(m.global.bright) > 3*mean(cooks.distance(m.global.bright)))]
# Model Summary
summary(m.global.bright)
```

#### Variance Inflation Factors
general definition:
https://online.stat.psu.edu/stat462/node/180/
on interactions and categorical variables:
https://statisticalhorizons.com/multicollinearity/
```{r}
vif(m.global.vis)
vif(m.global.bright)
```

Summary:
Interaction terms - ignoring.

### Dredge
```{r}
m.dredge.vis <- dredge(m.global.vis, beta = "none", rank = "AICc")
m.dredge.vis

m.dredge.bright <- dredge(m.global.bright, beta = "none", rank = "AICc")
m.dredge.bright
```

only models < 4 AICc are those with only lpe as the predictor.