---
title: "02-01-Model-Ranks"
author: "Michael Arbeider"
date: "2026-03-06"
---

This Code is used to pre-screen the General Model Suite.

# Set-up
## Packages
```{r}
rm(list = ls())
library(tidyverse) # most %>% 
theme_set(theme_bw()) # ggplot
library(patchwork) # for multi plot ggplot graphs
library(scales) # for label_comma()
```

## Functions
```{r}
source("00-Functions.R") # scale_x_reordered, rank.summary(), round2()
```


## Data
Reading in calibration data and then reading in all of the Model Rank outputs
```{r}
cal <-readRDS("r_objects/sockeye_calibration_data.rds")   

glimpse(cal)

# Combining all model rank outputs from other rmd files
 # Data path folder for Model Ranks
r_objects <- "./r_objects"
 # Pull all the data files
data_path <- dir(r_objects, full.names = TRUE)
  # Obtain the file names that inlcude Model Ranks
fileNames <- data_path[grep(pattern = "Model_Ranks", data_path)]
# remove the one time test of Harrison Individual MoR
fileNames <- fileNames[fileNames != "./r_objects/01_Model_Ranks_from_MoR_Harrison_Model.rds"]

 # Create a character string with the Model Type / Data Source
source <- sub(pattern = ".*from_", "", fileNames)
source <- sub(pattern = ".rds", "", source)

# looping through file names to read in RDS and create a "source" field based on the file name
for(i in 1:length(fileNames)){
  if(i==1) mod.ranks <- readRDS(fileNames[i])
  if(i==1) mod.ranks$source = source[1]
  if(i!=1) mod.ranks.temp <- readRDS(fileNames[i])
  if(i!=1) mod.ranks.temp$source = source[i]
  if(i!=1) mod.ranks = rbind(mod.ranks, mod.ranks.temp)
}
# removing the temporary binding dataframe that is a part of mod.ranks
rm(mod.ranks.temp) 

# joining the "classic" 1.8 expansion to every estimate
# this will be used to create a Scaled Error relative to the current 1.8 expansion
classic <- mod.ranks %>% 
  filter(source == "All_Data_Classic_Expansion") %>% 
  dplyr::select(usid, fit) %>% 
  rename(classic = fit) # calling the 1.8 fit "classic"

# joining the classic 1.8 and calculating the absolute difference between the 1.8 expansion and the High Accuracy estimate
mod.ranks <- mod.ranks %>% 
  left_join(., classic) %>% 
  mutate(abs.classic.error = abs(hpe - classic))

# removing "All_Data_Classic_Expansion" from the mod.ranks as we won't be selecting it
mod.ranks <- filter(mod.ranks, source != "All_Data_Classic_Expansion")

# joining other system specific information and creating Rank-Set levels
mod.ranks <- mod.ranks %>% 
  left_join(., dplyr::select(cal, usid, size, water_clarity)) %>% 
  # making some Rank.Set columns for filtering later
  mutate(agg.size = case_when(
    size %in% c("Small", "Medium") ~ "SM",
    size %in% c("Large", "X-Large") ~ "LXL"
    ),
    size.clarity = paste0(size, "_", water_clarity))

# making the same rank set columns in the cal data frame
cal <- cal %>% 
  mutate(agg.size = case_when(
    size %in% c("Small", "Medium") ~ "SM",
    size %in% c("Large", "X-Large") ~ "LXL"
    ),
    agg.size.clarity = paste0(agg.size, "_", water_clarity),
    size.clarity = paste0(size, "_", water_clarity))

# Save mod.ranks once
saveRDS(mod.ranks, file = "./r_objects/02_01_combined_model_LOO_rank_data.rds")
```

Remove the Low Abundance Models, these will be assessed in 02-02.
```{r}
mod.ranks <- mod.ranks %>% 
  filter(model != "MoR_Low_Abundance_Models")
```


Model Sample Size Table
```{r}
mod.rank.sum.by.model <- mod.ranks %>% 
  ungroup() %>% 
  count(source, model, lpe_method)

mod.ranks %>% 
  ungroup() %>% 
  count(source, model, lpe_method, size, water_clarity)

#write.csv(mod.rank.sum.by.model, file = "./tables/04_02_Model_Ranks_by_Model_Simple.csv",
#          row.names = F)

count(cal, lpe_method, size)
```

# Model Ranking by Rank Sets
```{r}
sample.minimum <- 6
```

## 01 Whole Dataset
by method
```{r}
for(j in c("Ground", "Aerial")){ # loop for each method
# create a SOURCE filter for only models that have the equivalent number of samples
    tempfilter2 <- mod.ranks %>% 
  filter(rank.set.level <= 1) %>% 
      count(source) %>%  
      filter(n == max(n))
   
    # apply the rank.summary() function filtering the mod.ranks by the above filters
    rank.temp <- rank.summary(filter(mod.ranks,
                                     source %in% tempfilter2$source,
                                     lpe_method == j))
    
    # Create a sys.meth.set or RANK.SET Name
    rank.temp <- rank.temp %>% 
      mutate(sys.meth.set = paste0("01_All_General - All - ", j),
             rank.filter = "All",
             rank.method = j,
             rank.set = "01_All_General",
             special.level = 1) # identifying that the special level is 1
    # special.level will be used to later to determine if the top model is from
    # the specialized level or if it is from a more aggregated level
    
    if(j == "Ground") rank.compare.general <- rank.temp
    if(j != "Ground") rank.compare.general <- rbind(rank.compare.general, rank.temp)
} # end of method loop j
```


## 02 Agg.Size
Creating Rank Summary's for various 

System-Method-Sets or sys.meth.set

Steps are to 
create a list of USIDS that are a part of the sys.meth.set
and then filter the model results to only include those USIDS
and then run the rank.summary() function on them.
Last step is to bind all of the results together

```{r}
agg.size <- as.character(unique(cal$agg.size))

for(i in agg.size){ # loop for every agg.size
  for(j in c("Ground", "Aerial")){ # loop for each method
    
    # create a USID filter for the agg.size-method Rank Set
    tempfilter1 <- cal %>% filter(agg.size == i,
                                 lpe_method == j)
    
    # create a SOURCE filter for only models that have the equivalent number of samples
    tempfilter2 <- mod.ranks %>% 
      filter(usid %in% tempfilter1$usid,
             rank.set.level <= 2) %>% 
      count(source) %>%  
      filter(n == max(n))
   
    # apply the rank.summary() function filtering the mod.ranks by the above filters
    rank.temp <- rank.summary(filter(mod.ranks,
                                     usid %in% tempfilter1$usid,
                                     source %in% tempfilter2$source))
    
    # Create a sys.meth.set or RANK.SET Name
    rank.temp <- rank.temp %>% 
      mutate(sys.meth.set = paste0("02_AggregateSize_General", " - ", i, " - ", j),
             rank.filter = i,
             rank.method = j,
             rank.set = "02_AggregateSize_General",
             special.level = 2)
    
    if(i == agg.size[1] & j == "Ground") rank.compare.agg.size <- rank.temp
    rank.compare.agg.size <- rbind(rank.compare.agg.size, rank.temp)
    
    rm(rank.temp)
    rm(tempfilter1)
    rm(tempfilter2)
  } # end of lpe_method loop
} # end of size.clarity loop

# removing the duplicates created from the first rbind
rank.compare.agg.size <- unique(rank.compare.agg.size)

# removing any rankings that have fewer than 10 samples
rank.compare.agg.size <- rank.compare.agg.size %>% filter(no.estimates >= sample.minimum)
```

## 03 Size
```{r}
size <- as.character(unique(cal$size))

for(i in size){ # loop for every size
  for(j in c("Ground", "Aerial")){ # loop for each method
    
    # create a USID filter for the size-method Rank Set
    tempfilter1 <- cal %>% filter(size == i,
                                 lpe_method == j)
    
    # create a SOURCE filter for only models that have the equivalent number of samples
    tempfilter2 <- mod.ranks %>% 
      filter(usid %in% tempfilter1$usid,
             rank.set.level <= 3) %>% 
      count(source) %>%  
      filter(n == max(n))
   
    # apply the rank.summary() function filtering the mod.ranks by the above filters
    rank.temp <- rank.summary(filter(mod.ranks,
                                     usid %in% tempfilter1$usid,
                                     source %in% tempfilter2$source))
    
    # Create a sys.meth.set or RANK.SET Name
    rank.temp <- rank.temp %>% 
      mutate(sys.meth.set = paste0("03_Size_Specific", " - ", i, " - ", j),
             rank.filter = i,
             rank.method = j,
             rank.set = "03_Size_Specific",
             special.level = 3)
    
    if(i == size[1] & j == "Ground") rank.compare.size <- rank.temp
    rank.compare.size <- rbind(rank.compare.size, rank.temp)
    
    rm(rank.temp)
    rm(tempfilter1)
    rm(tempfilter2)
  } # end of lpe_method loop
} # end of size.clarity loop

# removing the duplicates created from the first rbind
rank.compare.size <- unique(rank.compare.size)

# removing any rankings that have fewer than 10 samples
rank.compare.size <- rank.compare.size %>% filter(no.estimates >= sample.minimum)
```

## 04 Size-Water Clarity
```{r}
size.clarity <- as.character(unique(cal$size.clarity))

for(i in size.clarity){ # loop for every size.clarity
  for(j in c("Ground", "Aerial")){ # loop for each method
    # create a USID filter for the size.clarity-method Rank Set
    tempfilter1 <- cal %>% filter(size.clarity == i,
                                 lpe_method == j)
    
    # only run next chunks of code if tempfilter1 is not empty
    if(nrow(tempfilter1) >= 1){
    # create a SOURCE filter for only models that have the equivalent number of samples
    tempfilter2 <- mod.ranks %>% 
      filter(usid %in% tempfilter1$usid,
             rank.set.level <= 4) %>% 
      count(source) %>%  
      filter(n == max(n))
   
    # apply the rank.summary() function filtering the mod.ranks by the above filters
    rank.temp <- rank.summary(filter(mod.ranks,
                                     usid %in% tempfilter1$usid,
                                     source %in% tempfilter2$source))
    
    # Create a sys.meth.set or RANK.SET Name
    rank.temp <- rank.temp %>% 
      mutate(sys.meth.set = paste0("04_SizeClarity_Specific", " - ", i, " - ", j),
             rank.filter = i,
             rank.method = j,
             rank.set = "04_SizeClarity_Specific",
             special.level = 4)
    
    if(i == size.clarity[1] & j == "Ground") rank.compare.size.clarity <- rank.temp
    rank.compare.size.clarity <- rbind(rank.compare.size.clarity, rank.temp)
   
    rm(rank.temp)
    rm(tempfilter1)
    rm(tempfilter2)
   } # end of if statement
  } # end of lpe_method loop
} # end of size.clarity loop

# removing the duplicates created from the first rbind
rank.compare.size.clarity <- unique(rank.compare.size.clarity)

# removing any rankings that have fewer than 10 samples
rank.compare.size.clarity <- rank.compare.size.clarity %>% filter(no.estimates >= sample.minimum)
```

## 04.1 Clarity
```{r}
clarity <- as.character(unique(cal$water_clarity))

for(i in clarity){ # loop for every clarity
 
    # create a USID filter for the clarity-method Rank Set
    tempfilter1 <- cal %>% filter(water_clarity == i)
    
    # only run next chunks of code if tempfilter1 is not empty
    if(nrow(tempfilter1) >= 1){
    # create a SOURCE filter for only models that have the equivalent number of samples
    tempfilter2 <- mod.ranks %>% 
      filter(usid %in% tempfilter1$usid,
             rank.set.level <= 4) %>% 
      count(source) %>%  
      filter(n == max(n))
   
    # apply the rank.summary() function filtering the mod.ranks by the above filters
    rank.temp <- rank.summary(filter(mod.ranks,
                                     usid %in% tempfilter1$usid,
                                     source %in% tempfilter2$source))
    
    # Create a sys.meth.set or RANK.SET Name
    rank.temp <- rank.temp %>% 
      mutate(sys.meth.set = paste0("04.1_Clarity_Specific", " - ", i, "- Both"),
             rank.filter = i,
             rank.method = "Both",
             rank.set = "04.1_Clarity_Specific",
             special.level = 4.1)
    
    if(i == clarity[1]) rank.compare.clarity <- rank.temp
    rank.compare.clarity <- rbind(rank.compare.clarity, rank.temp)
   
    rm(rank.temp)
    rm(tempfilter1)
    rm(tempfilter2)
   } # end of if statement
  
} # end of clarity loop

# removing the duplicates created from the first rbind
rank.compare.clarity <- unique(rank.compare.clarity)

# removing any rankings that have fewer than 10 samples
rank.compare.clarity <- rank.compare.clarity %>% filter(no.estimates >= sample.minimum)
```



## 04.5 Specialized - General
This will be filtered out later but I wanted to still test the models as a whole
Won't be able to apply the results in a generalized sense.
```{r}
# pull out the list of rank.set.level = 5 for specialized that are not individual models
specialized.mods <- unique(mod.ranks$source[mod.ranks$rank.set.level == 6])

for(i in specialized.mods){ # loop for every size.clarity
  for(j in c("Ground", "Aerial")){ # loop for each method
    # create a USID filter for the specialized model
    tempfilter1 <- mod.ranks %>% filter(source == i,
                                 lpe_method == j)
    
    # only run next chunks of code if tempfilter1 is not empty
    if(nrow(tempfilter1) >= 1){
    # create a SOURCE filter for only models that have the equivalent number of samples
    tempfilter2 <- mod.ranks %>% 
      filter(usid %in% tempfilter1$usid,
             rank.set.level <= 6) %>% 
      count(source) %>%  
      filter(n == max(n))
   
    # apply the rank.summary() function filtering the mod.ranks by the above filters
    rank.temp <- rank.summary(filter(mod.ranks,
                                     usid %in% tempfilter1$usid,
                                     source %in% tempfilter2$source))
    
    # Create a sys.meth.set or RANK.SET Name
    rank.temp <- rank.temp %>% 
      mutate(sys.meth.set = paste0("04.5_Specialized", " - ", i, " - ", j),
             rank.filter = i,
             rank.method = j,
             rank.set = "04.5_Specialized_General",
             special.level = 4.5)
    
    if(i == specialized.mods[1] & j == "Ground") rank.compare.specialized.mods <- rank.temp
    rank.compare.specialized.mods <- rbind(rank.compare.specialized.mods, rank.temp)
   
    rm(rank.temp)
    rm(tempfilter1)
    rm(tempfilter2)
   } # end of if statement
  } # end of lpe_method loop
} # end of specialized.mods loop

# removing the duplicates created from the first rbind
rank.compare.specialized.mods <- unique(rank.compare.specialized.mods)

# removing any rankings that have fewer than 10 samples
rank.compare.specialized.mods <- rank.compare.specialized.mods %>% filter(no.estimates >= sample.minimum)
```

## 05 Individual - Simple
```{r}
# pull out the list of rank.set.level = 5 for specialized that are not individual models
individual.mods <- unique(mod.ranks$gazetted_stream_name[mod.ranks$rank.set.level == 5])

for(i in individual.mods){ # loop for every size.clarity
  for(j in c("Ground", "Aerial")){ # loop for each method
    # create a USID filter for the individual model
    tempfilter1 <- cal %>% filter(gazetted_stream_name == i,
                                 lpe_method == j)
    
    # only run next chunks of code if tempfilter1 is not empty
    if(nrow(tempfilter1) >= sample.minimum){
    # create a SOURCE filter for only models that have the equivalent number of samples
    tempfilter2 <- mod.ranks %>% 
      filter(usid %in% tempfilter1$usid,
             rank.set.level <= 5) %>% 
      count(source) %>%  
      filter(n == max(n))
   
    # apply the rank.summary() function filtering the mod.ranks by the above filters
    rank.temp <- rank.summary(filter(mod.ranks,
                                     usid %in% tempfilter1$usid,
                                     source %in% tempfilter2$source))
    
    # Create a sys.meth.set or RANK.SET Name
    rank.temp <- rank.temp %>% 
      mutate(sys.meth.set = paste0("05_Individual", " - ", i, " - ", j),
             rank.filter = i,
             rank.method = j,
             rank.set = "05_Individual_Specific",
             special.level = 5)
    
    if(i == individual.mods[1] & j == "Ground") rank.compare.individual.mods <- rank.temp
    rank.compare.individual.mods <- rbind(rank.compare.individual.mods, rank.temp)
   
    rm(rank.temp)
    rm(tempfilter1)
    rm(tempfilter2)
   } # end of if statement
  } # end of lpe_method loop
} # end of individual.mods loop

# removing the duplicates created from the first rbind
rank.compare.individual.mods <- unique(rank.compare.individual.mods)

# removing any rankings that have fewer than 10 samples
rank.compare.individual.mods <- rank.compare.individual.mods %>% filter(no.estimates >= sample.minimum)
```

## 06 Individual - Specialized
```{r}
# pull out the list of rank.set.level = 5 for specialized that are not individual models
specialized.mods <- unique(mod.ranks$source[mod.ranks$rank.set.level == 6])

for(i in specialized.mods){ # loop for every size.clarity
  for(j in c("Ground", "Aerial")){ # loop for each method
    # create a USID filter for the specialized model
    tempfilter0 <- mod.ranks %>% filter(source == i,
                                 lpe_method == j)
    
    temp.systems <- unique(tempfilter0$gazetted_stream_name) # list of systems to loop through
    
    for(k in temp.systems){ # loop for each system within the specialized model
      
      tempfilter1 <- mod.ranks %>% filter(source == i,
                                 lpe_method == j,
                            gazetted_stream_name == k)
    
    # only run next chunks of code if tempfilter1 is not empty
    if(nrow(tempfilter1) >= 1){
    # create a SOURCE filter for only models that have the equivalent number of samples
    tempfilter2 <- mod.ranks %>% 
      filter(usid %in% tempfilter1$usid,
             rank.set.level <= 6) %>% # up to 6 to include individual model if available
      count(source) %>%  
      filter(n == max(n))
   
    # apply the rank.summary() function filtering the mod.ranks by the above filters
    rank.temp <- rank.summary(filter(mod.ranks,
                                     usid %in% tempfilter1$usid,
                                     source %in% tempfilter2$source))
    
    # Create a sys.meth.set or RANK.SET Name
    rank.temp <- rank.temp %>% 
      mutate(sys.meth.set = paste0("06_", i, " - ", k, " - ", j),
             rank.filter = k,
             rank.method = j,
             rank.set = "06_Specialized_System",
             special.level = 6)
    
    if(i == specialized.mods[1] & j == "Ground" & k == "O' Ne-Ell Creek") rank.compare.specsys.mods <- rank.temp
    rank.compare.specsys.mods <- rbind(rank.compare.specsys.mods, rank.temp)
   
    rm(rank.temp)
    rm(tempfilter1)
    rm(tempfilter2)
      } # end of if statement
    } # end of temp.systems loop
  } # end of lpe_method loop
} # end of specialized.mods loop

# removing the duplicates created from the first rbind
rank.compare.specsys.mods <- unique(rank.compare.specsys.mods)

# removing any rankings that have fewer than minimum sample requirement
rank.compare.specsys.mods <- rank.compare.specsys.mods %>% filter(no.estimates >= sample.minimum)
```

# Rank Summary
Combining the rank sets
```{r}
# Generate the list of data frame names
rank.combined <- mget(ls(pattern = "^rank.compare\\."))

# Combine the data frames using rbind
rank.combined <- do.call(rbind, rank.combined)

rank.combined <- rank.combined %>% filter(special.level != 4.1)

saveRDS(rank.combined, file = "./r_objects/02_01_rank_combined_results.rds")
write.csv(rank.combined, file = "./tables/02_01-rank_combined_results.csv", row.names = F)
```

## Performance Metric Sensitivity
I wanted to test what set of performance Metrics may be "best" to use.
In this case, I plan to identify which metrics often result in the same model
selection or if there is a set of metrics that results in a different choice 
more often.
```{r}
# selecting just the identifier information sys.meth.set:rank.set, source
# and selecting the different metric-set results, the rel.sums
rank.method.compare <- rank.combined %>% 
  dplyr::select(sys.meth.set:rank.set, source, rel.sum:rel.sum.5) %>% 
  # pivoting the data into a long format for analysis below
  pivot_longer(cols = starts_with("rel"),
               names_to = "method", # Method is the name of the metric-sets
               values_to = "score")

# Analysis comparing the top results
rank.method.compare.tops <- rank.method.compare %>% 
  # within each Rank.Set, select the top scoring Source (Model) per method (metric-set)
  group_by(sys.meth.set, method) %>% 
  filter(score == min(score)) %>%  # filters to the top scoring per metric set
  # creating a count of the source per Rank.Set within the top scorers
  group_by(sys.meth.set, source) %>% 
  mutate(source.pair = n()) %>% 
  # within each Rank.Set, identifying "outliers"
  # outliers are defined here as not being the most arrived at source (model)
  group_by(sys.meth.set) %>% 
  mutate(outlier = ifelse(source.pair == max(source.pair), FALSE, TRUE))
# in this case, if each method had selected a different source, then none would 
# be flagged as an outlier because all would be == 1 which would be the max in 
# that scenario

# creating a ranked order of rel.sums (method) for plotting reasons 
rel.method <- unique(rank.method.compare.tops$method[order(rank.method.compare.tops$method)])
rank.method.compare.tops$method <- factor(rank.method.compare.tops$method, levels = rel.method)
# if they are ranked, then they will plot in order in each of the stacked bar
# plots that are below

# First Stacked Bar shows by each source that was selected
# is there a trend in which different source selected?
rank.method.compare.tops %>% 
  ggplot(aes(x = sys.meth.set,
              colour = method,
             fill = source)) +
  geom_bar(stat = "count") +
  coord_flip()

# Second stacked Bar highlights which method had outliers (or not)
rank.method.compare.tops %>% 
  ggplot(aes(x = sys.meth.set,
              colour = method,
             fill = outlier)) +
  geom_bar(stat = "count") +
  coord_flip()

# Calculating the proportion of outlying source selections each method produced
rank.method.compare.tops %>% 
  ungroup() %>% 
  count(method, outlier) %>% 
  group_by(method) %>% 
  mutate(pr.outlier = n / sum(n)) %>% 
  filter(outlier == FALSE) %>% 
  arrange(desc(pr.outlier))

# rel.Sum had the fewest number of outliers
# rel.sumMASE was second place
```

Selecting rel.sum, which uses abs.MPE, MAPE, RMSE, and r.cor

## Save Top Model Results
```{r}
rank.top.results <- rank.combined %>% 
  # for each rank.set
  group_by(sys.meth.set) %>% 
  # select the top ranked (lowest value) source (model)
  filter(rel.sum == min(rel.sum),
         rank.set != "04.5_Specialized_General") %>% 
  rename(top.model = source) %>% 
  mutate(is.n.robust = no.estimates >= 20,
         is.n.special.robust = no.estimates >= 10,
         is.cv.robust = mean.CV <= 50,
         is.special = special.level == rank.set.level,
         keep = case_when(
           is.n.robust == TRUE & is.cv.robust == TRUE ~ TRUE,
           is.n.special.robust == TRUE & is.cv.robust == TRUE & is.special == TRUE ~ TRUE,
           TRUE ~ FALSE
         ))

saveRDS(rank.top.results, file = "./r_objects/02_01_rank_top_results.rds")

rank.selected <- rank.top.results %>% 
  filter(keep == TRUE)
```

### Summary Plots
Just plotting top 5 models in each rank set to reduce the number of colours
```{r}
rank.combined <- rank.combined %>% 
  group_by(sys.meth.set) %>% 
  mutate(rank = rank(rel.sum))

rank.combined.reduced <- rank.combined %>% 
  filter(rank <= 5) %>% 
  ungroup()
```


Colourblind safe colours
https://stackoverflow.com/questions/57153428/r-plot-color-combinations-that-are-colorblind-accessible
```{r}
col.mods<-arrange(count(rank.combined.reduced, source), desc(n))

print(col.mods, n = 30)

col.mods$col = NA

col.mods$col = c("#DDCC77", "#CC6677", "black", "#888888", # 4
                 "#88CCEE", "#117733", "#332288", "#661100", # 8
                 "#44AA99", "#999933", "#882255","#AA4499",  # 12
                 "#DDCC10", "#CC6610", "#6699CC", "grey25", # 16
                 "#999975", "#332200", "#44AA50", "red", # 20
                 "grey") # 21

col.mods <- arrange(col.mods, source)
```


```{r}
# making a reference for looping and filtering within the loop
mod.ranks <- left_join(mod.ranks,
                       dplyr::select(cal, usid, size))

# making a list of rank sets to loop through to make figures for
sys.meth.set <- rank.combined.reduced %>% 
  dplyr::select(no.estimates, sys.meth.set:special.level) %>% 
  unique()

for(i in 1:nrow(sys.meth.set)){
  
  loop.dat <- rank.combined.reduced[rank.combined.reduced$sys.meth.set == as.character(sys.meth.set$sys.meth.set[i]), ]
  loop.river <- sys.meth.set$rank.filter[i] # loop.river = river or aggregate level
  loop.set <- sys.meth.set$rank.set[i] # loop model/rank set that is being plotted
  loop.method <- sys.meth.set$rank.method[i] # loop method
  loop.source <- loop.dat$source[loop.dat$rel.sum == min(loop.dat$rel.sum)] # top model source
  loop.sp.level <- sys.meth.set$special.level[i] # rank set special level, for later filtering step
  loop.cols <- col.mods$col[col.mods$source %in% loop.dat$source] # colours for plotting
  loop.keep <- ifelse(rank.top.results$keep[rank.top.results$sys.meth.set == as.character(sys.meth.set$sys.meth.set[i])] == TRUE,
                      " - Accepted", " - Rejected")
  
  # Paired down data to the model set data and top model fits
  loop.dat.top <- filter(mod.ranks,
                         lpe_method == loop.method, # filter down to method
                         source == loop.source) # filter to top model
  # need to further filter based on the aggregate, size, size.clarity, river or special set
  if(loop.sp.level == 2) loop.dat.top <- filter(loop.dat.top,
                                                agg.size == loop.river)
  if(loop.sp.level == 3) loop.dat.top <- filter(loop.dat.top,
                                                size == loop.river)
  if(loop.sp.level == 4) loop.dat.top <- filter(loop.dat.top,
                                                size.clarity == loop.river)
  if(loop.sp.level >= 5) loop.dat.top <- filter(loop.dat.top, # individual rivers
                                                gazetted_stream_name == loop.river)
  if(loop.sp.level == 4.5) { # 5 are the partial aggregate special sets.
    loop.usid <- mod.ranks %>% filter(source == loop.river)
    loop.dat.top <- filter(loop.dat.top, usid %in% loop.usid$usid)
  }
      
  # Relative Rank Sum
  pl1 <- ggplot(loop.dat,
         aes(x = reorder_within(source, rel.sum,
                                sys.meth.set, median), 
             y = rel.sum, fill = source)) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_x_reordered() +
    scale_fill_manual(values = loop.cols) +
    theme(axis.text.x = element_blank()) +
    ylim(c(0,1)) +
    labs(x = "", y = "Average \n Relative Model Rank \n (1=Worst, 0=Best)")
  
  # Mean Percent Error
  pl2 <- ggplot(loop.dat,
         aes(x = reorder_within(source, abs.mean.percent.error,
                                sys.meth.set, median), 
             y = mean.percent.error, fill = source)) +
    geom_bar(stat = "identity", position = "dodge") +
    geom_hline(yintercept = 0.15, col = "grey") +
    geom_hline(yintercept = -0.15, col = "grey") +
    geom_hline(yintercept = 0.50, col = "black") +
    geom_hline(yintercept = -0.50, col = "black") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    scale_x_reordered() +
    scale_fill_manual(values = loop.cols) +
    geom_hline(yintercept = 0) +
    theme(axis.text.x = element_blank()) +
    labs(x = "", y = "Mean Percent \n Error (MPE)")
  
  # Mean Absolute Percent Error
  pl3 <- ggplot(loop.dat,
         aes(x = reorder_within(source, mean.abs.per.error,
                                sys.meth.set, median), 
             y = mean.abs.per.error, fill = source)) +
    geom_bar(stat = "identity", position = "dodge") +
    geom_hline(yintercept = 0.15, col = "grey") +
    geom_hline(yintercept = 0.50, col = "black") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    scale_x_reordered() +
    scale_fill_manual(values = loop.cols) +
    theme(axis.text.x = element_blank()) +
    labs(x = "", y = "Mean Absolute Percent \n Error (MAPE)")
  
  # Root mean squared error
  pl4 <- ggplot(loop.dat,
         aes(x = reorder_within(source, root.mean.sqrd.error,
                                sys.meth.set, median), 
             y = root.mean.sqrd.error, fill = source)) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_x_reordered() +
    scale_fill_manual(values = loop.cols) +
    theme(axis.text.x = element_blank()) +
    labs(x = "", y = "Root Mean Squared \n Error (RMSE)")

  # R correlation coefficient
  pl6 <- ggplot(loop.dat,
         aes(x = reorder_within(source, -r.correlation,
                                sys.meth.set, median), 
             y = r.correlation, fill = source)) +
    geom_bar(stat = "identity", position = "dodge") +
    scale_x_reordered() +
    geom_hline(yintercept = 0.80, col = "grey") +
    geom_hline(yintercept = 1, col = "black") +
    scale_fill_manual(values = loop.cols) +
    theme(axis.text.x = element_blank()) +
    labs(x = "", y = "Correlation Coefficient \n (R)")
  
  # LPE vs. HAE and Fit graph
  pl7 <- ggplot(loop.dat.top,
                aes(x = classic/1.8, y = hpe)) +
    geom_point(alpha = 0.75) +
    geom_point(aes(y = calibrated.estimate), colour = "red", alpha = 0.60) +
    scale_x_continuous(labels = label_comma()) +
  scale_y_continuous(labels = label_comma()) +
    theme(axis.text.y = element_text(angle = 90, hjust = 0.5, vjust = 1),
          panel.grid.major = element_line(colour = "grey")) +
    #geom_vline(xintercept = 75000, alpha = 0.25, lty = 2) +
    labs(x = "Low Precision Estimate (LPE)",
         y = "High Accuracy Estimate \n (HAE-black) \n & Top Model Fit (Red)",
         title = paste0(loop.source, loop.keep))
  
  # combining the plots using patchwork
  (pl2 | pl3) / (pl4 | pl6) / (pl1) +
    plot_annotation(paste0("Model Performance Results for \n ", sys.meth.set$sys.meth.set[i], 
                                                   " Set [", sys.meth.set$no.estimates[i], "]"),
                    theme=theme(plot.title=element_text(hjust=0.5))) +
    plot_layout(heights = c(1,1,1),
                guides = 'collect') & 
      theme(legend.position = 'top') &
    guides(fill = guide_legend(nrow = 3))
  
  ggsave(paste0("./figures/02_Rank_Summaries/", 
                sys.meth.set$sys.meth.set[i], ".png"),
         height = 7.2, width = 8, dpi = 150)
  
}

```

```{r}

```

