---
title: "03-01-Model-Code"
author: "Michael Arbeider"
date: "2026-03-06"
---

# Set-up
## Packages
```{r}
rm(list = ls())
library(tidyverse) # most %>% 
library(scales) # for label_comma()
```

# Data
## Models
```{r}
# Combining all models into a list called mods
 # Data path folder for Model Ranks
r_objects <- "./r_objects/models"
 # Pull all the data files
data_path <- dir(r_objects, full.names = TRUE)
  # Obtain the file names that include .rds, which are the models themselves
fileNames <- data_path[grep(pattern = ".rds", data_path)]
 # Create a character string with the Model Names
individual.names <- sub(pattern = ".*models/", "", fileNames)
individual.names <- sub(pattern = ".rds", "", individual.names)
# remove the model_specifications from the individual names and filenames
individual.names <- individual.names[-which(individual.names == "model_specifications")]
fileNames <- fileNames[-grep(pattern = "model_specifications", fileNames)]


# prime blank list
mods <- list()

# looping through file names to read in RDS and create a "source" field based on the file name
for(i in 1:length(fileNames)){
  mods[[i]] <- readRDS(fileNames[i])
  names(mods)[i] <- individual.names[i]
  }
```
### Model Specs Functions
a series of functions for each model type to apply the correct data input
transformations, model, and fit transformations (either due to a link function
or other transformation, e.g. Estu model applied a cube root to hpe when the 
model was fit).
Using mod.specs to identify what transformations to make for each function.
```{r}
# model specifications
mod.specs <- readRDS("./r_objects/models/model_specifications.rds")
unique(mod.specs$model)

# blank list to hold functions that apply models and data transformations
mod.spec.fn <- list()
```
#### All_Clarity_Model
```{r}
mod.spec.fn$All_Clarity_Model <- function(datin){
  # apply data input transformations
  datin$lnlpe.sc = (log(datin$lpe) - 
                      mod.specs$transform[mod.specs$model == "All_Clarity_Model" & mod.specs$type == "center"]) / 
    mod.specs$transform[mod.specs$model == "All_Clarity_Model" & mod.specs$type == "scale"]
  
  # Force Substrate Shade to be Medium if it is Dark when size is small and the method is ground
  if(datin$size == "Small" & datin$lpe_method == "Ground" & datin$substrate_shade == "Dark") datin$substrate_shade = "Medium"
  
  # apply model and get predictions (preds)
  preds <- as.data.frame(predict(mods[["All_Clarity_Model"]], type="link", 
                   dplyr::select(datin, -gazetted_stream_name),
                      re.form=NA, se.fit=T))
  # apply data transformations to predictions (calibrated.estimates) to get error bounds.
  preds <- mutate(preds,
                fit.cv = round2(sqrt(exp(se.fit^2)-1)*100,1),
                calibrated.se.lower = exp(fit-se.fit),
                calibrated.se.upper = exp(fit+se.fit),
                calibrated.estimate = exp(fit))
  # attach results to data input file
  
  datin <- mutate(datin,
                  fit = preds$fit,
                  se.fit = preds$se.fit,
                  fit.cv = preds$fit.cv,
                  calibrated.se.lower = preds$calibrated.se.lower,
                  calibrated.se.upper = preds$calibrated.se.upper,
                  calibrated.estimate = preds$calibrated.estimate)
  # return the data input file with predictions attached.
  return(datin)
}
# annotations above are not repeated for each model function.
```
#### SM_Data_Mode
```{r}
mod.spec.fn$SM_Data_Model <- function(datin){
  datin$lnlpe.sc = (log(datin$lpe) - 
                      mod.specs$transform[mod.specs$model == "SM_Data_Model" & mod.specs$type == "center"]) / 
    mod.specs$transform[mod.specs$model == "SM_Data_Model" & mod.specs$type == "scale"]
  
  preds <- as.data.frame(predict(mods[["SM_Data_Model"]], type="link", 
                   dplyr::select(datin, -gazetted_stream_name),
                      re.form=NA, se.fit=T))
  
  preds <- mutate(preds,
                fit.cv = round2(sqrt(exp(se.fit^2)-1)*100,1),
                calibrated.se.lower = exp(fit-se.fit),
                calibrated.se.upper = exp(fit+se.fit),
                calibrated.estimate = exp(fit))
  
  datin <- mutate(datin,                   
                  fit = preds$fit,                   
                  se.fit = preds$se.fit,                   
                  fit.cv = preds$fit.cv,                   
                  calibrated.se.lower = preds$calibrated.se.lower,                   
                  calibrated.se.upper = preds$calibrated.se.upper,                   
                  calibrated.estimate = preds$calibrated.estimate)
  
  return(datin)
}
```
#### LXL_Data_Model
```{r}
mod.spec.fn$LXL_Data_Model <- function(datin){
  datin$lnlpe.sc = (log(datin$lpe) - 
                      mod.specs$transform[mod.specs$model == "LXL_Data_Model" & mod.specs$type == "center"]) / 
    mod.specs$transform[mod.specs$model == "LXL_Data_Model" & mod.specs$type == "scale"]
  
  preds <- as.data.frame(predict(mods[["LXL_Data_Model"]], type="link", 
                   dplyr::select(datin, -gazetted_stream_name),
                      re.form=NA, se.fit=T))
  
  preds <- mutate(preds,
                fit.cv = round2(sqrt(exp(se.fit^2)-1)*100,1),
                calibrated.se.lower = exp(fit-se.fit),
                calibrated.se.upper = exp(fit+se.fit),
                calibrated.estimate = exp(fit))
  
  datin <- mutate(datin,                   
                  fit = preds$fit,                   
                  se.fit = preds$se.fit,                   
                  fit.cv = preds$fit.cv,                   
                  calibrated.se.lower = preds$calibrated.se.lower,                   
                  calibrated.se.upper = preds$calibrated.se.upper,                   
                  calibrated.estimate = preds$calibrated.estimate)
  
  return(datin)
}
```
#### Early_Stu_DeadAlive_Model
```{r}
mod.spec.fn$Early_Stu_DeadAlive_Model <- function(datin){
  datin$Cubelpe.sc = ((datin$lpe)^(1/3) - 
                      mod.specs$transform[mod.specs$model == "Early_Stu_DeadAlive_Model" & 
                                            mod.specs$type == "center" & mod.specs$target == "lpe"]) / 
    mod.specs$transform[mod.specs$model == "Early_Stu_DeadAlive_Model" & 
                          mod.specs$type == "scale" & mod.specs$target == "lpe"]
  datin$Cubedeadvlive.sc = ( (datin$ground_carcs_total_above/datin$peak_live_above)^(1/3) - 
                      mod.specs$transform[mod.specs$model == "Early_Stu_DeadAlive_Model" & 
                                            mod.specs$type == "center" & mod.specs$target == "deadvlive"]) / 
    mod.specs$transform[mod.specs$model == "Early_Stu_DeadAlive_Model" & 
                          mod.specs$type == "scale" & mod.specs$target == "deadvlive"]
  
  preds <- as.data.frame(predict(mods[["Early_Stu_DeadAlive_Model"]], type="response", 
                   dplyr::select(datin, -gazetted_stream_name),
                      re.form=NA, se.fit=T))
  
  preds <- mutate(preds,
                fit.cv = round2(sqrt(exp(se.fit^2)-1)*100,1),
                calibrated.se.lower = (fit-se.fit)^3,
                calibrated.se.upper = (fit+se.fit)^3,
                calibrated.estimate = fit^3)
  
  datin <- mutate(datin,                   
                  fit = preds$fit,                   
                  se.fit = preds$se.fit,                   
                  fit.cv = preds$fit.cv,                   
                  calibrated.se.lower = preds$calibrated.se.lower,                   
                  calibrated.se.upper = preds$calibrated.se.upper,                   
                  calibrated.estimate = preds$calibrated.estimate)
  
  return(datin)
}
```

#### All_Systems_Discharge_Model
```{r}
# HERE #
mod.spec.fn$All_Systems_Discharge_Model <- function(datin){
  center <- mod.specs$transform[mod.specs$model == "All_Systems_Discharge_Model" &
                                 mod.specs$target == paste(datin$gazetted_stream_name, datin$run_timing_group, sep = "-") &
                                mod.specs$type == "mean"]
  sd <- mod.specs$transform[mod.specs$model == "All_Systems_Discharge_Model" &
                                 mod.specs$target == paste(datin$gazetted_stream_name, datin$run_timing_group, sep = "-") &
                                mod.specs$type == "sd"]
  
  datin$ln.lpe.sc = (log(datin$lpe) - 
                      mod.specs$transform[mod.specs$model == "All_Systems_Discharge_Model" & 
                                            mod.specs$type == "center" & mod.specs$target == "lpe"]) / 
    mod.specs$transform[mod.specs$model == "All_Systems_Discharge_Model" & 
                          mod.specs$type == "scale" & mod.specs$target == "lpe"]
  datin$discharge.sc = ( datin$water_discharge - center ) / sd
  
  preds <- as.data.frame(predict(mods[["All_Systems_Discharge_Model"]], type="link", 
                   dplyr::select(datin, -gazetted_stream_name),
                      re.form=NA, se.fit=T))
  
  preds <- mutate(preds,
                fit.cv = round2(sqrt(exp(se.fit^2)-1)*100,1),
                calibrated.se.lower = exp(fit-se.fit),
                calibrated.se.upper = exp(fit+se.fit),
                calibrated.estimate = exp(fit))
  
  datin <- mutate(datin,                   
                  fit = preds$fit,                   
                  se.fit = preds$se.fit,                   
                  fit.cv = preds$fit.cv,                   
                  calibrated.se.lower = preds$calibrated.se.lower,                   
                  calibrated.se.upper = preds$calibrated.se.upper,                   
                  calibrated.estimate = preds$calibrated.estimate)
  
  return(datin)
}
```
#### Stellako_Discharge_Model
```{r}
mod.spec.fn$Stellako_Discharge_Model <- function(datin){
  datin$lnlpe.sc = ( log(datin$lpe) - 
                      mod.specs$transform[mod.specs$model == "Stellako_Discharge_Model" & 
                                            mod.specs$type == "center" & mod.specs$target == "lpe"]) / 
    mod.specs$transform[mod.specs$model == "Stellako_Discharge_Model" & 
                          mod.specs$type == "scale" & mod.specs$target == "lpe"]
  datin$discharge.sc = ( datin$water_discharge - 
                      mod.specs$transform[mod.specs$model == "Stellako_Discharge_Model" & 
                                            mod.specs$type == "center" & mod.specs$target == "discharge"]) /
    mod.specs$transform[mod.specs$model == "Stellako_Discharge_Model" & 
                          mod.specs$type == "scale" & mod.specs$target == "discharge"]
  
  preds <- as.data.frame(predict(mods[["Stellako_Discharge_Model"]], type="link", 
                   dplyr::select(datin, -gazetted_stream_name),
                      re.form=NA, se.fit=T))
  
  preds <- mutate(preds,
                fit.cv = round2(sqrt(exp(se.fit^2)-1)*100,1),
                calibrated.se.lower = exp(fit-se.fit),
                calibrated.se.upper = exp(fit+se.fit),
                calibrated.estimate = exp(fit))
  
  datin <- mutate(datin,                   
                  fit = preds$fit,                   
                  se.fit = preds$se.fit,                   
                  fit.cv = preds$fit.cv,                   
                  calibrated.se.lower = preds$calibrated.se.lower,                   
                  calibrated.se.upper = preds$calibrated.se.upper,                   
                  calibrated.estimate = preds$calibrated.estimate)
  
  return(datin)
}
```
#### Individual_Data_Model
```{r}
mod.spec.fn$Individual_Data_Model <- function(datin, mod.i){
  # apply data input transformations
  datin$lnlpe.sc = (log(datin$lpe))
  
  # apply model and get predictions (preds)
  preds <- as.data.frame(predict(mods[[mod.i]], type="link", 
                   dplyr::select(datin, -gazetted_stream_name),
                      re.form=NA, se.fit=T))
  # apply data transformations to predictions (calibrated.estimates) to get error bounds.
  preds <- mutate(preds,
                fit.cv = round2(sqrt(exp(se.fit^2)-1)*100,1),
                calibrated.se.lower = exp(fit-se.fit),
                calibrated.se.upper = exp(fit+se.fit),
                calibrated.estimate = exp(fit))
  # attach results to data input file
  datin <- mutate(datin,                   
                  fit = preds$fit,                   
                  se.fit = preds$se.fit,                   
                  fit.cv = preds$fit.cv,                   
                  calibrated.se.lower = preds$calibrated.se.lower,                   
                  calibrated.se.upper = preds$calibrated.se.upper,                   
                  calibrated.estimate = preds$calibrated.estimate)
  # return the data input file with predictions attached.
  return(datin)
}
```
#### All_Minimal_Model
```{r}
mod.spec.fn$All_Minimal_Model <- function(datin){
  # apply data input transformations
  datin$lnlpe.sc = (log(datin$lpe) - 
                      mod.specs$transform[mod.specs$model == "All_Minimal_Model" & mod.specs$type == "center"]) / 
    mod.specs$transform[mod.specs$model == "All_Minimal_Model" & mod.specs$type == "scale"]
  
  # apply model and get predictions (preds)
  preds <- as.data.frame(predict(mods[["All_Minimal_Model"]], type="link", 
                   dplyr::select(datin, -gazetted_stream_name),
                      re.form=NA, se.fit=T))
  # apply data transformations to predictions (calibrated.estimates) to get error bounds.
  preds <- mutate(preds,
                fit.cv = round2(sqrt(exp(se.fit^2)-1)*100,1),
                calibrated.se.lower = exp(fit-se.fit),
                calibrated.se.upper = exp(fit+se.fit),
                calibrated.estimate = exp(fit))
  # attach results to data input file
  datin <- mutate(datin,                   
                  fit = preds$fit,                   
                  se.fit = preds$se.fit,                   
                  fit.cv = preds$fit.cv,                   
                  calibrated.se.lower = preds$calibrated.se.lower,                   
                  calibrated.se.upper = preds$calibrated.se.upper,                   
                  calibrated.estimate = preds$calibrated.estimate)
  # return the data input file with predictions attached.
  return(datin)
}
```

#### All_ROM_GenB_Model
```{r}
mod.spec.fn$All_ROM_GenB_Model <- function(datin){
  # apply data input transformations
 datin$lnlpe = log(datin$lpe)
   # apply model and get predictions (preds)
  preds <- data.frame(fit = mods[["All_ROM_GenB_Model"]]$rom*datin$lnlpe,
                      se.fit = datin$lnlpe*mods[["All_ROM_GenB_Model"]]$se)
  # apply data transformations to predictions (calibrated.estimates) to get error bounds.
  preds <- mutate(preds,
                  fit.cv = round2(sqrt(exp(se.fit^2)-1)*100,1),
                  calibrated.se.lower = exp(fit-se.fit),
                calibrated.se.upper = exp(fit+se.fit),
                calibrated.estimate = exp(fit))
  # attach results to data input file
  datin <- mutate(datin,                   
                  fit = preds$fit,                   
                  se.fit = preds$se.fit,                   
                  fit.cv = preds$fit.cv,                   
                  calibrated.se.lower = preds$calibrated.se.lower,                   
                  calibrated.se.upper = preds$calibrated.se.upper,                   
                  calibrated.estimate = preds$calibrated.estimate)
  # return the data input file with predictions attached.
  return(datin)
}
```

#### All_ROM_Gen_Model
```{r}
mod.spec.fn$All_ROM_Gen_Model <- function(datin){
  # apply data input transformations
  datin$lnlpe = log(datin$lpe)
 grouping <- paste("All", datin$lpe_method, sep = ".")
 grouping.row = which(mods[["All_ROM_Gen_Model"]]$grouping == grouping)
   # apply model and get predictions (preds)
  preds <- data.frame(fit = mods[["All_ROM_Gen_Model"]]$rom[grouping.row]*datin$lnlpe,
                      se.fit = datin$lnlpe*mods[["All_ROM_Gen_Model"]]$se[grouping.row])
  # apply data transformations to predictions (calibrated.estimates) to get error bounds.
  preds <- mutate(preds,
                  fit.cv = round2(sqrt(exp(se.fit^2)-1)*100,1),
                  calibrated.se.lower = exp(fit-se.fit),
                calibrated.se.upper = exp(fit+se.fit),
                calibrated.estimate = exp(fit))
  # attach results to data input file
  datin <- mutate(datin,                   
                  fit = preds$fit,                   
                  se.fit = preds$se.fit,                   
                  fit.cv = preds$fit.cv,                   
                  calibrated.se.lower = preds$calibrated.se.lower,                   
                  calibrated.se.upper = preds$calibrated.se.upper,                   
                  calibrated.estimate = preds$calibrated.estimate)
  # return the data input file with predictions attached.
  return(datin)
}
```

#### All_ROM_Agg_Model
```{r}
mod.spec.fn$All_ROM_Agg_Model <- function(datin){
  # apply data input transformations
 grouping <- paste(datin$aggr, datin$lpe_method, sep = ".")
 grouping.row = which(mods[["All_ROM_Agg_Model"]]$grouping == grouping)
 datin$lnlpe = log(datin$lpe)
   # apply model and get predictions (preds)
  preds <- data.frame(fit = mods[["All_ROM_Agg_Model"]]$rom[grouping.row]*datin$lnlpe,
                      se.fit = datin$lnlpe*mods[["All_ROM_Agg_Model"]]$se[grouping.row])
  # apply data transformations to predictions (calibrated.estimates) to get error bounds.
  preds <- mutate(preds,
                  se.fit = se.fit, # placeholder need to EDIT
                fit.cv = round2(sqrt(exp(se.fit^2)-1)*100,1),
                calibrated.se.lower = exp(fit-se.fit),
                calibrated.se.upper = exp(fit+se.fit),
                calibrated.estimate = exp(fit))
  # attach results to data input file
    datin <- mutate(datin,
                  fit = preds$fit,
                  se.fit = preds$se.fit,
                  fit.cv = preds$fit.cv,
                  calibrated.se.lower = preds$calibrated.se.lower,
                  calibrated.se.upper = preds$calibrated.se.upper,
                  calibrated.estimate = preds$calibrated.estimate)
  # return the data input file with predictions attached.
  return(datin)
}
```

#### All_ROM_Size_Model
```{r}
mod.spec.fn$All_ROM_Size_Model <- function(datin){
  # apply data input transformations
 grouping <- paste(datin$size, datin$lpe_method, sep = ".")
   grouping.row = which(mods[["All_ROM_Size_Model"]]$grouping == grouping)
   datin$lnlpe = log(datin$lpe)
   # apply model and get predictions (preds)
  preds <- data.frame(fit = mods[["All_ROM_Size_Model"]]$rom[grouping.row]*datin$lnlpe,
                      se.fit = datin$lnlpe*mods[["All_ROM_Size_Model"]]$se[grouping.row])
  # apply data transformations to predictions (calibrated.estimates) to get error bounds.
  preds <- mutate(preds,
                  se.fit = se.fit, # placeholder need to EDIT
                fit.cv = round2(sqrt(exp(se.fit^2)-1)*100,1),
                calibrated.se.lower = exp(fit-se.fit),
                calibrated.se.upper = exp(fit+se.fit),
                calibrated.estimate = exp(fit))
  # attach results to data input file
  datin <- datin <- mutate(datin,
                  fit = preds$fit,
                  se.fit = preds$se.fit,
                  fit.cv = preds$fit.cv,
                  calibrated.se.lower = preds$calibrated.se.lower,
                  calibrated.se.upper = preds$calibrated.se.upper,
                  calibrated.estimate = preds$calibrated.estimate)
  # return the data input file with predictions attached.
  return(datin)
}
```

#### MoR_LM_Clarity_Model
```{r}
mod.spec.fn$MoR_LM_Clarity_Model <- function(datin){
  # apply model and get predictions (preds)
  # In this case, it is predicting the ratio
  datin$lnlpe = log(datin$lpe)
  preds <- as.data.frame(predict(mods[["MoR_LM_Clarity_Model"]], 
                      datin,
                      re.form=NA, se.fit=T))
  # apply data transformations to predictions (calibrated.estimates) to get error bounds.
  preds <- mutate(preds,
                  fit = fit*datin$lpe,
                  se.fit = se.fit*datin$lpe,
                fit.cv = round2(sqrt(exp(se.fit^2)-1)*100,1),
                calibrated.se.lower = (fit-se.fit),
                calibrated.se.upper = (fit+se.fit),
                calibrated.estimate = (fit))
  # attach results to data input file
  datin <- mutate(datin,
                  fit = preds$fit,
                  se.fit = preds$se.fit,
                  fit.cv = preds$fit.cv,
                  calibrated.se.lower = preds$calibrated.se.lower,
                  calibrated.se.upper = preds$calibrated.se.upper,
                  calibrated.estimate = preds$calibrated.estimate)
  # return the data input file with predictions attached.
  return(datin)
}
```

#### MoR_Clarity_Model
```{r}
mod.spec.fn$MoR_Clarity_Mode <- function(datin){
  # apply data input transformations
 # apply data input transformations
 datin$lnlpe = log(datin$lpe)
  modf <- mods[["MoR_Clarity_Model"]] %>% 
    dplyr::filter(lpe_method == datin$lpe_method,
                  size == datin$size,
                  water_clarity == datin$water_clarity)
  
     # apply model and get predictions (preds)
  preds <- data.frame(fit = modf$MoR*datin$lpe,
                      se.fit = modf$MoR.SD*datin$lpe)
  
  # apply data transformations to predictions (calibrated.estimates) to get error bounds.
  preds <- mutate(preds,
                  se.fit = se.fit,
                fit.cv = round2(se.fit / fit * 100, 1),
                calibrated.se.lower = (fit-se.fit),
                calibrated.se.upper = (fit+se.fit),
                calibrated.estimate = (fit))
  # attach results to data input file
  datin <- mutate(datin,                   
                  fit = preds$fit,                   
                  se.fit = preds$se.fit,                   
                  fit.cv = preds$fit.cv,                   
                  calibrated.se.lower = preds$calibrated.se.lower,                   
                  calibrated.se.upper = preds$calibrated.se.upper,                   
                  calibrated.estimate = preds$calibrated.estimate)

  # return the data input file with predictions attached.
  return(datin)
}
```

#### MoR_Low_Abundance_Models
```{r}
mod.spec.fn$MoR_Low_Abundance_Models <- function(datin){
  # apply data input transformations
 # apply data input transformations
 datin$lnlpe = log(datin$lpe)
  modf <- mods[["MoR_Low_Abundance_Models"]] %>% 
    dplyr::filter(lpe_method == datin$lpe_method,
                  water_clarity == datin$water_clarity)
  
     # apply model and get predictions (preds)
  preds <- data.frame(fit = modf$MoR*datin$lpe,
                      se.fit = modf$MoR.SD*datin$lpe)
  
  # apply data transformations to predictions (calibrated.estimates) to get error bounds.
  preds <- mutate(preds,
                  se.fit = se.fit,
                fit.cv = round2(se.fit / fit * 100, 1),
                calibrated.se.lower = (fit-se.fit),
                calibrated.se.upper = (fit+se.fit),
                calibrated.estimate = (fit))
  # attach results to data input file
  datin <- mutate(datin,                   
                  fit = preds$fit,                   
                  se.fit = preds$se.fit,                   
                  fit.cv = preds$fit.cv,                   
                  calibrated.se.lower = preds$calibrated.se.lower,                   
                  calibrated.se.upper = preds$calibrated.se.upper,                   
                  calibrated.estimate = preds$calibrated.estimate)

  # return the data input file with predictions attached.
  return(datin)
}
```

#### RoM_Clarity_Model
```{r}
mod.spec.fn$RoM_Clarity_Model <- function(datin){
  # apply data input transformations
 # apply data input transformations
 datin$lnlpe = log(datin$lpe)
  modf <- mods[["RoM_Clarity_Model"]] %>% 
    dplyr::filter(lpe_method == datin$lpe_method,
                  size == datin$size,
                  water_clarity == datin$water_clarity)
  
     # apply model and get predictions (preds)
  preds <- data.frame(fit = modf$RoM*datin$lpe,
                      se.fit = modf$RoM.SD)
  
  # apply data transformations to predictions (calibrated.estimates) to get error bounds.
  preds <- mutate(preds,
                  se.fit = se.fit, # placeholder need to EDIT
                fit.cv = round2(se.fit / fit * 100, 1),
                calibrated.se.lower = (fit-se.fit),
                calibrated.se.upper = (fit+se.fit),
                calibrated.estimate = (fit))
  # attach results to data input file
  datin <- mutate(datin,                   
                  fit = preds$fit,                   
                  se.fit = preds$se.fit,                   
                  fit.cv = preds$fit.cv,                   
                  calibrated.se.lower = preds$calibrated.se.lower,                   
                  calibrated.se.upper = preds$calibrated.se.upper,                   
                  calibrated.estimate = preds$calibrated.estimate)

  # return the data input file with predictions attached.
  return(datin)
}
```
#### Inverse Models
```{r}
cal_4apply <- function(y0, # Low Precision Estimate
                       mod, # simple linear model 
                       mean.res=FALSE, interval="Wald")
  # Apply calibration function from INVESTR PACKAGE
{
  x0 <- investr::calibrate(mod,y0,interval=interval, mean.response = mean.res)
  return(x0[1:4]) # return the High Precision estimate, confidence interval, and Standard Error
}

saveRDS(cal_4apply, "./r_objects/03_cal_4apply.rds")
```
##### Inverse_LXL_Aerial_Data_Model
```{r}
mod.spec.fn$Inverse_LXL_Aerial_Data_Model <- function(datin){
  # apply data input transformations
  datin$lnlpe = (log(datin$lpe))
  
  # apply model and get predictions (preds)
  m.temp <- mods$Inverse_LXL_Aerial_Data_Model
    
    p.temp <- matrix(unlist(t(sapply(cal.noL$lnlpe[i], cal_4apply, mod=m.temp))), ncol = 4)
    p.temp <- as.data.frame( p.temp )
    colnames(p.temp) <- c("fit", "lower", "upper", "se.fit")
    
    preds <- data.frame(fit = p.temp$fit,
                        se.fit = p.temp$se.fit)

  # apply data transformations to predictions (calibrated.estimates) to get error bounds.
  preds <- mutate(preds,
                fit.cv = round2(sqrt(exp(se.fit^2)-1)*100,1),
                calibrated.se.lower = exp(fit-se.fit),
                calibrated.se.upper = exp(fit+se.fit),
                calibrated.estimate = exp(fit))
  # attach results to data input file
  datin <- mutate(datin,                   
                  fit = preds$fit,                   
                  se.fit = preds$se.fit,                   
                  fit.cv = preds$fit.cv,                   
                  calibrated.se.lower = preds$calibrated.se.lower,                   
                  calibrated.se.upper = preds$calibrated.se.upper,                   
                  calibrated.estimate = preds$calibrated.estimate)
  # return the data input file with predictions attached.
  return(datin)
}
```

##### Inverse_LXL_Both_Data_Model
```{r}
mod.spec.fn$Inverse_LXL_Both_Data_Model <- function(datin){
  # apply data input transformations
  datin$lnlpe = (log(datin$lpe))
  
  # apply model and get predictions (preds)
  m.temp <- mods$Inverse_LXL_Both_Data_Model
    
    p.temp <- matrix(unlist(t(sapply(cal.noL$lnlpe[i], cal_4apply, mod=m.temp))), ncol = 4)
    p.temp <- as.data.frame( p.temp )
    colnames(p.temp) <- c("fit", "lower", "upper", "se.fit")
    
    preds <- data.frame(fit = p.temp$fit,
                        se.fit = p.temp$se.fit)

  # apply data transformations to predictions (calibrated.estimates) to get error bounds.
  preds <- mutate(preds,
                fit.cv = round2(sqrt(exp(se.fit^2)-1)*100,1),
                calibrated.se.lower = exp(fit-se.fit),
                calibrated.se.upper = exp(fit+se.fit),
                calibrated.estimate = exp(fit))
  # attach results to data input file
  datin <- mutate(datin,                   
                  fit = preds$fit,                   
                  se.fit = preds$se.fit,                   
                  fit.cv = preds$fit.cv,                   
                  calibrated.se.lower = preds$calibrated.se.lower,                   
                  calibrated.se.upper = preds$calibrated.se.upper,                   
                  calibrated.estimate = preds$calibrated.estimate)
  # return the data input file with predictions attached.
  return(datin)
}
```

##### Inverse_LXL_Ground_Data_Model
```{r}
mod.spec.fn$Inverse_LXL_Ground_Data_Model <- function(datin){
  # apply data input transformations
  datin$lnlpe = (log(datin$lpe))
  
  # apply model and get predictions (preds)
  m.temp <- mods$Inverse_LXL_Ground_Data_Model
    
    p.temp <- matrix(unlist(t(sapply(cal.noL$lnlpe[i], cal_4apply, mod=m.temp))), ncol = 4)
    p.temp <- as.data.frame( p.temp )
    colnames(p.temp) <- c("fit", "lower", "upper", "se.fit")
    
    preds <- data.frame(fit = p.temp$fit,
                        se.fit = p.temp$se.fit)

  # apply data transformations to predictions (calibrated.estimates) to get error bounds.
  preds <- mutate(preds,
                fit.cv = round2(sqrt(exp(se.fit^2)-1)*100,1),
                calibrated.se.lower = exp(fit-se.fit),
                calibrated.se.upper = exp(fit+se.fit),
                calibrated.estimate = exp(fit))
  # attach results to data input file
  datin <- mutate(datin,                   
                  fit = preds$fit,                   
                  se.fit = preds$se.fit,                   
                  fit.cv = preds$fit.cv,                   
                  calibrated.se.lower = preds$calibrated.se.lower,                   
                  calibrated.se.upper = preds$calibrated.se.upper,                   
                  calibrated.estimate = preds$calibrated.estimate)
  # return the data input file with predictions attached.
  return(datin)
}
```

##### Inverse_SM_Aerial_Data_Model
```{r}
mod.spec.fn$Inverse_SM_Aerial_Data_Model <- function(datin){
  # apply data input transformations
  datin$lnlpe = (log(datin$lpe))
  
  # apply model and get predictions (preds)
  m.temp <- mods$Inverse_SM_Aerial_Data_Model
    
    p.temp <- matrix(unlist(t(sapply(cal.noL$lnlpe[i], cal_4apply, mod=m.temp))), ncol = 4)
    p.temp <- as.data.frame( p.temp )
    colnames(p.temp) <- c("fit", "lower", "upper", "se.fit")
    
    preds <- data.frame(fit = p.temp$fit,
                        se.fit = p.temp$se.fit)

  # apply data transformations to predictions (calibrated.estimates) to get error bounds.
  preds <- mutate(preds,
                fit.cv = round2(sqrt(exp(se.fit^2)-1)*100,1),
                calibrated.se.lower = exp(fit-se.fit),
                calibrated.se.upper = exp(fit+se.fit),
                calibrated.estimate = exp(fit))
  # attach results to data input file
  datin <- mutate(datin,                   
                  fit = preds$fit,                   
                  se.fit = preds$se.fit,                   
                  fit.cv = preds$fit.cv,                   
                  calibrated.se.lower = preds$calibrated.se.lower,                   
                  calibrated.se.upper = preds$calibrated.se.upper,                   
                  calibrated.estimate = preds$calibrated.estimate)
  # return the data input file with predictions attached.
  return(datin)
}
```

##### Inverse_SM_Both_Data_Model
```{r}
mod.spec.fn$Inverse_SM_Both_Data_Model <- function(datin){
  # apply data input transformations
  datin$lnlpe = (log(datin$lpe))
  
  # apply model and get predictions (preds)
  m.temp <- mods$Inverse_SM_Both_Data_Model
    
    p.temp <- matrix(unlist(t(sapply(cal.noL$lnlpe[i], cal_4apply, mod=m.temp))), ncol = 4)
    p.temp <- as.data.frame( p.temp )
    colnames(p.temp) <- c("fit", "lower", "upper", "se.fit")
    
    preds <- data.frame(fit = p.temp$fit,
                        se.fit = p.temp$se.fit)

  # apply data transformations to predictions (calibrated.estimates) to get error bounds.
  preds <- mutate(preds,
                fit.cv = round2(sqrt(exp(se.fit^2)-1)*100,1),
                calibrated.se.lower = exp(fit-se.fit),
                calibrated.se.upper = exp(fit+se.fit),
                calibrated.estimate = exp(fit))
  # attach results to data input file
  datin <- mutate(datin,                   
                  fit = preds$fit,                   
                  se.fit = preds$se.fit,                   
                  fit.cv = preds$fit.cv,                   
                  calibrated.se.lower = preds$calibrated.se.lower,                   
                  calibrated.se.upper = preds$calibrated.se.upper,                   
                  calibrated.estimate = preds$calibrated.estimate)
  # return the data input file with predictions attached.
  return(datin)
}
```

##### Inverse_SM_Ground_Data_Model
```{r}
mod.spec.fn$Inverse_SM_Ground_Data_Model <- function(datin){
  # apply data input transformations
  datin$lnlpe = (log(datin$lpe))
  
  # apply model and get predictions (preds)
  m.temp <- mods$Inverse_SM_Ground_Data_Model
    
    p.temp <- matrix(unlist(t(sapply(cal.noL$lnlpe[i], cal_4apply, mod=m.temp))), ncol = 4)
    p.temp <- as.data.frame( p.temp )
    colnames(p.temp) <- c("fit", "lower", "upper", "se.fit")
    
    preds <- data.frame(fit = p.temp$fit,
                        se.fit = p.temp$se.fit)

  # apply data transformations to predictions (calibrated.estimates) to get error bounds.
  preds <- mutate(preds,
                fit.cv = round2(sqrt(exp(se.fit^2)-1)*100,1),
                calibrated.se.lower = exp(fit-se.fit),
                calibrated.se.upper = exp(fit+se.fit),
                calibrated.estimate = exp(fit))
  # attach results to data input file
  datin <- mutate(datin,                   
                  fit = preds$fit,                   
                  se.fit = preds$se.fit,                   
                  fit.cv = preds$fit.cv,                   
                  calibrated.se.lower = preds$calibrated.se.lower,                   
                  calibrated.se.upper = preds$calibrated.se.upper,                   
                  calibrated.estimate = preds$calibrated.estimate)
  # return the data input file with predictions attached.
  return(datin)
}
```

#### Individual_Inverse_Data_Model
```{r}
mod.spec.fn$Individual_Inverse_Data_Model <- function(datin, mod.i){
  # apply data input transformations
  datin$lnlpe = (log(datin$lpe))
  
  # apply model and get predictions (preds)
  preds <- as.data.frame(predict(mods[[mod.i]], type="link", 
                   dplyr::select(datin, -gazetted_stream_name),
                      re.form=NA, se.fit=T))
  # apply data transformations to predictions (calibrated.estimates) to get error bounds.
  preds <- mutate(preds,
                fit.cv = round2(sqrt(exp(se.fit^2)-1)*100,1),
                calibrated.se.lower = exp(fit-se.fit),
                calibrated.se.upper = exp(fit+se.fit),
                calibrated.estimate = exp(fit))
  # attach results to data input file
  datin <- mutate(datin,                   
                  fit = preds$fit,                   
                  se.fit = preds$se.fit,                   
                  fit.cv = preds$fit.cv,                   
                  calibrated.se.lower = preds$calibrated.se.lower,                   
                  calibrated.se.upper = preds$calibrated.se.upper,                   
                  calibrated.estimate = preds$calibrated.estimate)
  # return the data input file with predictions attached.
  return(dSatin)
}
```

# Save Model Function List
```{r}
saveRDS(mod.spec.fn, "./r_objects/03_mod_spec_fn.rds")
saveRDS(mods, "./r_objects/03_mods.rds")
```

