---
title: "03-03-Calibrated-Summaries"
author: "Michael Arbeider"
date: "2026-03-06"
---

# Set-up
## Packages
```{r}
rm(list = ls())
library(tidyverse) # most %>% 
theme_set(theme_bw()) # ggplot
library(patchwork) # for multi plot ggplot graphs
library(scales) # for label_comma()
library(stringr) # str_extract
source("00-Functions.R") # scale_x_reordered, rank.summary(), round2()
library(car) # leveneTest
library(rstatix) # pairwise, identify_outliers
library(lmtest) # lrtest(), likelihood ratio test
cal_4apply <- readRDS("./r_objects/03_cal_4apply.rds")
mods <- readRDS("./r_objects/03_mods.rds")
mod.spec.fn <- readRDS("./r_objects/03_mod_spec_fn.rds")
mod.specs <- readRDS("./r_objects/models/model_specifications.rds")
```

# Data
## Model Selection Results
Reading in the model selection results.
```{r}
mod.select.full <- readRDS(file = "./r_objects/02_01_rank_combined_results.rds") %>% 
  filter(special.level != 4.5)
mod.selection <- readRDS(file = "./r_objects/02_01_rank_top_results.rds")
mod.selection <- filter(mod.selection,
                        keep == TRUE)
```

### Model summaries
```{r}
# Models that passed
mod.selection %>% 
  filter(!str_detect(sys.meth.set, "All_Systems_Discharge")) %>% 
  ungroup() %>% 
  count(top.model)

# add a relative rank within each rank-set
mod.select.full <- mod.select.full %>% 
  group_by(sys.meth.set) %>% 
  mutate(rank = rank(rel.sum))

# count of model occurrence within the top 5
mod.select.full %>% 
  filter(rank <= 5) %>% 
  ungroup() %>% 
  count(source) %>% 
  arrange(n) %>% 
  print(., n = 30)

# count of model occurrence within the top 5 excluding the top
mod.select.full %>% 
  filter(rank <= 5,
         rank != 1) %>% 
  ungroup() %>% 
  count(source) %>% 
  arrange(n)

# count of model occurrence second from the top
mod.select.full %>% 
  filter(rank == 2) %>% 
  ungroup() %>% 
  count(source) %>% 
  arrange(n)

mod.select.full %>% 
  count(sys.meth.set)
```

## Calibration Set & Data
```{r}
cal <-readRDS("r_objects/sockeye_calibration_data.rds")

cal <- cal %>% 
  mutate(aggr = ifelse(size %in% c("Large", "X-Large"), "LONLY", "NoL"),
         water_clarity = ifelse(water_clarity == "Pt_Turb_Tannic", 
                                       "Turbid/Tannic", water_clarity))

cal.out <- readRDS("./r_objects/03_02_calibrated_series.rds") %>% 
  mutate(water_clarity = ifelse(water_clarity == "Pt_Turb_Tannic", "Turbid/Tannic", water_clarity),
         usid = paste(gazetted_stream_name, year, lpe_method, sep = "-"))
cal.out.long <- readRDS("./r_objects/03_02_calibrated_series_long.rds") %>% 
  mutate(water_clarity = ifelse(water_clarity == "Pt_Turb_Tannic", "Turbid/Tannic", water_clarity),
         usid = paste(gazetted_stream_name, year, lpe_method, sep = "-"),
         usid2 = paste(gazetted_stream_name, year, lpe_method, type, sep = "-"),
         type = as.factor(type),
         abin = as.factor(abin))
```
### Relative to 1.8
```{r}
cal.out <- cal.out %>% 
  mutate(water_clarity = ifelse(water_clarity == "Pt_Turb_Tannic", "Turbid/Tannic", water_clarity),
         calibrated.ci.lower = case_when(
           calibration.model == "Early_Stu_DeadAlive_Model" ~ (fit - se.fit*1.96)^3,
           TRUE ~ exp(fit - se.fit*1.96)
         ),
         calibrated.ci.upper = case_when(
           calibration.model == "Early_Stu_DeadAlive_Model" ~ (fit + se.fit*1.96)^3,
           TRUE ~ exp(fit + se.fit*1.96)
         ),
         per.change.class = (calibrated.estimate - classic)/classic,
         abs.per.change.class = abs(per.change.class),
         is.classic.in.se = classic >= calibrated.se.lower & classic <= calibrated.se.upper,
         is.per.change.within = abs.per.change.class < 0.15,
         is.hpe.within.se = hpe >= calibrated.se.lower & hpe <= calibrated.se.upper,
         is.hpe.within.ci = hpe >= calibrated.ci.lower & hpe <= calibrated.ci.upper,
         is.both.true = is.classic.in.se == TRUE & is.per.change.within == TRUE,
         calibrated.difference = calibrated.estimate - hpe,
         classic.difference = classic - hpe,
         calibrated.difference.ln = log(calibrated.estimate) - log(hpe),
         #ln.calibrated.difference = log(calibrated.difference),
         classic.difference.ln = log(classic) - log(hpe),
         #ln.classic.difference = log(classic.difference),
         per.error = calibrated.difference / hpe,
         classic.per.error = classic.difference / hpe,
         abs.per.error = abs(per.error),
         classic.abs.per.error = abs(classic.per.error),
         cal.cv = (calibrated.se.upper - calibrated.se.lower) / calibrated.estimate,
         true.CV = sqrt(exp(se.fit^2)-1))
```

# Summary Statistics
```{r}
cal.out.summary.bins <- cal.out.long %>% 
  group_by(abin, type) %>% 
  summarise(n = n(),
            MPE = mean(per.error),
            #Abs.MPE = abs(MPE),
            MAPE = mean(abs.per.error),
            RMSE = sqrt(mean(squared.error)),
            R.Cor = cor(estimate, hpe, method = "pearson"),
            total.expanded = sum(estimate),
            avg.high.accuracy = mean(hpe))

cal.out.summary <- cal.out.long %>% 
  group_by(abin = "All", type) %>% 
  summarise(n = n(),
            MPE = mean(per.error),
            #Abs.MPE = abs(MPE),
            MAPE = mean(abs.per.error),
            RMSE = sqrt(mean(squared.error)),
            R.Cor = cor(estimate, hpe, method = "pearson"),
            total.expanded = sum(estimate),
            avg.high.accuracy = mean(hpe)) %>% 
  rbind(., cal.out.summary.bins) %>% 
  mutate(MPE = round2(MPE*100, 1),
         MAPE = round2(MAPE*100, 1),
         RMSE = round2(RMSE, 0),
         R.Cor = round2(R.Cor*100, 1),
         total.expanded = round2(total.expanded, 0),
         avg.high.accuracy = round2(avg.high.accuracy, 0)) ; cal.out.summary
```
## Outliers
```{r}
cal.out.high.error <- filter(cal.out,
                             abs.per.error > 0.5)

# Summary statistics reported in the report.
# # Aerial outliers
cal.out.high.error %>% count(lpe_method) %>% mutate(n / sum(n))
# Aerial Occurrence in database
cal.out %>% count(lpe_method) %>% mutate(n / sum(n))
# Water Clarity Outliers
cal.out.high.error %>% count(water_clarity) %>% mutate(n / sum(n))
# Clarity occurrence in database
cal.out %>% count(water_clarity) %>% mutate(n / sum(n))
# Size Outliers
cal.out.high.error %>% count(size) %>% mutate(n / sum(n))
# Size occurrence in database
cal.out %>% count(size) %>% mutate(n / sum(n))

cal.out.high.error %>% count(lpe_method, water_clarity) %>% mutate(n / sum(n))
cal.out %>% count(lpe_method, water_clarity) %>% mutate(n / sum(n))

# Summary of Outliers by Model
cal.out.high.error %>% count(calibration.model) %>% 
  mutate(prop = n / sum(n)) %>% 
  select(-n) %>% 
  left_join(., cal.out %>% count(calibration.model) %>% mutate(prop2 = n / sum(n))) %>% 
  mutate(diff = (prop - prop2)*100)

# By Model and method / clarity
cal.out.high.error %>% 
  group_by(lpe_method, water_clarity, size, gazetted_stream_name, calibration.model) %>% 
  summarize(high.error = n()) %>% 
  left_join(count(cal, lpe_method, water_clarity, gazetted_stream_name)) %>% 
  mutate(per.high = high.error/n*100) %>% 
  arrange(lpe_method, desc(per.high)) 

# Default 1.8 number of outliers
filter(cal.out.long, type == "historical", abs.per.error > 0.5) %>% 
  count()

# number of overlapping outliers
filter(cal.out.long, abs.per.error > 0.5) %>% 
  count(usid) %>% filter(n == 2)
# number of new outliers
23-12

# percent remained outliers
12 / 28

# non-overlapping outliers
cal.out.high.error %>% 
  filter(!usid %in% cal.out.long$usid[cal.out.long$type == "historical" & 
                                         cal.out.long$abs.per.error > 0.5]) %>% 
  count(size, lpe_method, water_clarity)

# Aerial
9/11
```

## Precision
```{r}
# Count of high accuracy estimates within SE and CI
count(cal.out, is.hpe.within.se) %>%  mutate(n / sum(n))
count(cal.out, is.hpe.within.ci) %>%  mutate(n / sum(n))

# Average CV of estimates with hpe within SE and CI
cal.out %>% group_by(is.hpe.within.se) %>% 
  summarise(mean.cv = mean(fit.cv))
cal.out %>% group_by(is.hpe.within.ci) %>% 
  summarise(mean.cv = mean(fit.cv))

# Summary of outliers CVs and if they captured the true estimate.
cal.out %>% filter(abs.per.error > 0.5) %>% 
  group_by(is.hpe.within.se) %>% 
  summarise(n = n(), mean.cv = mean(fit.cv)) %>%  mutate(n / sum(n))
cal.out %>% filter(abs.per.error > 0.5) %>% 
  group_by(is.hpe.within.ci) %>% 
  summarise(n = n(), mean.cv = mean(fit.cv))  %>% mutate(n / sum(n))

# Precision Summary
count(cal.out, water_clarity, is.hpe.within.se) %>%  
  group_by(water_clarity) %>% 
  mutate(n / sum(n))

count(cal.out, lpe_method, is.hpe.within.se) %>%  
  group_by(lpe_method) %>% 
  mutate(n / sum(n))

count(cal.out, size, is.hpe.within.se) %>%  
  group_by(size) %>% 
  mutate(n / sum(n))

count(cal.out, calibration.model, is.hpe.within.se) %>% 
  filter(is.hpe.within.se == FALSE) %>% 
  mutate(prop = n / 162)

lm.precision <- lm(fit.cv ~ is.hpe.within.se, data = cal.out)
summary(lm.precision)


cal.out %>% group_by(is.hpe.within.se) %>% 
  summarise(mean.cv = mean(true.CV))

cal.out %>% group_by(is.hpe.within.ci) %>% 
  summarise(mean.cv = mean(fit.cv))

ggplot(cal.out, 
       aes(x = is.hpe.within.se, y = fit.cv)) +
  geom_boxplot()

lm.precision.model <- lm(fit.cv ~ calibration.model, data = cal.out)
summary(lm.precision.model)

ggplot(cal.out, 
       aes(x = calibration.model, y = fit.cv,
           fill = is.hpe.within.se)) +
  geom_boxplot()


ggplot(cal.out, 
       aes(x = log(calibrated.estimate), y = is.hpe.within.se,
           colour = calibration.model)) +
  geom_point()
```

## Performance Metrics
By Abundance Bins and Water Clarity
```{r}
cal.out.summary.wc <- cal.out.long %>% 
  group_by(abin, type, water_clarity) %>% 
  summarise(n= n(),
            MPE = mean(per.error),
            #Abs.MPE = abs(MPE),
            MAPE = mean(abs.per.error),
            RMSE = sqrt(mean(squared.error)),
            R.Cor = cor(estimate, hpe, method = "pearson"),
            total.expanded = sum(estimate),
            avg.high.accuracy = mean(hpe)) %>% 
  mutate(MPE = round2(MPE*100, 1),
         MAPE = round2(MAPE*100, 1),
         RMSE = round2(RMSE, 0),
         R.Cor = round2(R.Cor*100, 1),
         total.expanded = round2(total.expanded, 0),
         avg.high.accuracy = round2(avg.high.accuracy, 0)) ; cal.out.summary.wc

write.csv(cal.out.summary.wc, 
          file = "./tables/03_03-Comparison with Historical Summary Table with Water Clarity.csv",
          row.names = F)

p.MPE <- ggplot(cal.out.long,
                aes(x = abin, y = per.error,
                     fill = type)) +
    geom_hline(yintercept = 0) +
  geom_boxplot(alpha = 0.85) +
  facet_wrap(~water_clarity) +
  scale_fill_manual(name = "Estimate",
                    values = c("salmon", "grey")) +
  labs(x = "Abundance Bin", y = "Percent Error") ; p.MPE

p.MAPE <- ggplot(cal.out.long,
                aes(x = abin, y = abs.per.error,
                     fill = type)) +
  geom_boxplot(alpha = 0.85) +
  facet_wrap(~water_clarity) +
  scale_fill_manual(name = "Estimate",
                    values = c("salmon", "grey")) +
  #geom_hline(yintercept = 0) +
  labs(x = "Abundance Bin", y = "Absolute Percent Error") ; p.MAPE
```

### Combined
```{r}
cal.out.summary$water_clarity <- "Combined"
cal.out.summary.combined <- rbind(cal.out.summary.wc, cal.out.summary) %>% 
  mutate(water_clarity = factor(water_clarity,
                                levels = c("Clear", "Turbid/Tannic", "Combined"
                                ))) %>% 
  arrange(abin, water_clarity)

write.csv(cal.out.summary.combined, 
          file = "./tables/03_03-Comparison-with-Historical-Summary-Table-Combined.csv",
          row.names = F)
```

### Extra Detail
```{r}
cal.out.summary.extra <- cal.out.long %>% 
  group_by(abin, type, size, water_clarity) %>% 
  summarise(n= n(),
            MPE = mean(per.error),
            #Abs.MPE = abs(MPE),
            MAPE = mean(abs.per.error),
            RMSE = sqrt(mean(squared.error)),
            total.expanded = sum(estimate),
            avg.high.accuracy = mean(hpe)) %>% 
  mutate(MPE = round2(MPE*100, 1),
         MAPE = round2(MAPE*100, 1),
         RMSE = round2(RMSE, 0),
         total.expanded = round2(total.expanded, 0),
         avg.high.accuracy = round2(avg.high.accuracy, 0)) ; cal.out.summary.extra

cal.out.summary.extra <- cal.out.summary.extra %>% 
  mutate(water_clarity = factor(water_clarity,
                                levels = c("Clear", "Turbid/Tannic", "Combined"
                                )),
         size = factor(size,
                       levels = c("Small", "Medium",
                                  "Large", "X-Large"))) %>% 
  arrange(abin, size, water_clarity)

write.csv(cal.out.summary.extra, 
          file = "./tables/03_03-Comparison-with-Historical-Summary-Table-Extra.csv",
          row.names = F)
```

```{r}
cal.out.summary.extra <- cal.out.long %>% 
  group_by(abin, type, size, lpe_method, water_clarity) %>% 
  summarise(n= n(),
            MPE = mean(per.error),
            #Abs.MPE = abs(MPE),
            MAPE = mean(abs.per.error),
            RMSE = sqrt(mean(squared.error)),
            total.expanded = sum(estimate),
            avg.high.accuracy = mean(hpe)) %>% 
  mutate(MPE = round2(MPE*100, 1),
         MAPE = round2(MAPE*100, 1),
         RMSE = round2(RMSE, 0),
         total.expanded = round2(total.expanded, 0),
         avg.high.accuracy = round2(avg.high.accuracy, 0)) ; cal.out.summary.extra

cal.out.summary.extra <- cal.out.summary.extra %>% 
  mutate(water_clarity = factor(water_clarity,
                                levels = c("Clear", "Turbid/Tannic", "Combined"
                                )),
         size = factor(size,
                       levels = c("Small", "Medium",
                                  "Large", "X-Large"))) %>% 
  arrange(abin, size, lpe_method, water_clarity)

write.csv(cal.out.summary.extra, 
          file = "./tables/03_03-Comparison-with-Historical-Summary-Table-Extra-Method.csv",
          row.names = F)
```


### ANOVAs and T-Tests
#### Percent Error
```{r}
cal.lm.mpe <- lm(per.error ~ 0 + type*abin, data = cal.out.long)
summary(cal.lm.mpe)

plot(fitted(cal.lm.mpe), residuals(cal.lm.mpe))
abline(h = 0)

plot(cal.lm.mpe)

aov.mpe <- aov(per.error ~ type*abin, data = cal.out.long)
summary(aov.mpe)
TukeyHSD(aov.mpe, which = c("type", "abin"))

# outliers?
cal.out.long %>%
     group_by(abin, type) %>%
     identify_outliers(per.error) %>% 
  filter(type == "updated")

# Homogeneity of variance test from https://www.datanovia.com/en/lessons/mixed-anova-in-r/#post-hoc-tests
cal.out.long %>%
  group_by(abin) %>%
  levene_test(per.error ~ type)

pair.mpe <- cal.out.long %>%
  group_by(abin) %>%
  pairwise_t_test(per.error ~ type, paired = TRUE, 
                  p.adjust.method = "bonferroni")
pair.mpe

t.mpe <- cal.out.long %>%
 # group_by(abin, type) %>%
  t_test(per.error~type, paired = TRUE)
t.mpe

# Updated relative to zero
t.mpe.0u <- cal.out.long %>%
  filter(type == "updated") %>%
  t_test(per.error ~ 1, mu = 0)
t.mpe.0u

# Default relative to zero
t.mpe.0d <- cal.out.long %>%
  filter(type == "historical") %>%
  t_test(per.error ~ 1, mu = 0)
t.mpe.0d

# Updated by Abuns relative to zero
pair.mpe.0u <- cal.out.long %>%
  filter(type == "updated") %>% 
  group_by(abin) %>%
  t_test(per.error ~ 1, mu = 0)
pair.mpe.0u

# Default by Abuns relative to zero
pair.mpe.0d <- cal.out.long %>%
  filter(type == "historical") %>% 
  group_by(abin) %>%
  t_test(per.error ~ 1, mu = 0)
pair.mpe.0d

lm.mpe.dat <- expand.grid(abin = unique(cal.out.long$abin),
                       type = unique(cal.out.long$type)) %>% 
  mutate(abin = as.character(abin),
         type = as.character(type))

lm.mpe.dat <- cbind(lm.mpe.dat, 
                 as.data.frame(
                   predict(cal.lm.mpe, newdata = lm.mpe.dat, 
                           type = "response", se.fit = TRUE)
                   )
                 )
```
#### Absolute Percent Error
```{r}
cal.lm.mape <- lm(abs.per.error ~ type*abin, data = cal.out.long)
summary(cal.lm.mape)

plot(fitted(cal.lm.mape), residuals(cal.lm.mape))
abline(h = 0)

# plot(cal.lm.mape)

aov.mape <- aov(abs.per.error ~ type*abin, data = cal.out.long)
summary(aov.mape)
TukeyHSD(aov.mape, which = c("type", "abin"))

# outliers?
cal.out.long %>%
     group_by(abin, type) %>%
     identify_outliers(abs.per.error) %>% 
  filter(type == "updated")

# Homogeneity of variance test from https://www.datanovia.com/en/lessons/mixed-anova-in-r/#post-hoc-tests
cal.out.long %>%
  group_by(abin) %>%
  levene_test(abs.per.error ~ type)

pair.mape <- cal.out.long %>%
  group_by(abin) %>%
  pairwise_t_test(abs.per.error ~ type, paired = TRUE, 
                  p.adjust.method = "bonferroni")
pair.mape

t.mape <- cal.out.long %>%
 # group_by(abin, type) %>%
  t_test(abs.per.error~type, paired = TRUE)
t.mape

# Updated relative to zero
t.mape.0u <- cal.out.long %>%
  filter(type == "updated") %>%
  t_test(abs.per.error ~ 1, mu = 0, alternative = "greater")
t.mape.0u

# Default relative to zero
t.mape.0d <- cal.out.long %>%
  filter(type == "historical") %>%
  t_test(abs.per.error ~ 1, mu = 0, alternative = "greater")
t.mape.0d

# Updated by Abuns relative to zero
pair.mape.0u <- cal.out.long %>%
  filter(type == "updated") %>% 
  group_by(abin) %>%
  t_test(abs.per.error ~ 1, mu = 0, alternative = "greater")
pair.mape.0u

# Default by Abuns relative to zero
pair.mape.0d <- cal.out.long %>%
  filter(type == "historical") %>% 
  group_by(abin) %>%
  t_test(abs.per.error ~ 1, mu = 0, alternative = "greater")
pair.mape.0d

lm.mape.dat <- expand.grid(abin = unique(cal.out.long$abin),
                       type = unique(cal.out.long$type)) %>% 
  mutate(abin = as.character(abin),
         type = as.character(type))

lm.mape.dat <- cbind(lm.mape.dat, 
                 as.data.frame(
                   predict(cal.lm.mape, newdata = lm.mape.dat, 
                           type = "response", se.fit = TRUE)
                   )
                 )
```

### LM + Water Clarity
#### Percent Error
```{r}
cal.lm.mpe.wc <- lm(per.error ~ type*abin*water_clarity, data = cal.out.long)
summary(cal.lm.mpe.wc)

plot(fitted(cal.lm.mpe.wc), residuals(cal.lm.mpe.wc))
abline(h = 0)

# plot(cal.lm.mpe.wc)

aov.mpe.wc <- aov(per.error ~ type*abin*water_clarity, data = cal.out.long)
summary(aov.mpe.wc)
TukeyHSD(aov.mpe.wc, which = c("type", "abin", "water_clarity"))

# outliers?
cal.out.long %>%
     group_by(abin, water_clarity, type) %>%
     identify_outliers(per.error) %>% 
  filter(type == "updated")

# Homogeneity of variance test from https://www.datanovia.com/en/lessons/mixed-anova-in-r/#post-hoc-tests
cal.out.long %>%
  group_by(abin, water_clarity) %>%
  levene_test(per.error ~ type)

pair.mpe.wc <- cal.out.long %>%
  group_by(abin, water_clarity) %>%
  pairwise_t_test(per.error ~ type, paired = TRUE, 
                  p.adjust.method = "bonferroni")
pair.mpe.wc

# Updated by Abuns relative to zero
pair.mpe.wc.0u <- cal.out.long %>%
  filter(type == "updated") %>% 
  group_by(abin, water_clarity) %>%
  t_test(per.error ~ 1, mu = 0)
pair.mpe.wc.0u

# Default by Abuns relative to zero
pair.mpe.wc.0d <- cal.out.long %>%
  filter(type == "historical") %>% 
  group_by(abin, water_clarity) %>%
  t_test(per.error ~ 1, mu = 0)
pair.mpe.wc.0d

lm.mpe.wc.dat <- expand.grid(abin = unique(cal.out.long$abin),
                       type = unique(cal.out.long$type),
                       water_clarity = unique(cal.out.long$water_clarity)) %>% 
  mutate(abin = as.character(abin),
         type = as.character(type),
         water_clarity = as.character(water_clarity))

lm.mpe.wc.dat <- cbind(lm.mpe.wc.dat, 
                 as.data.frame(
                   predict(cal.lm.mpe.wc, newdata = lm.mpe.wc.dat, 
                           type = "response", se.fit = TRUE)
                   )
                 )
```

#### Absolute Percent Error
```{r}
cal.lm.mape.wc <- lm(abs.per.error ~ type*abin*water_clarity, data = cal.out.long)
summary(cal.lm.mape.wc)

plot(fitted(cal.lm.mape.wc), residuals(cal.lm.mape.wc))
abline(h = 0)

# plot(cal.lm.mape.wc)

aov.mape.wc <- aov(abs.per.error ~ type*abin*water_clarity, data = cal.out.long)
summary(aov.mape.wc)
TukeyHSD(aov.mape.wc, which = c("type", "abin", "water_clarity"))

# outliers?
cal.out.long %>%
     group_by(abin, water_clarity, type) %>%
     identify_outliers(abs.per.error) %>% 
  filter(type == "updated")

# Homogeneity of variance test from https://www.datanovia.com/en/lessons/mixed-anova-in-r/#post-hoc-tests
cal.out.long %>%
  group_by(abin, water_clarity) %>%
  levene_test(abs.per.error ~ type)

pair.mape.wc <- cal.out.long %>%
  group_by(abin, water_clarity) %>%
  pairwise_t_test(abs.per.error ~ type, paired = TRUE, 
                  p.adjust.method = "bonferroni")
pair.mape.wc

# Updated by Abuns relative to zero
pair.mape.wc.0u <- cal.out.long %>%
  filter(type == "updated") %>% 
  group_by(abin, water_clarity) %>%
  t_test(abs.per.error ~ 1, mu = 0, alternative = "greater")
pair.mape.wc.0u

# Default by Abuns relative to zero
pair.mape.wc.0d <- cal.out.long %>%
  filter(type == "historical") %>% 
  group_by(abin, water_clarity) %>%
  t_test(abs.per.error ~ 1, mu = 0, alternative = "greater")
pair.mape.wc.0d

lm.mape.wc.dat <- expand.grid(abin = unique(cal.out.long$abin),
                       type = unique(cal.out.long$type),
                       water_clarity = unique(cal.out.long$water_clarity)) %>% 
  mutate(abin = as.character(abin),
         type = as.character(type),
         water_clarity = as.character(water_clarity))

lm.mape.wc.dat <- cbind(lm.mape.wc.dat, 
                 as.data.frame(
                   predict(cal.lm.mape.wc, newdata = lm.mape.wc.dat, 
                           type = "response", se.fit = TRUE)
                   )
                 )
```

#### Summary Table
```{r}
cal.out.mpe.mape.summary <- cal.out.summary.combined %>% 
  mutate(type = ifelse(type == "historical", "MPE", "MAPE")) %>% 
  dplyr::select(abin:n) %>% 
  rename(metric = type) %>% 
  mutate(updated.to.default = NA,
         updated.to.zero = NA,
         default.to.zero = NA)

### Relative to Historical ###
# MPE All Combined Relative to Historical
cal.out.mpe.mape.summary <- within(cal.out.mpe.mape.summary, {
     updated.to.default[abin == "All" & metric == "MPE" & 
                             water_clarity == "Combined"] <- t.mpe$p})
# MPE Abins Relative to Historical
cal.out.mpe.mape.summary <- within(cal.out.mpe.mape.summary, {
     updated.to.default[abin != "All" & metric == "MPE" & 
                             water_clarity == "Combined"] <- pair.mpe$p})
# MAPE All Combined Relative to Historical
cal.out.mpe.mape.summary <- within(cal.out.mpe.mape.summary, {
     updated.to.default[abin == "All" & metric == "MAPE" & 
                             water_clarity == "Combined"] <- t.mape$p})
# MAPE Abins Relative to Historical
cal.out.mpe.mape.summary <- within(cal.out.mpe.mape.summary, {
     updated.to.default[abin != "All" & metric == "MAPE" & 
                             water_clarity == "Combined"] <- pair.mape$p})
# MPE Abins & Water Clarity Relative to Historical
cal.out.mpe.mape.summary <- within(cal.out.mpe.mape.summary, {
     updated.to.default[abin != "All" & metric == "MPE" & 
                             water_clarity != "Combined"] <- pair.mpe.wc$p})
# MAPE Abins & Water Clarity Relative to Historical
cal.out.mpe.mape.summary <- within(cal.out.mpe.mape.summary, {
     updated.to.default[abin != "All" & metric == "MAPE" & 
                             water_clarity != "Combined"] <- pair.mape.wc$p})

### Updated Relative to Zero ###
# MPE All Combined
cal.out.mpe.mape.summary <- within(cal.out.mpe.mape.summary, {
     updated.to.zero[abin == "All" & metric == "MPE" & 
                             water_clarity == "Combined"] <- t.mpe.0u$p})
# MPE Abins 
cal.out.mpe.mape.summary <- within(cal.out.mpe.mape.summary, {
     updated.to.zero[abin != "All" & metric == "MPE" & 
                             water_clarity == "Combined"] <- pair.mpe.0u$p})
# MAPE All Combined 
cal.out.mpe.mape.summary <- within(cal.out.mpe.mape.summary, {
     updated.to.zero[abin == "All" & metric == "MAPE" & 
                             water_clarity == "Combined"] <- t.mape.0u$p})
# MAPE Abins 
cal.out.mpe.mape.summary <- within(cal.out.mpe.mape.summary, {
     updated.to.zero[abin != "All" & metric == "MAPE" & 
                             water_clarity == "Combined"] <- pair.mape.0u$p})
# MPE Abins & Water Clarity 
cal.out.mpe.mape.summary <- within(cal.out.mpe.mape.summary, {
     updated.to.zero[abin != "All" & metric == "MPE" & 
                             water_clarity != "Combined"] <- pair.mpe.wc.0u$p})
# MAPE Abins & Water Clarity 
cal.out.mpe.mape.summary <- within(cal.out.mpe.mape.summary, {
     updated.to.zero[abin != "All" & metric == "MAPE" & 
                             water_clarity != "Combined"] <- pair.mape.wc.0u$p})

### Deafult Relative to Zero ###
# MPE All Combined
cal.out.mpe.mape.summary <- within(cal.out.mpe.mape.summary, {
     default.to.zero[abin == "All" & metric == "MPE" & 
                             water_clarity == "Combined"] <- t.mpe.0d$p})
# MPE Abins 
cal.out.mpe.mape.summary <- within(cal.out.mpe.mape.summary, {
     default.to.zero[abin != "All" & metric == "MPE" & 
                             water_clarity == "Combined"] <- pair.mpe.0d$p})
# MAPE All Combined 
cal.out.mpe.mape.summary <- within(cal.out.mpe.mape.summary, {
     default.to.zero[abin == "All" & metric == "MAPE" & 
                             water_clarity == "Combined"] <- t.mape.0d$p})
# MAPE Abins 
cal.out.mpe.mape.summary <- within(cal.out.mpe.mape.summary, {
     default.to.zero[abin != "All" & metric == "MAPE" & 
                             water_clarity == "Combined"] <- pair.mape.0d$p})
# MPE Abins & Water Clarity 
cal.out.mpe.mape.summary <- within(cal.out.mpe.mape.summary, {
     default.to.zero[abin != "All" & metric == "MPE" & 
                             water_clarity != "Combined"] <- pair.mpe.wc.0d$p})
# MAPE Abins & Water Clarity 
cal.out.mpe.mape.summary <- within(cal.out.mpe.mape.summary, {
     default.to.zero[abin != "All" & metric == "MAPE" & 
                             water_clarity != "Combined"] <- pair.mape.wc.0d$p})

# Clean
cal.out.mpe.mape.summary.clean <- cal.out.mpe.mape.summary %>% 
  mutate_at(c("updated.to.default","updated.to.zero","default.to.zero"), 
            round2, digits = 4) %>% 
  mutate(
    across(c(updated.to.default, updated.to.zero, default.to.zero),
           ~ if_else(. < 0.0025,
                     "< 0.0025*",
                     as.character(.))
           )
  )

write.csv(cal.out.mpe.mape.summary.clean, 
          file = "./tables/03_03-Comparison-with-Historical-Summary-Statistics-Combined.csv",
          row.names = F)
```

### Through Time
```{r}
hist(cal.out$year, breaks = 3)

cal.out.summary.years <- cal.out.long %>% 
  mutate(ybin = case_when(
    year < 2004 ~ "1988-2003",
    year >= 2004 & year <= 2010 ~ "2004-2010",
    year > 2010 ~ "2011-2022"
  )) %>% 
  group_by(ybin, type) %>% 
  summarise(n = n(),
            MPE = mean(per.error),
            #Abs.MPE = abs(MPE),
            MAPE = mean(abs.per.error),
            RMSE = sqrt(mean(squared.error)),
            R.Cor = cor(estimate, hpe, method = "pearson"),
            Abundance = sum(estimate)) %>% 
  mutate(MPE = round2(MPE*100, 1),
         MAPE = round2(MAPE*100, 1),
         RMSE = round2(RMSE, 0),
         R.Cor = round2(R.Cor*100, 1),
         Abundance = round2(Abundance, 0)) ; cal.out.summary.years

write.csv(cal.out.summary.years, 
          file = "./tables/03_03-Comparison-with-Historical-Summary-Table-Year-Bins.csv",
          row.names = F)

ggplot(cal.out.long,
       aes(x = year, y = abs.per.error)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(type~abin)

ggplot(cal.out.long,
       aes(x = year, y = per.error)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(type~abin)
```
#### LM Tests
Summary Table to Populate
```{r}
table.temporal.results <- data.frame(
  "Metric" = c(rep("MPE", 4), rep("MAPE", 4)),
  "Calibration.Method" = c(rep(c(rep("Default",2),rep("Updated",2)),2)),
  "Model" = c(rep(c("Temporal General", 
                  "Interaction w/Abundance"),4)),
  "Ratio.p.value" = NA,
  "R2" = NA,
  "Slope" = NA,
  "SE" = NA
)
```
General Temporal Trend
```{r}
# MPE of Default
lm.t.mpe.D <- lm(per.error ~ year, 
               data = cal.out.long[cal.out.long$type == "historical", ])
summary(lm.t.mpe.D)
# MPE of Updated
lm.t.mpe.U <- lm(per.error ~ year, 
               data = cal.out.long[cal.out.long$type == "updated", ])
summary(lm.t.mpe.U)
# Save R2, Slopes, and Errors
table.temporal.results$R2[table.temporal.results$Metric == "MPE" & 
                            table.temporal.results$Model == "Temporal General"] = c(summary(lm.t.mpe.D)$adj.r.squared,
                                                                                    summary(lm.t.mpe.U)$adj.r.squared)
table.temporal.results$Slope[table.temporal.results$Metric == "MPE" & 
                            table.temporal.results$Model == "Temporal General"] = c(coef(lm.t.mpe.D)[[2]],
                                                                                    coef(lm.t.mpe.U)[[2]])
table.temporal.results$SE[table.temporal.results$Metric == "MPE" & 
                            table.temporal.results$Model == "Temporal General"] = c(summary(lm.t.mpe.D)$coefficients[2,2],
                                                                                    summary(lm.t.mpe.U)$coefficients[2,2])
# MAPE of Default
lm.t.mape.D <- lm(abs.per.error ~ year, 
               data = cal.out.long[cal.out.long$type == "historical", ])
summary(lm.t.mape.D)
# MAPE of Updated
lm.t.mape.U <- lm(abs.per.error ~ year, 
               data = cal.out.long[cal.out.long$type == "updated", ])
summary(lm.t.mape.U)

table.temporal.results$R2[table.temporal.results$Metric == "MAPE" & 
                            table.temporal.results$Model == "Temporal General"] = c(summary(lm.t.mape.D)$adj.r.squared,
                                                                                    summary(lm.t.mape.U)$adj.r.squared)
table.temporal.results$Slope[table.temporal.results$Metric == "MAPE" & 
                            table.temporal.results$Model == "Temporal General"] = c(coef(lm.t.mape.D)[[2]],
                                                                                    coef(lm.t.mape.U)[[2]])
table.temporal.results$SE[table.temporal.results$Metric == "MAPE" & 
                            table.temporal.results$Model == "Temporal General"] = c(summary(lm.t.mape.D)$coefficients[2,2],
                                                                                    summary(lm.t.mape.U)$coefficients[2,2])

table.temporal.results <- table.temporal.results %>% 
  mutate_at(c("R2","Slope","SE"), round2, digits = 3)
```
With Size Bin covariates
```{r}
# MPE of Default
lm.ta.mpe.D <- lm(per.error ~ year*abin, 
               data = cal.out.long[cal.out.long$type == "historical", ])
lm.ta.mpe.D.sum <- summary(lm.ta.mpe.D); lm.ta.mpe.D.sum
# Slopes & SEs for each Abundance Bin
slope.ta.mpe.D.b1 <- coef(lm.ta.mpe.D.sum)["year", "Estimate"]
se.ta.mpe.D.b1    <- coef(lm.ta.mpe.D.sum)["year", "Std. Error"]
# For b2:
slope.ta.mpe.D.b2 <- slope.ta.mpe.D.b1 + coef(lm.ta.mpe.D.sum)["year:abin25-75k", "Estimate"]
se.ta.mpe.D.b2    <- sqrt(se.ta.mpe.D.b1^2 + 
                            coef(lm.ta.mpe.D.sum)["year:abin25-75k", "Std. Error"]^2)
# For b3:
slope.ta.mpe.D.b3 <- slope.ta.mpe.D.b1 + coef(lm.ta.mpe.D.sum)["year:abin75k+", "Estimate"]
se.ta.mpe.D.b3    <- sqrt(se.ta.mpe.D.b1^2 + 
                   coef(lm.ta.mpe.D.sum)["year:abin75k+", "Std. Error"]^2)
# MPE of Updated
lm.ta.mpe.U <- lm(per.error ~ year*abin, 
               data = cal.out.long[cal.out.long$type == "updated", ])
lm.ta.mpe.U.sum <- summary(lm.ta.mpe.U); lm.ta.mpe.U.sum
# Slopes & SEs for each Abundance Bin
slope.ta.mpe.U.b1 <- coef(lm.ta.mpe.U.sum)["year", "Estimate"]
se.ta.mpe.U.b1    <- coef(lm.ta.mpe.U.sum)["year", "Std. Error"]
# For b2:
slope.ta.mpe.U.b2 <- slope.ta.mpe.U.b1 + coef(lm.ta.mpe.U.sum)["year:abin25-75k", "Estimate"]
se.ta.mpe.U.b2    <- sqrt(se.ta.mpe.U.b1^2 + 
                            coef(lm.ta.mpe.U.sum)["year:abin25-75k", "Std. Error"]^2)
# For b3:
slope.ta.mpe.U.b3 <- slope.ta.mpe.U.b1 + coef(lm.ta.mpe.U.sum)["year:abin75k+", "Estimate"]
se.ta.mpe.U.b3    <- sqrt(se.ta.mpe.U.b1^2 + 
                   coef(lm.ta.mpe.U.sum)["year:abin75k+", "Std. Error"]^2)

# Save R2, Slopes, and Errors
table.temporal.results$R2[table.temporal.results$Metric == "MPE" & 
                            table.temporal.results$Model == "Interaction w/Abundance"] = round2(c(summary(lm.ta.mpe.D)$adj.r.squared,
                                                                                    summary(lm.ta.mpe.U)$adj.r.squared), 3)

table.temporal.results$Slope[table.temporal.results$Metric == "MPE" & 
                            table.temporal.results$Model == "Interaction w/Abundance"] = c(paste(round2(c(slope.ta.mpe.D.b1,
                                                                                                          slope.ta.mpe.D.b2,
                                                                                                          slope.ta.mpe.D.b3), 3), collapse = "; "),
                                                                                           paste(round2(c(slope.ta.mpe.U.b1,
                                                                                                          slope.ta.mpe.U.b2,
                                                                                                          slope.ta.mpe.U.b3), 3), collapse = "; "))
table.temporal.results$SE[table.temporal.results$Metric == "MPE" & 
                            table.temporal.results$Model == "Interaction w/Abundance"] = c(paste(round2(c(se.ta.mpe.D.b1,
                                                                                                          se.ta.mpe.D.b2,
                                                                                                          se.ta.mpe.D.b3), 3), collapse = "; "),
                                                                                           paste(round2(c(se.ta.mpe.U.b1,
                                                                                                          se.ta.mpe.U.b2,
                                                                                                          se.ta.mpe.U.b3), 3), collapse = "; "))


# MAPE of Default
lm.ta.mape.D <- lm(abs.per.error ~ year*abin, 
               data = cal.out.long[cal.out.long$type == "historical", ])
lm.ta.mape.D.sum <- summary(lm.ta.mape.D); lm.ta.mape.D.sum
# Slopes & SEs for each Abundance Bin
slope.ta.mape.D.b1 <- coef(lm.ta.mape.D.sum)["year", "Estimate"]
se.ta.mape.D.b1    <- coef(lm.ta.mape.D.sum)["year", "Std. Error"]
# For b2:
slope.ta.mape.D.b2 <- slope.ta.mape.D.b1 + coef(lm.ta.mape.D.sum)["year:abin25-75k", "Estimate"]
se.ta.mape.D.b2    <- sqrt(se.ta.mape.D.b1^2 + 
                            coef(lm.ta.mape.D.sum)["year:abin25-75k", "Std. Error"]^2)
# For b3:
slope.ta.mape.D.b3 <- slope.ta.mape.D.b1 + coef(lm.ta.mape.D.sum)["year:abin75k+", "Estimate"]
se.ta.mape.D.b3    <- sqrt(se.ta.mape.D.b1^2 + 
                   coef(lm.ta.mape.D.sum)["year:abin75k+", "Std. Error"]^2)
# MAPE of Updated
lm.ta.mape.U <- lm(abs.per.error ~ year*abin, 
               data = cal.out.long[cal.out.long$type == "updated", ])
lm.ta.mape.U.sum <- summary(lm.ta.mape.U); lm.ta.mape.U.sum
# Slopes & SEs for each Abundance Bin
slope.ta.mape.U.b1 <- coef(lm.ta.mape.U.sum)["year", "Estimate"]
se.ta.mape.U.b1    <- coef(lm.ta.mape.U.sum)["year", "Std. Error"]
# For b2:
slope.ta.mape.U.b2 <- slope.ta.mape.U.b1 + coef(lm.ta.mape.U.sum)["year:abin25-75k", "Estimate"]
se.ta.mape.U.b2    <- sqrt(se.ta.mape.U.b1^2 + 
                            coef(lm.ta.mape.U.sum)["year:abin25-75k", "Std. Error"]^2)
# For b3:
slope.ta.mape.U.b3 <- slope.ta.mape.U.b1 + coef(lm.ta.mape.U.sum)["year:abin75k+", "Estimate"]
se.ta.mape.U.b3    <- sqrt(se.ta.mape.U.b1^2 + 
                   coef(lm.ta.mape.U.sum)["year:abin75k+", "Std. Error"]^2)

# Save R2, Slopes, and Errors
table.temporal.results$R2[table.temporal.results$Metric == "MAPE" & 
                            table.temporal.results$Model == "Interaction w/Abundance"] = round2(c(summary(lm.ta.mape.D)$adj.r.squared,
                                                                                    summary(lm.ta.mape.U)$adj.r.squared), 3)

table.temporal.results$Slope[table.temporal.results$Metric == "MAPE" & 
                            table.temporal.results$Model == "Interaction w/Abundance"] = c(paste(round2(c(slope.ta.mape.D.b1,
                                                                                                          slope.ta.mape.D.b2,
                                                                                                          slope.ta.mape.D.b3), 3), collapse = "; "),
                                                                                           paste(round2(c(slope.ta.mape.U.b1,
                                                                                                          slope.ta.mape.U.b2,
                                                                                                          slope.ta.mape.U.b3), 3), collapse = "; "))
table.temporal.results$SE[table.temporal.results$Metric == "MAPE" & 
                            table.temporal.results$Model == "Interaction w/Abundance"] = c(paste(round2(c(se.ta.mape.D.b1,
                                                                                                          se.ta.mape.D.b2,
                                                                                                          se.ta.mape.D.b3), 3), collapse = "; "),
                                                                                           paste(round2(c(se.ta.mape.U.b1,
                                                                                                          se.ta.mape.U.b2,
                                                                                                          se.ta.mape.U.b3), 3), collapse = "; "))
```
Intercept/Null Models
```{r}
# MPE of Default
lm.mpe.D <- lm(per.error ~ 1, 
               data = cal.out.long[cal.out.long$type == "historical", ])
summary(lm.mpe.D)
# MPE of Updated
lm.mpe.U <- lm(per.error ~ 1, 
               data = cal.out.long[cal.out.long$type == "updated", ])
summary(lm.mpe.U)


# MAPE of Default
lm.mape.D <- lm(abs.per.error ~ 1, 
               data = cal.out.long[cal.out.long$type == "historical", ])
summary(lm.mape.D)
# MAPE of Updated
lm.mape.U <- lm(abs.per.error ~ 1, 
               data = cal.out.long[cal.out.long$type == "updated", ])
summary(lm.mape.U)
```

#### LRT
P < 0.05 means that the more complicated model significantly adds explanatory power
```{r}
# MPE
# Default 1.8 Series tests
# Is the interaction with abundance bins significant
lrt.ta.mpe.D <- lrtest(lm.t.mpe.D, lm.ta.mpe.D)
lrt.ta.mpe.D # significant
# Is a temporal covariate significant
lrt.t.mpe.D <- lrtest(lm.mpe.D, lm.t.mpe.D)
lrt.t.mpe.D # significant
# Updated Series Tests
# Is the interaction with abundance bins significant
lrt.ta.mpe.U <- lrtest(lm.t.mpe.U, lm.ta.mpe.U)
lrt.ta.mpe.U # NOT significant
# Is a temporal covariate signifiant
lrt.t.mpe.U <- lrtest(lm.mpe.U, lm.t.mpe.U)
lrt.t.mpe.U # NOT significant
# Interpretation - There is not significant temporal trend in the Updated Series

# save Ratio P Values
table.temporal.results$Ratio.p.value[table.temporal.results$Metric == "MPE"] = 
  round2(c(lrt.t.mpe.D$`Pr(>Chisq)`[2], lrt.ta.mpe.D$`Pr(>Chisq)`[2], 
         lrt.t.mpe.U$`Pr(>Chisq)`[2], lrt.ta.mpe.U$`Pr(>Chisq)`[2]), 3)

# MAPE
# Default 1.8 Series tests
# Is the interaction with abundance bins significant
lrt.ta.mape.D <- lrtest(lm.t.mape.D, lm.ta.mape.D)
lrt.ta.mape.D # NOT significant
# Is a temporal covariate significant
lrt.t.mape.D <- lrtest(lm.mape.D, lm.t.mape.D)
lrt.t.mape.D # significant
# Updated Series Tests
# Is the interaction with abundance bins significant
lrt.ta.mape.U <- lrtest(lm.t.mape.U, lm.ta.mape.U)
lrt.ta.mape.U # NOT significant
# Is a temporal covariate significant
lrt.t.mape.U <- lrtest(lm.mape.U, lm.t.mape.U)
lrt.t.mape.U # significant

# Save P values
table.temporal.results$Ratio.p.value[table.temporal.results$Metric == "MAPE"] = 
  round2(c(lrt.t.mape.D$`Pr(>Chisq)`[2], lrt.ta.mape.D$`Pr(>Chisq)`[2], 
         lrt.t.mape.U$`Pr(>Chisq)`[2], lrt.ta.mape.U$`Pr(>Chisq)`[2]), 3)

table.temporal.results <- table.temporal.results %>% 
  mutate(Ratio.p.value = ifelse(Ratio.p.value == 0, "< 0.001",
                                Ratio.p.value))

# Interpretation - There is no significant temporal trend within abundance bins
# But there is a temporal trend on Average for MAPE in both the Default and Updated.
# 
# Save Table of Results
write.csv(table.temporal.results, file = "./tables/03_03-Temporal-Trend-Regressions-LRT.csv",
          row.names = F)

```

# Plots
## By Model
```{r}
ggplot(cal.out.long,
       aes(x = hpe, y = per.error,
           colour = type)) +
  geom_point() +
  scale_colour_manual(name = "Estimate Type",
                    values = c("salmon", "grey"),
                    labels = c("Default", "Updated")) +
  geom_hline(yintercept = 0) +
  facet_wrap(calibration.model~abin, scales = "free_x")

ggplot(cal.out.long,
       aes(x = abin, y = per.error,
           fill = type)) +
  geom_boxplot() +
  scale_fill_manual(name = "Estimate Type",
                    values = c("salmon", "grey"),
                    labels = c("Default", "Updated")) +
  geom_hline(yintercept = 0) +
  facet_wrap(~calibration.model, scales = "free_x") +
  labs(x = "High Accuracy Abundance Bins", y = "Percent Error")

ggsave("./figures/03_Model_Performance/Percent_Error_By_Model_By_Abundance_Bins.png",
       height = 7, width = 7, dpi = 95)
```

## Early Stuart
```{r}
estu <- cal.out.long %>% 
  filter(run_timing_group == "Early Stuart")

cal.estu <- cal %>% filter(run_timing_group == "Early Stuart")

cal %>% filter(size == "Medium") %>% count(lpe_method, water_clarity)

count(estu, type, gazetted_stream_name)

count(estu, type, size, lpe_method, water_clarity)
count(estu, type, water_clarity, calibration.model)

estu %>% 
  filter(type == "historical") %>% 
  ggplot(aes(x = size, y = hpe/lpe, fill = size, colour = water_clarity)) +
  geom_hline(yintercept = 1.8, col = "red", alpha = 0.5) +
  geom_boxplot() +
  scale_fill_manual(values = "#D55E00") +
  scale_colour_manual(values = c("gray","black")) +
  facet_wrap(~lpe_method) +
  labs(y = "High Accuracy / Peak Count")

ggsave("./figures/03_Model_Performance/Early_Stuart_Story.png",
       height = 4, width = 5, dpi = 95)

estu.summary <- estu %>% 
  group_by(abin, type, water_clarity) %>% 
  summarise(MPE = mean(per.error),
            #Abs.MPE = abs(MPE),
            MAPE = mean(abs.per.error),
            RMSE = sqrt(mean(squared.error)),
            R.Cor = cor(estimate, hpe, method = "pearson"),
            sample = n()) ;estu.summary

estu %>% 
  group_by(type) %>% 
  summarise(MPE = mean(per.error),
            #Abs.MPE = abs(MPE),
            MAPE = mean(abs.per.error),
            RMSE = sqrt(mean(squared.error)),
            R.Cor = cor(estimate, hpe, method = "pearson"),
            sample = n(),
            ratio = mean(hpe/lpe),
            ratio.m = median(hpe/lpe))
```

## Summary Figures
### MPE and MAPE
```{r}
# combining model results by adding a field for water clarity to the combined one
lm.mpe.dat$water_clarity <- "Combined"
lm.mpe.dat.combined <- rbind(lm.mpe.wc.dat, lm.mpe.dat)
# combining model results by adding a field for water clarity to the combined one
lm.mape.dat$water_clarity <- "Combined"
lm.mape.dat.combined <- rbind(lm.mape.wc.dat, lm.mape.dat)

p.dodge <- 0.6
err.bar.wd <- 0.4

p.MPE.combined <- ggplot(lm.mpe.dat.combined,
       aes(x = abin, y = fit, shape = water_clarity, colour = type)) +
  geom_hline(yintercept = 0, colour = "grey15", alpha = 0.5) +
  geom_point(position = position_dodge(width = p.dodge), size = 1.75) +
  geom_errorbar(aes(ymin = fit-se.fit, ymax = fit+se.fit), 
                position = position_dodge(width = p.dodge), width = err.bar.wd) +
  scale_colour_manual(name = "Estimate Type",
                    values = c("salmon", "grey"),
                    labels = c("Default", "Updated")) +
  scale_shape_manual(name = "Water Clarity",
                     values = c(15, 16, 17),
                    labels = c("Clear", "Combined", "Turbid/Tannic")) +
  labs(x = "High Accuracy Abundance Estimate Bins", y = "Fit of MPE") +
  ylim(c(-0.52, 0.5))

p.MAPE.combined <- ggplot(lm.mape.dat.combined,
       aes(x = abin, y = fit, shape = water_clarity, colour = type)) +
  geom_point(position = position_dodge(width = p.dodge), size = 1.75) +
  geom_errorbar(aes(ymin = fit-se.fit, ymax = fit+se.fit), 
                position = position_dodge(width = p.dodge), width = err.bar.wd) +
  scale_colour_manual(name = "Estimate Type",
                    values = c("salmon", "grey"),
                    labels = c("Default", "Updated")) +
  scale_shape_manual(name = "Water Clarity",
                     values = c(15, 16, 17),
                    labels = c("Clear", "Combined", "Turbid/Tannic")) +
  labs(x = "High Accuracy Abundance Estimate Bins", y = "Fit of MAPE") +
  ylim(c(0,0.5))

p.MPE.combined + p.MAPE.combined +
    plot_annotation(tag_levels = 'A') +
  plot_layout(ncol = 2, guides = "collect",
              axes = "collect") &
  theme(legend.position='right', plot.tag = element_text(size = 10),
        legend.title = element_text(face = "bold"))

ggsave("./figures/03_Model_Performance/F1_Fitted_Performance_Metrics_By_Abun_and_WC.png",
       height = 4, width = 9, dpi = 95)
```


### Performance relative to 1.8
```{r}
options(scipen=10000)
# natural Calibrated vs High Accuracy
f1a <- ggplot(cal.out,
       aes(x = hpe,
           y = calibrated.estimate)) +
  geom_vline(xintercept = 75000, alpha = 0.25, lty = 2) +
  geom_vline(xintercept = 25000, alpha = 0.25, lty = 2) +
  geom_point(aes(y = classic), alpha = 0.25, colour = "salmon") +
  geom_abline(slope = 1, intercept = 0, colour = "gray") +
  geom_point(alpha = 0.5) +
  geom_errorbar(aes(ymin = calibrated.se.lower, ymax = calibrated.se.upper),
                alpha = 0.25) +
  scale_x_continuous(labels = scales::comma) +
  scale_y_continuous(labels = scales::comma) +
  labs(x = "High Accuracy Estimates",
       y = "Caibrated Estimates \n and Standard Errors")

# logarithmic Calibrated vs High Accuracy
f1b <- ggplot(cal.out,
       aes(x = log(hpe),
           y = log(calibrated.estimate))) +
  geom_vline(xintercept = log(75000), alpha = 0.25, lty = 2) +
  geom_vline(xintercept = log(25000), alpha = 0.25, lty = 2) +
  geom_point(aes(y = log(classic)), alpha = 0.25, colour = "salmon") +
  geom_abline(slope = 1, intercept = 0, colour = "gray") +
  geom_point(alpha = 0.5) +
  geom_errorbar(aes(ymin = log(calibrated.se.lower), ymax = log(calibrated.se.upper)),
                alpha = 0.25) +
  labs(x = "Natural Log of High Accuracy Estimates",
       y = "Natural Log of Caibrated Estimates \n and Standard Errors")

# Classic Expansion Residuals
f1c <- ggplot(cal.out,
              aes(x = log(hpe),
                  y = classic.per.error)) +
  geom_vline(xintercept = log(75000), alpha = 0.25, lty = 2) +
  geom_vline(xintercept = log(25000), alpha = 0.25, lty = 2) +
  geom_hline(yintercept = 0, colour = "gray") +
  geom_hline(yintercept = c(0.25, -0.25), alpha = 0.25, col = "grey") +
  geom_hline(yintercept = c(0.5, -0.5), alpha = 0.25, col = "blue") +
  geom_point(alpha = 0.5, col = "salmon") +
  ylim(c(-2.5,2.5)) +
  labs(x = "Natural Log of High Accuracy Estimates",
       y = "Percent Error of Classic Expansion")

# New Calibration Residuals
f1d <- ggplot(cal.out,
              aes(x = log(hpe),
                  y = per.error)) +
  geom_vline(xintercept = log(75000), alpha = 0.25, lty = 2) +
  geom_vline(xintercept = log(25000), alpha = 0.25, lty = 2) +
  geom_hline(yintercept = 0, colour = "gray") +
  geom_hline(yintercept = c(0.25, -0.25), alpha = 0.25, col = "grey") +
  geom_hline(yintercept = c(0.5, -0.5), alpha = 0.25, col = "blue") +
  geom_point(alpha = 0.5) +
  ylim(c(-2.5,2.5)) +
  labs(x = "Natural Log of High Accuracy Estimates",
       y = "Percent Error of New Expansion")

(f1a + f1b) / (f1c + f1d) + plot_annotation(tag_levels = 'A') +
  theme(plot.tag = element_text(size = 6))

ggsave("./figures/03_Model_Performance/F2_CalibratedvHPE_Calibration_Set_Results.png",
       height = 8, width = 10, dpi = 150)
```